---
title: "Integrated Analysis of All Samples"
author: "Altair Wei"
date: '2022-10-24'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
reactable::reactable(iris)
source("../scripts/LoadUtils.R", chdir = TRUE)
```

## Load Datasets

```{r}
obj <- readRDS(Sys.glob(
  "../results/ObjectCache/IntegrationAndAnnotation/obj_annotated_*.rds"))
```

## Marker Genes of Cell Populations

### Identify Global Marker Genes

```{r}
marker_cosg <- COSG::cosg(
 obj,
 groups = 'all',
 assay = 'RNA',
 slot = 'data',
 mu = 1,
 n_genes_user = 300
)
```

```{r}
markers <- formatCosgTable(marker_cosg)
```

### List of Marker Genes {.tabset}

```{r message=FALSE, warning=FALSE}
anno <- brookite::pull_annotation(
  unique(markers$names),
  mart_info = list(
    mart = "plants_mart",
    dataset = "taestivum_eg_gene",
    host = "https://plants.ensembl.org"
  ),
  dest_attrs = c(
    Description = "description",
    GO_ID = "go_id",
    GO_Name = "name_1006",
    GO_Level = "namespace_1003"
  ))

markers_list <- dplyr::left_join(markers, anno, by = c("names" = "ensembl_gene_id")) |>
  dplyr::select(names, scores, cluster, Description, GO_Name)
markers_list <- split(markers_list, markers_list$cluster)
```

```{r results='asis'}
for (cls in names(markers_list)) {
  df <- markers_list[[cls]]
  df_id <- paste0("markers_anno_table_", cls)
  cat("####", cls, "\n\n")
  print(
    htmltools::tagList(
      htmltools::tags$button(
        "Download as CSV",
        onclick = sprintf(
          "Reactable.downloadDataCSV('%s', '%s')",
          df_id, df_id)
      ),
      reactable::reactable(
        df,
        defaultPageSize = 5,
        elementId = df_id
      )
    )
  )
}
```


### Expression Projection {.tabset}

```{r fig.height=6, fig.width=9, results='asis'}
for (cls in names(markers_list)) {
  df <- markers_list[[cls]] |> dplyr::select(names, scores)
  df_id <- paste0("markers_anno_table_", cls)
  cat("####", cls, "\n\n")
  tops <- suppressMessages(
    dplyr::arrange(df, dplyr::desc(scores)) |>
    dplyr::top_n(6) |>
    dplyr::pull(names))
  p <- Seurat::FeaturePlot(obj, features = tops, order = TRUE,
                           reduction = "umap", ncol = 3) &
    ggplot2::coord_fixed()
  print(p)
  cat("\n\n")
}
```

## Local Markers of Related Subpopulations

### Characteristic of mesophylls

#### Expression distribution

```{r}
avg_expr_df <- Seurat::AverageExpression(obj, slot = "data")$RNA |>
  log1p() |>
  as.data.frame()
```

```{r}
vip_anno <- c(
  "TraesCS1B02G406200" = "N-like",
  "TraesCS1B02G413100" = "N-like",
  "TraesCS1D02G183600" = "N-like",
  "TraesCS2A02G242400" = "Transposase",
  "TraesCS2A02G490900" = "RuBisCO",
  "TraesCS3A02G405700" = "N-like",
  "TraesCS3A02G430500" = "N-like",
  "TraesCS4A02G272300" = "N-like",
  "TraesCS3B02G186100" = "RuBisCO",
  "TraesCS3B02G523500" = "RuBisCO",
  "TraesCS5A02G165400" = "RuBisCO",
  "TraesCS5B02G039000" = "Transposase",
  "TraesCS5B02G423500" = "N-like",
  "TraesCS5D02G010200" = "RuBisCO",
  "TraesCS5D02G169900" = "RuBisCO",
  "TraesCS5D02G425100" = "RuBisCO",
  "TraesCS6B02G402800" = "Transposase",
  "TraesCS7D02G255100" = "N-like",
  "TraesCSU02G243900"  = "N-like",
  "rrn23-2" = "chloroplast rRNA",
  "TraesCS3B02G563300" = "psbM",
  "TraesCS1B02G404500" = "psbZ",
  "TraesCS2D02G573100" = "psbZ",
  "TraesCS2D02G573000" = "psbC",
  "TraesCS5D02G008300" = "psbD",
  "TraesCS5D02G456600" = "Ribosomal protein S3",
  "TraesCS5D02G010000" = "Ycf4",
  "TraesCS2D02G566300" = "Ribosomal protein S3",
  "TraesCS2B02G487300" = "Ribosomal protein S14"
)
```

```{r}
vip_genes <- avg_expr_df[names(vip_anno), c("Me_1", "Me_2")]
vip_genes$anno <- vip_anno
```

```{r fig.height=7, fig.width=7}
avg_expr_df$gene <- rownames(avg_expr_df)
p <- ggplot2::ggplot(avg_expr_df, ggplot2::aes(x = Me_2, y = Me_1, label = gene)) +
  ggplot2::geom_point() +
  ggplot2::geom_abline(intercept = 0, slope = 1, color = "blue") +
  #ggplot2::geom_abline(intercept = 0, slope = 0.6, color = "blue") +
  #ggplot2::geom_hline(yintercept = 3.5, color = "blue") +
  #ggplot2::geom_vline(xintercept = 1.5, color = "blue") +
  ggrepel::geom_text_repel(
    data = vip_genes,
    mapping = ggplot2::aes(x = Me_2, y = Me_1, label = anno),
    size = 2, bg.color = "white") +
  cowplot::theme_cowplot()

p
```

#### Differential expression

```{r}
de_chl <- xfun::cache_rds(
  file = "de_marker_chlorenchyma.rds",
  dir = "../results/ObjectCache/IntegrationAndAnnotation/",
  expr = local({
    chlorenchyma <- c("Me_1", "Me_2", "Me_3", "Va_1")
    names(chlorenchyma) <- chlorenchyma
    lapply(chlorenchyma, function(cls) Seurat::FindMarkers(
        obj, ident.1 = cls, ident.2 = chlorenchyma[chlorenchyma != cls],
        min.pct = 0.25, only.pos = TRUE)
    )
  })
)
```

#### Enrichment of local markers {.tabset}

```{r}
me_marker_list <- lapply(de_chl,
  function(df) rownames(df[df$p_val_adj < 0.01,]))

ora_me_marker <- lapply(
  X = c(BP = "BP", MF = "MF", CC = "CC"),
  FUN = function(ont)
    clusterProfiler::compareCluster(
      me_marker_list,
      fun = "enrichGO",
      OrgDb = "org.Taestivum.iwgsc.db",
      keyType = "GID",
      ont = ont,
      universe = rownames(obj)
    )
)
```

##### Biological process

```{r eval = FALSE}
p <- ora_me_marker$BP |>
  enrichplot::pairwise_termsim(
    showCategory = NULL) |>
  enrichplot::treeplot(
    showCategory = NULL,
    size = "count",
    clusterPanel.params = list(
      clusterPanel = "dotplot"),
    cluster.params = list(
      n = 12, label_words_n = 0, label_format = 5),
    offset.params = list(bar_tree = 60, tiplab = 32),
    hilight.params = list()) +
  ggplot2::scale_y_continuous(
    expand = ggplot2::expansion(mult = c(0, 0), add = c(15, 0))) +
  NULL

plotpowerpoint(
  ggobj = p,
  file = "../results/DataPlots/local_marker_ora_me_marker.pptx",
  template = "../data/60x120.pptx")
```

```{r}
reprterms <- ora_me_marker$BP@compareClusterResult |>
  dplyr::select(Description, ID) |>
  dplyr::distinct() |>
  tibble::deframe()

reprterms <- reprterms[
  extractMarkedTerms("../results/DataPlots/ds_ora_comp_1dpi.pptx")]
```

```{r fig.height=10, fig.width=10}
ora_me_marker$BP |>
  dplyr::filter(ID %in% reprterms) |>
  enrichplot::dotplot(
    showCategory = NULL,
    size = "count", label_format = 50)
```

##### Molecular function

```{r fig.height=12, fig.width=10}
ora_me_marker$MF |>
  enrichplot::pairwise_termsim(showCategory = NULL) |>
  enrichRepresent() |>
  enrichplot::dotplot(
    showCategory = NULL,
    size = "count", label_format = 50)
```

##### Cellular component

```{r fig.height=6, fig.width=10}
ora_me_marker$CC |>
  enrichplot::pairwise_termsim(showCategory = NULL) |>
  enrichRepresent() |>
  enrichplot::dotplot(
    showCategory = NULL,
    size = "count", label_format = 50)
```

### Characteristic of outer sheath

#### Differential expression

```{r}
de_ospa <- xfun::cache_rds(
  file = "de_marker_outersheath_parenchyma.rds",
  dir = "../results/ObjectCache/IntegrationAndAnnotation/",
  expr = local({
    ospa <- c("Va_2", "Va_3", "Va_4", "Me_4", "Me_5", "Me_6")
    names(ospa) <- ospa
    lapply(ospa, function(cls) Seurat::FindMarkers(
        obj, ident.1 = cls, ident.2 = ospa[ospa != cls],
        min.pct = 0.25, only.pos = TRUE)
    )
  })
)
```

#### Enrichment of local markers {.tabset}

```{r}
va_marker_list <- lapply(de_ospa,
  function(df) rownames(df[df$p_val_adj < 0.01,]))

ora_va_marker <- lapply(
  X = c(BP = "BP", MF = "MF", CC = "CC"),
  FUN = function(ont)
    clusterProfiler::compareCluster(
      va_marker_list,
      fun = "enrichGO",
      OrgDb = "org.Taestivum.iwgsc.db",
      keyType = "GID",
      ont = ont,
      universe = rownames(obj)
    )
)
```

##### Biological proess

```{r eval = FALSE}
p <- ora_va_marker$BP |>
  enrichplot::pairwise_termsim(
    showCategory = NULL) |>
  enrichplot::treeplot(
    showCategory = NULL,
    size = "count",
    clusterPanel.params = list(
      clusterPanel = "dotplot"),
    cluster.params = list(
      n = 12, label_words_n = 0, label_format = 5),
    offset.params = list(bar_tree = 48, tiplab = 50),
    hilight.params = list()) +
  ggplot2::scale_y_continuous(
    expand = ggplot2::expansion(mult = c(0, 0), add = c(15, 0))) +
  NULL

plotpowerpoint(
  ggobj = p,
  file = "../results/DataPlots/local_marker_ora_va_marker.pptx",
  template = "../data/60x140.pptx")
```

```{r}
reprterms <- ora_va_marker$BP@compareClusterResult |>
  dplyr::select(Description, ID) |>
  dplyr::distinct() |>
  tibble::deframe()

reprterms <- reprterms[
  extractMarkedTerms("../results/DataPlots/local_marker_ora_va_marker.pptx")]
```

```{r fig.height=16, fig.width=8}
ora_va_marker$BP |>
  dplyr::filter(ID %in% reprterms) |>
  enrichplot::dotplot(
    showCategory = NULL,
    size = "count", label_format = 50)
```

```{r fig.height=16, fig.width=16}
ora_va_marker$BP |>
  dplyr::filter(ID %in% reprterms) |>
  enrichplot::pairwise_termsim(
    showCategory = NULL) |>
  enrichplot::emapplot(
    showCategory = NULL,
    font.size = 14,
    node_label = "category",
    pie.params = list(legend_n = 3, pie = "Count"),
    layout.params = list(layout = "fr"),
    edge.params = list(min = 0.01),
    cex.params = list(category_label = 1.6)) +
  ggplot2::guides(fill = ggplot2::guide_legend(title = "Cell Clusters")) +
  ggplot2::scale_fill_brewer(palette = "BrBG")
```

##### Molecular function

```{r fig.height=12, fig.width=10}
ora_me_marker$MF |>
  enrichplot::pairwise_termsim(showCategory = NULL) |>
  enrichRepresent() |>
  enrichplot::dotplot(
    showCategory = NULL,
    size = "count", label_format = 50)
```

##### Cellular component

```{r fig.height=6, fig.width=10}
ora_me_marker$CC |>
  enrichplot::pairwise_termsim(showCategory = NULL) |>
  enrichRepresent() |>
  enrichplot::dotplot(
    showCategory = NULL,
    size = "count", label_format = 50)
```

```{r}
knitr::knit_exit()
```

### Combined ORA visualization

```{r}
marker_df <- purrr::imap(
    .x = list(
      "Mesophyll" = me_marker_list,
      "Outer sheath" = va_marker_list),
    .f = function(malist, uppername) {
      purrr::imap(
          .x = malist,
          .f = function(genes, name) data.frame(
            gene = genes, cell_type = name, tissue = uppername)) |>
        dplyr::bind_rows()
    }) |>
  dplyr::bind_rows()
```

```{r}
ego <- clusterProfiler::compareCluster(
  gene ~ tissue + cell_type, data = marker_df,
  OrgDb = orgdb, fun = "enrichGO",
  universe = rownames(obj),
  ont = "BP", keyType = "GID")
```

```{r fig.height=18, fig.width=14}
selected_terms <- unique(c(selected_me_go_terms, selected_os_go_terms))

p <- ego |>
  dplyr::filter(ID %in% selected_terms) |>
  enrichplot::dotplot(
    showCategory = length(selected_terms),
    x = "cell_type", size = "count",
    label_format = 100, font.size = 14) +
  ggplot2::facet_grid(~ tissue, scales = "free_x") +
  ggplot2::xlab("cell cluster") +
  ggplot2::theme(strip.text = ggplot2::element_text(size = 14)) +
  NULL

p
```

```{r fig.height=12, fig.width=14}
p <- ego |>
  dplyr::filter(ID %in% selected_terms) |>
  enrichplot::pairwise_termsim() |>
  enrichplot::emapplot(
    showCategory = length(selected_terms),
    node_label = "category", layout = "fr",
    legend_n = 3, font.size = 14, min_edge = 0.01,
    cex_label_category = 1.6) +
  ggthemes::scale_fill_tableau(
    palette = "Tableau 20",
    labels = c("Me_1", "Me_2", "Va_2", "Va_3", "Va_4")) +
  ggplot2::guides(fill = ggplot2::guide_legend(title = "cell cluster")) +
  ggplot2::theme_void(14)

p
```

