---
title: "Co-varying Neighborhood Analysis"
author: "Altair Wei"
date: "2022/5/15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, autodep = TRUE)
library(magrittr)
library(Matrix)
library(ggplot2) # Fix {ggtreeExtra} error
source("../scripts/AbundanceUtilities.R")
source("../scripts/EnrichmentUtilities.R")
source("../scripts/TrajectoryUtilities.R")
```

## 算法探索

- `batch` 是一个比 replicates 更大的概念。
- 目前 {CNA} 不支持多元无序变量。
- `ncorrs` is the vector of neighborhood coefficients
- 我觉得应该将数据集拆分，而不是应用 `isPNR2` 或者 `isTR4` 这种类型的二元变量。

## 加载数据

```{r}
obj <- readRDS(Sys.glob("../results/ObjectCache/IntegrationAndAnnotation/obj_annotated_*.rds"))
```

```{r message=FALSE, warning=FALSE, results="hide"}
library(org.Taestivum.iwgsc.db)
orgdb <- org.Taestivum.iwgsc.db
```

```{r}
obj$cellType <- Seurat::Idents(obj)
sce <- Seurat::as.SingleCellExperiment(obj, assay = "RNA")
sce <- muscat::prepSCE(sce,
  kid = "ident", # subpopulation assignments
  gid = "group", # group IDs (ctrl/stim)
  sid = "sample", # sample IDs (ctrl/stim.1234)
  drop = FALSE
)
```

```{r}
rm(obj)
gc()
```

## MOCK 发育规律

```{r}
sce_mock <- sce[, sce$treatment == "MOCK"]

sce_mock$sample_id <- droplevels(sce_mock$sample_id)
sce_mock$group_id <- droplevels(sce_mock$group_id)
sce_mock$cluster_id <- droplevels(sce_mock$cluster_id)
```

### 降维可视化 {.tabset}

```{r}
sce_mock <- runPHATE(sce_mock, dimred = "HARMONY", seed = 1130L)
```

#### UMAP

```{r}
scater::plotReducedDim(sce_mock, dimred = "UMAP",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```


#### t-SNE

```{r}
scater::plotReducedDim(sce_mock, dimred = "TSNE",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

#### PHATE

```{r}
scater::plotReducedDim(sce_mock, dimred = "PHATE",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

### 时间序列降维 {.tabset}

#### UMAP

```{r fig.height=7, fig.width=28}
scater::plotReducedDim(
    sce_mock, dimred = "UMAP",
    colour_by = "cellType",
    text_by = "cellType", text_size = 5,
    other_fields = "time") +
  ggplot2::coord_fixed() +
  ggplot2::facet_wrap(~ time, nrow = 1) +
  ggplot2::guides(color = ggplot2::guide_legend(override.aes = list(size = 4))) +
  NULL
```

#### t-SNE

```{r fig.height=7, fig.width=28}
scater::plotReducedDim(
    sce_mock, dimred = "TSNE",
    colour_by = "cellType",
    text_by = "cellType", text_size = 5,
    other_fields = "time") +
  ggplot2::coord_fixed() +
  ggplot2::facet_wrap(~ time, nrow = 1) +
  ggplot2::guides(color = ggplot2::guide_legend(override.aes = list(size = 4))) +
  NULL
```

### 运行 CNA 方法

关于有序离散变量：https://github.com/immunogenomics/cna/issues/12

> Continuous variables should work fine -- just plug them in. Ordered categorical variables are a bit trickier -- you can always just code the categories as {1,2,3...} and plug them in since they're ordered, but be aware that they will be fit via a linear model so the assumption is that whatever enrichment differentiates category 1 from category 2 will also equally differentiate category 2 from category 3.

CNA 假设有序离散变量之间的富集差异是线性的。

```{r}
sce_mock$time_val <- as.numeric(factor(
  sce_mock$time, levels = c("0DPI", "1DPI", "2DPI", "3DPI")))

cnaObj <- createCNAObject(
  sce_mock, 
  samplem_key = "sample_id",
  samplem_vars = c(
    "time", "time_val"
  )
)

cnaRes <- rcna::association(
  data = cnaObj,
  y = cnaObj$samplem$time_val,
  batches = NULL,
  return_nam = TRUE
)

sce_mock$time_ncorrs <- cnaRes$ncorrs[colnames(sce_mock), , drop=TRUE]
sce_mock$time_ncorrs_fdr05 <- rep(0, ncol(sce_mock))
if (!is.null(cnaRes$fdr_5p_t)) {
    idx_passed <- which(abs(sce_mock$time_ncorrs) >= cnaRes$fdr_5p_t)
    sce_mock$time_ncorrs_fdr05[idx_passed] <- sce_mock$time_ncorrs[idx_passed]
}

sce_mock$time_ncorrs_fdr10 <- rep(0, ncol(sce_mock))
if (!is.null(cnaRes$fdr_10p_t)) {
    idx_passed <- which(abs(sce_mock$time_ncorrs) >= cnaRes$fdr_10p_t)
    sce_mock$time_ncorrs_fdr10[idx_passed] <- sce_mock$time_ncorrs[idx_passed]
}
```



### 表型关联结果 {.tabset}

```{r}
blue_to_orange <- rev(ggthemes::ggthemes_data[[
  c("tableau", "color-palettes", "ordered-diverging", "Orange-Blue Diverging")]][["value"]])
```

在下面的图中，每个细胞根据其邻域系数着色，即颜色对应于每个细胞相应邻域的丰度与样本属性的相关性。由于我们将时间编码成了数字，因此从蓝色到橙红色代表了 0DPI ~ 3DPI 的邻域系数。

#### UMAP

```{r}
scater::plotReducedDim(
    sce_mock[, sample(seq_len(ncol(sce_mock)))],
    dimred = "UMAP",
    point_size = 0.5,
    colour_by = "time_ncorrs",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::scale_color_gradientn(colours = blue_to_orange) +
  ggplot2::coord_fixed()
```

#### t-SNE

```{r}
scater::plotReducedDim(
    sce_mock[, sample(seq_len(ncol(sce_mock)))],
    dimred = "TSNE",
    point_size = 0.5,
    colour_by = "time_ncorrs",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::scale_color_gradientn(colours = blue_to_orange) +
  ggplot2::coord_fixed()
```

#### PHATE

```{r}
scater::plotReducedDim(
    sce_mock[, sample(seq_len(ncol(sce_mock)))],
    dimred = "PHATE",
    point_size = 0.5,
    colour_by = "time_ncorrs",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::scale_color_gradientn(colours = blue_to_orange) +
  ggplot2::coord_fixed()
```

### 邻域系数

#### 细胞群体分布

```{r message=FALSE, warning=FALSE, dev="svg"}
data.frame(time_ncorrs = sce_mock$time_ncorrs,
           cellType = sce_mock$cellType) |>
  ggplot2::ggplot(ggplot2::aes(
    x = time_ncorrs, y = forcats::fct_reorder(cellType, time_ncorrs),
    fill = cellType)) +
  ggplot2::ylab("cellType") +
  ggridges::geom_density_ridges(scale = 4) +
  ggplot2::geom_vline(xintercept = cnaRes$fdr_5p_t, linetype="dashed") +
  ggplot2::geom_vline(xintercept = -cnaRes$fdr_5p_t, linetype="dashed") +
  ggplot2::scale_y_discrete(limits=rev, expand = ggplot2::expansion(add = c(0, 4))) +
  ggridges::theme_ridges()
```

#### 关联基因富集分析

```{r}
ncorrs <- as(matrix(cnaRes$ncorrs, ncol = 1), "dgCMatrix")
genes <- qlcMatrix::corSparse(ncorrs, t(logcounts(sce_mock)))
genes <- structure(as.numeric(genes), names = rownames(logcounts(sce_mock)))
genes <- sort(genes, decreasing = TRUE)
```

```{r}
gsea <- clusterProfiler::gseGO(
  genes, ont = "BP",
  OrgDb = org.Taestivum.iwgsc.db,
  keyType = "GID")

gsea <- clusterProfiler::simplify(gsea, cutoff = 0.5)
```

```{r fig.height=10, fig.width=10, dev="svg"}
p <- gsea |>
  enrichplot::pairwise_termsim() |>
  enrichplot::treeplot(
    showCategory = 40, nCluster = 12,
    offset = 20, offset_tiplab = 1,
    label_format_cladelab = 5, color = "NES",
    geneClusterPanel = "dotplot", nWords = 0) +
  ggplot2::scale_y_continuous(expand = ggplot2::expansion(mult = c(0, 0), add = c(6, 0))) +
  ggplot2::scale_color_gradient2(low = "blue", mid = "white", high = "red", name = "NES") +
  ggplot2::coord_cartesian()
p
```

```{r}
printGSEATable(gsea)
```

```{r fig.height=12, fig.width=14}
enrichplot::ridgeplot(gsea, showCategory = 100, label_format = 100) +
  ggplot2::ggtitle("Correlation distributions of core enriched genes") +
  ggplot2::xlab("Pearson's R")
```

### 邻域负荷

#### 方差解释

查看源代码可知，`NAM_nbhdXpc` 等于 `NAM_embeddings`，而 `NAM_sampleXpc` 等于 `NAM_loadings` 。

```{r}
data.frame(sig2 = cnaRes$NAM_svs / sum(cnaRes$NAM_svs)) |>
  tibble::rowid_to_column('PC') |>
  ggplot2::ggplot(ggplot2::aes(PC, sig2)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme_classic(base_size = 14) +
  ggplot2::labs(y = 'Fraction of variance explained', x = 'NAM PC') +
  ggplot2::scale_x_continuous(breaks = 1:20)
```

#### 细胞投影

```{r}
nampcs <- paste0("NAM", colnames(cnaRes$NAM_embeddings))
colData(sce_mock)[, nampcs] <- as.data.frame(cnaRes$NAM_embeddings)
```

```{r fig.height=8, fig.width=16}
lapply(nampcs, function(pcname) {
  scater::plotReducedDim(
    sce_mock,
    dimred = "UMAP",
    point_size = 0.5,
    colour_by = pcname,
    text_by = "cellType",
    text_size = 3,
    show_violin = TRUE) +
  ggplot2::scale_color_gradientn(colours = blue_to_orange) +
  ggplot2::ggtitle(pcname) +
  ggplot2::coord_fixed()
}) |> patchwork::wrap_plots(nrow = 2)
```

### 样本负荷

#### 样本间的距离

```{r}
rownames(cnaRes$NAM_loadings) <- as.character(cnaObj$samplem[[cnaObj$samplem_key]])

df_to_plot <- data.frame(
  PC1 = cnaRes$NAM_loadings[, 1],
  PC2 = cnaRes$NAM_loadings[, 2],
  PC3 = cnaRes$NAM_loadings[, 3],
  Time = cnaObj$samplem$time
)

plotly::plot_ly(
  df_to_plot,
  x = ~ PC1, y = ~ PC2, z = ~ PC3,
  color = ~ Time,
  #colors = cellTypeColors,
  type = "scatter3d",
  mode = "markers",
  marker = list(size = 4)
)
```

#### 细胞类型丰度关联

> Correlation across samples between the sample loadings for each NAM-PC and the abundances of each of the five main cell types in the dataset.

```{r}
cluster_prop <- t(prop.table(table(sce_mock$cluster_id, sce_mock$sample_id), 2))
cluster_prop
```

```{r}
corres <- psych::corr.test(
  x = cnaRes$NAM_loadings[rownames(cluster_prop),],
  y = cluster_prop, method = "pearson"
)
ComplexHeatmap::Heatmap(
    corres$r,
    name = "Pearson \nCorrelation",
    col = c("blue", "white", "red"),
    row_names_side = "right",
    row_dend_side = "left",
    column_names_side = "bottom",
    column_dend_side = "top",
    column_names_rot = 45,
    cell_fun = function(j, i, x, y, width, height, fill) {
      if (corres$p[i, j] < 0.05)
        grid::grid.text(
          sprintf("%.2f", corres$r[i, j]),
          x, y, gp = grid::gpar(fontsize = 10)
        )
    }
  )
```

## PH-1 侵染反应

```{r}
sce_pnr2 <- sce[, sce$treatment == "PNR2" | sce$time == "0DPI"]

sce_pnr2$sample_id <- droplevels(sce_pnr2$sample_id)
sce_pnr2$group_id <- droplevels(sce_pnr2$group_id)
sce_pnr2$cluster_id <- droplevels(sce_pnr2$cluster_id)
```

### 降维可视化 {.tabset}

#### UMAP

```{r}
scater::plotReducedDim(sce_pnr2, dimred = "UMAP",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

#### t-SNE

```{r}
scater::plotReducedDim(sce_pnr2, dimred = "TSNE",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

### 时间序列降维 {.tabset}

#### UMAP

```{r fig.height=7, fig.width=28}
scater::plotReducedDim(
    sce_pnr2, dimred = "UMAP",
    colour_by = "cellType",
    text_by = "cellType", text_size = 5,
    other_fields = "time") +
  ggplot2::coord_fixed() +
  ggplot2::facet_wrap(~ time, nrow = 1) +
  ggplot2::guides(color = ggplot2::guide_legend(override.aes = list(size = 4))) +
  NULL
```

#### t-SNE

```{r fig.height=7, fig.width=28}
scater::plotReducedDim(
    sce_pnr2, dimred = "TSNE",
    colour_by = "cellType",
    text_by = "cellType", text_size = 5,
    other_fields = "time") +
  ggplot2::coord_fixed() +
  ggplot2::facet_wrap(~ time, nrow = 1) +
  ggplot2::guides(color = ggplot2::guide_legend(override.aes = list(size = 4))) +
  NULL
```

### 运行 CNA 方法

```{r}
sce_pnr2$time_val <- as.numeric(factor(
  sce_pnr2$time, levels = c("0DPI", "1DPI", "2DPI", "3DPI")))

cnaObj <- createCNAObject(
  sce_pnr2, 
  samplem_key = "sample_id",
  samplem_vars = c(
    "time", "time_val"
  )
)

cnaRes <- rcna::association(
  data = cnaObj,
  y = cnaObj$samplem$time_val,
  batches = NULL,
  return_nam = TRUE
)

sce_pnr2$time_ncorrs <- cnaRes$ncorrs[colnames(sce_pnr2), , drop=TRUE]
sce_pnr2$time_ncorrs_fdr05 <- rep(0, ncol(sce_pnr2))
if (!is.null(cnaRes$fdr_5p_t)) {
    idx_passed <- which(abs(sce_pnr2$time_ncorrs) >= cnaRes$fdr_5p_t)
    sce_pnr2$time_ncorrs_fdr05[idx_passed] <- sce_pnr2$time_ncorrs[idx_passed]
}

sce_pnr2$time_ncorrs_fdr10 <- rep(0, ncol(sce_pnr2))
if (!is.null(cnaRes$fdr_10p_t)) {
    idx_passed <- which(abs(sce_pnr2$time_ncorrs) >= cnaRes$fdr_10p_t)
    sce_pnr2$time_ncorrs_fdr10[idx_passed] <- sce_pnr2$time_ncorrs[idx_passed]
}
```

### 表型关联结果 {.tabset}

```{r}
blue_to_orange <- rev(ggthemes::ggthemes_data[[
  c("tableau", "color-palettes", "ordered-diverging", "Orange-Blue Diverging")]][["value"]])
```

#### UMAP

```{r}
scater::plotReducedDim(
    sce_pnr2[, sample(seq_len(ncol(sce_pnr2)))],
    dimred = "UMAP",
    point_size = 0.5,
    colour_by = "time_ncorrs",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::scale_color_gradientn(colours = blue_to_orange) +
  ggplot2::coord_fixed()
```
#### t-SNE

```{r}
scater::plotReducedDim(
    sce_pnr2[, sample(seq_len(ncol(sce_pnr2)))],
    dimred = "TSNE",
    point_size = 0.5,
    colour_by = "time_ncorrs",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::scale_color_gradientn(colours = blue_to_orange) +
  ggplot2::coord_fixed()
```

### 邻域系数

#### 细胞群体分布

```{r message=FALSE, warning=FALSE, dev="svg"}
data.frame(time_ncorrs = sce_pnr2$time_ncorrs,
           cellType = sce_pnr2$cellType) |>
  ggplot2::ggplot(ggplot2::aes(
    x = time_ncorrs, y = forcats::fct_reorder(cellType, time_ncorrs),
    fill = cellType)) +
  ggplot2::ylab("cellType") +
  ggridges::geom_density_ridges(scale = 4) +
  ggplot2::geom_vline(xintercept = cnaRes$fdr_5p_t, linetype="dashed") +
  ggplot2::geom_vline(xintercept = -cnaRes$fdr_5p_t, linetype="dashed") +
  ggplot2::scale_y_discrete(limits=rev, expand = ggplot2::expansion(add = c(0, 4))) +
  ggridges::theme_ridges()
```

#### 关联基因富集分析

```{r}
ncorrs <- as(matrix(cnaRes$ncorrs, ncol = 1), "dgCMatrix")
genes <- qlcMatrix::corSparse(ncorrs, t(logcounts(sce_pnr2)))
genes <- structure(as.numeric(genes), names = rownames(logcounts(sce_pnr2)))
genes <- sort(genes, decreasing = TRUE)
```

```{r}
gsea <- clusterProfiler::gseGO(
  genes, ont = "BP",
  OrgDb = org.Taestivum.iwgsc.db,
  keyType = "GID")

gsea <- clusterProfiler::simplify(gsea, cutoff = 0.5)
```

```{r fig.height=10, fig.width=10, dev="svg"}
p <- gsea |>
  enrichplot::pairwise_termsim() |>
  enrichplot::treeplot(
    showCategory = 40, nCluster = 12,
    offset = 20, offset_tiplab = 1,
    label_format_cladelab = 5, color = "NES",
    geneClusterPanel = "dotplot", nWords = 0) +
  ggplot2::scale_y_continuous(expand = ggplot2::expansion(mult = c(0, 0), add = c(6, 0))) +
  ggplot2::scale_color_gradient2(low = "blue", mid = "white", high = "red", name = "NES") +
  ggplot2::coord_cartesian()
p
```

```{r}
printGSEATable(gsea)
```

```{r fig.height=12, fig.width=14}
enrichplot::ridgeplot(gsea, showCategory = 100, label_format = 100) +
  ggplot2::ggtitle("Correlation distributions of core enriched genes") +
  ggplot2::xlab("Pearson's R")
```

### 邻域负荷

#### 方差解释

查看源代码可知，`NAM_nbhdXpc` 等于 `NAM_embeddings`，而 `NAM_sampleXpc` 等于 `NAM_loadings` 。

```{r}
data.frame(sig2 = cnaRes$NAM_svs / sum(cnaRes$NAM_svs)) |>
  tibble::rowid_to_column('PC') |>
  ggplot2::ggplot(ggplot2::aes(PC, sig2)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme_classic(base_size = 14) +
  ggplot2::labs(y = 'Fraction of variance explained', x = 'NAM PC') +
  ggplot2::scale_x_continuous(breaks = 1:20)
```

#### 细胞投影

```{r}
nampcs <- paste0("NAM", colnames(cnaRes$NAM_embeddings))
colData(sce_pnr2)[, nampcs] <- as.data.frame(cnaRes$NAM_embeddings)
```

```{r fig.height=8, fig.width=16}
lapply(nampcs, function(pcname) {
  scater::plotReducedDim(
    sce_pnr2,
    dimred = "UMAP",
    point_size = 0.5,
    colour_by = pcname,
    text_by = "cellType",
    text_size = 3,
    show_violin = TRUE) +
  ggplot2::scale_color_gradientn(colours = blue_to_orange) +
  ggplot2::ggtitle(pcname) +
  ggplot2::coord_fixed()
}) |> patchwork::wrap_plots(nrow = 2)
```

### 样本负荷

#### 样本间的距离

```{r}
rownames(cnaRes$NAM_loadings) <- as.character(cnaObj$samplem[[cnaObj$samplem_key]])

df_to_plot <- data.frame(
  PC1 = cnaRes$NAM_loadings[, 1],
  PC2 = cnaRes$NAM_loadings[, 2],
  PC3 = cnaRes$NAM_loadings[, 3],
  Time = cnaObj$samplem$time
)

plotly::plot_ly(
  df_to_plot,
  x = ~ PC1, y = ~ PC2, z = ~ PC3,
  color = ~ Time,
  #colors = cellTypeColors,
  type = "scatter3d",
  mode = "markers",
  marker = list(size = 4)
)
```

#### 细胞类型丰度关联

> Correlation across samples between the sample loadings for each NAM-PC and the abundances of each of the five main cell types in the dataset.

```{r}
cluster_prop <- t(prop.table(table(sce_pnr2$cluster_id, sce_pnr2$sample_id), 2))
cluster_prop
```

```{r}
corres <- psych::corr.test(
  x = cnaRes$NAM_loadings[rownames(cluster_prop),],
  y = cluster_prop, method = "pearson"
)
ComplexHeatmap::Heatmap(
    corres$r,
    name = "Pearson \nCorrelation",
    col = c("blue", "white", "red"),
    row_names_side = "right",
    row_dend_side = "left",
    column_names_side = "bottom",
    column_dend_side = "top",
    column_names_rot = 45,
    cell_fun = function(j, i, x, y, width, height, fill) {
      if (corres$p[i, j] < 0.05)
        grid::grid.text(
          sprintf("%.2f", corres$r[i, j]),
          x, y, gp = grid::gpar(fontsize = 10)
        )
    }
  )
```

## TR4 侵染反应

```{r}
sce_tr4 <- sce[, sce$treatment == "TR4" | sce$time == "0DPI"]

sce_tr4$sample_id <- droplevels(sce_tr4$sample_id)
sce_tr4$group_id <- droplevels(sce_tr4$group_id)
sce_tr4$cluster_id <- droplevels(sce_tr4$cluster_id)
```

### 降维可视化 {.tabset}

#### UMAP

```{r}
scater::plotReducedDim(sce_tr4, dimred = "UMAP",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

#### t-SNE

```{r}
scater::plotReducedDim(sce_tr4, dimred = "TSNE",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

### 时间序列降维 {.tabset}

#### UMAP

```{r fig.height=7, fig.width=28}
scater::plotReducedDim(
    sce_tr4, dimred = "UMAP",
    colour_by = "cellType",
    text_by = "cellType", text_size = 5,
    other_fields = "time") +
  ggplot2::coord_fixed() +
  ggplot2::facet_wrap(~ time, nrow = 1) +
  ggplot2::guides(color = ggplot2::guide_legend(override.aes = list(size = 4))) +
  NULL
```

#### t-SNE

```{r fig.height=7, fig.width=28}
scater::plotReducedDim(
    sce_tr4, dimred = "TSNE",
    colour_by = "cellType",
    text_by = "cellType", text_size = 5,
    other_fields = "time") +
  ggplot2::coord_fixed() +
  ggplot2::facet_wrap(~ time, nrow = 1) +
  ggplot2::guides(color = ggplot2::guide_legend(override.aes = list(size = 4))) +
  NULL
```

### 运行 CNA 方法

关于有序离散变量：https://github.com/immunogenomics/cna/issues/12

> Continuous variables should work fine -- just plug them in. Ordered categorical variables are a bit trickier -- you can always just code the categories as {1,2,3...} and plug them in since they're ordered, but be aware that they will be fit via a linear model so the assumption is that whatever enrichment differentiates category 1 from category 2 will also equally differentiate category 2 from category 3.

CNA 假设有序离散变量之间的富集差异是线性的。

```{r}
sce_tr4$time_val <- as.numeric(factor(
  sce_tr4$time, levels = c("0DPI", "1DPI", "2DPI", "3DPI")))

cnaObj <- createCNAObject(
  sce_tr4, 
  samplem_key = "sample_id",
  samplem_vars = c(
    "time", "time_val"
  )
)

cnaRes <- rcna::association(
  data = cnaObj,
  y = cnaObj$samplem$time_val,
  batches = NULL,
  return_nam = TRUE
)

sce_tr4$time_ncorrs <- cnaRes$ncorrs[colnames(sce_tr4), , drop=TRUE]
sce_tr4$time_ncorrs_fdr05 <- rep(0, ncol(sce_tr4))
if (!is.null(cnaRes$fdr_5p_t)) {
    idx_passed <- which(abs(sce_tr4$time_ncorrs) >= cnaRes$fdr_5p_t)
    sce_tr4$time_ncorrs_fdr05[idx_passed] <- sce_tr4$time_ncorrs[idx_passed]
}

sce_tr4$time_ncorrs_fdr10 <- rep(0, ncol(sce_tr4))
if (!is.null(cnaRes$fdr_10p_t)) {
    idx_passed <- which(abs(sce_tr4$time_ncorrs) >= cnaRes$fdr_10p_t)
    sce_tr4$time_ncorrs_fdr10[idx_passed] <- sce_tr4$time_ncorrs[idx_passed]
}
```


### 表型关联结果 {.tabset}

```{r}
blue_to_orange <- rev(ggthemes::ggthemes_data[[
  c("tableau", "color-palettes", "ordered-diverging", "Orange-Blue Diverging")]][["value"]])
```

#### UMAP

```{r}
scater::plotReducedDim(
    sce_tr4[, sample(seq_len(ncol(sce_tr4)))],
    dimred = "UMAP",
    point_size = 0.5,
    colour_by = "time_ncorrs",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::scale_color_gradientn(colours = blue_to_orange) +
  ggplot2::coord_fixed()
```

#### t-SNE

```{r}
scater::plotReducedDim(
    sce_tr4[, sample(seq_len(ncol(sce_tr4)))],
    dimred = "TSNE",
    point_size = 0.5,
    colour_by = "time_ncorrs",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::scale_color_gradientn(colours = blue_to_orange) +
  ggplot2::coord_fixed()
```

### 邻域系数

#### 细胞群体分布

```{r message=FALSE, warning=FALSE, dev="svg"}
data.frame(time_ncorrs = sce_tr4$time_ncorrs,
           cellType = sce_tr4$cellType) |>
  ggplot2::ggplot(ggplot2::aes(
    x = time_ncorrs, y = forcats::fct_reorder(cellType, time_ncorrs),
    fill = cellType)) +
  ggplot2::ylab("cellType") +
  ggridges::geom_density_ridges(scale = 4) +
  ggplot2::geom_vline(xintercept = cnaRes$fdr_5p_t, linetype="dashed") +
  ggplot2::geom_vline(xintercept = -cnaRes$fdr_5p_t, linetype="dashed") +
  ggplot2::scale_y_discrete(limits=rev, expand = ggplot2::expansion(add = c(0, 4))) +
  ggridges::theme_ridges()
```

#### 关联基因富集分析

```{r}
ncorrs <- as(matrix(cnaRes$ncorrs, ncol = 1), "dgCMatrix")
genes <- qlcMatrix::corSparse(ncorrs, t(logcounts(sce_tr4)))
genes <- structure(as.numeric(genes), names = rownames(logcounts(sce_tr4)))
genes <- sort(genes, decreasing = TRUE)
```

```{r}
gsea <- clusterProfiler::gseGO(
  genes, ont = "BP",
  OrgDb = org.Taestivum.iwgsc.db,
  keyType = "GID")

gsea <- clusterProfiler::simplify(gsea, cutoff = 0.5)
```

```{r fig.height=10, fig.width=10, dev="svg"}
p <- gsea |>
  enrichplot::pairwise_termsim() |>
  enrichplot::treeplot(
    showCategory = 40, nCluster = 12,
    offset = 20, offset_tiplab = 1,
    label_format_cladelab = 5, color = "NES",
    geneClusterPanel = "dotplot", nWords = 0) +
  ggplot2::scale_y_continuous(expand = ggplot2::expansion(mult = c(0, 0), add = c(6, 0))) +
  ggplot2::scale_color_gradient2(low = "blue", mid = "white", high = "red", name = "NES") +
  ggplot2::coord_cartesian()
p
```

```{r}
printGSEATable(gsea)
```

```{r fig.height=12, fig.width=14}
enrichplot::ridgeplot(gsea, showCategory = 100, label_format = 100) +
  ggplot2::ggtitle("Correlation distributions of core enriched genes") +
  ggplot2::xlab("Pearson's R")
```

### 邻域负荷

#### 方差解释

查看源代码可知，`NAM_nbhdXpc` 等于 `NAM_embeddings`，而 `NAM_sampleXpc` 等于 `NAM_loadings` 。

```{r}
data.frame(sig2 = cnaRes$NAM_svs / sum(cnaRes$NAM_svs)) |>
  tibble::rowid_to_column('PC') |>
  ggplot2::ggplot(ggplot2::aes(PC, sig2)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::theme_classic(base_size = 14) +
  ggplot2::labs(y = 'Fraction of variance explained', x = 'NAM PC') +
  ggplot2::scale_x_continuous(breaks = 1:20)
```

#### 细胞投影

```{r}
nampcs <- paste0("NAM", colnames(cnaRes$NAM_embeddings))
colData(sce_tr4)[, nampcs] <- as.data.frame(cnaRes$NAM_embeddings)
```

```{r fig.height=8, fig.width=16}
lapply(nampcs, function(pcname) {
  scater::plotReducedDim(
    sce_tr4,
    dimred = "UMAP",
    point_size = 0.5,
    colour_by = pcname,
    text_by = "cellType",
    text_size = 3,
    show_violin = TRUE) +
  ggplot2::scale_color_gradientn(colours = blue_to_orange) +
  ggplot2::ggtitle(pcname) +
  ggplot2::coord_fixed()
}) |> patchwork::wrap_plots(nrow = 2)
```

### 样本负荷

#### 样本间的距离

```{r}
rownames(cnaRes$NAM_loadings) <- as.character(cnaObj$samplem[[cnaObj$samplem_key]])

df_to_plot <- data.frame(
  PC1 = cnaRes$NAM_loadings[, 1],
  PC2 = cnaRes$NAM_loadings[, 2],
  PC3 = cnaRes$NAM_loadings[, 3],
  Time = cnaObj$samplem$time
)

plotly::plot_ly(
  df_to_plot,
  x = ~ PC1, y = ~ PC2, z = ~ PC3,
  color = ~ Time,
  #colors = cellTypeColors,
  type = "scatter3d",
  mode = "markers",
  marker = list(size = 4)
)
```

#### 细胞类型丰度关联

> Correlation across samples between the sample loadings for each NAM-PC and the abundances of each of the five main cell types in the dataset.

```{r}
cluster_prop <- t(prop.table(table(sce_tr4$cluster_id, sce_tr4$sample_id), 2))
cluster_prop
```

```{r}
corres <- psych::corr.test(
  x = cnaRes$NAM_loadings[rownames(cluster_prop),],
  y = cluster_prop, method = "pearson"
)
ComplexHeatmap::Heatmap(
    corres$r,
    name = "Pearson \nCorrelation",
    col = c("blue", "white", "red"),
    row_names_side = "right",
    row_dend_side = "left",
    column_names_side = "bottom",
    column_dend_side = "top",
    column_names_rot = 45,
    cell_fun = function(j, i, x, y, width, height, fill) {
      if (corres$p[i, j] < 0.05)
        grid::grid.text(
          sprintf("%.2f", corres$r[i, j]),
          x, y, gp = grid::gpar(fontsize = 10)
        )
    }
  )
```

