---
title: "Decompose the composition of coleoptile niches"
author: "Altair Wei"
date: "2022/2/23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

## 准备 LCM-Seq 数据

```{r}
lcm_counts <- readr::read_tsv("../results/LCMSeq/gene_quanti/counts/all.featureCounts", comment = "#")
lcm_counts <- lcm_counts[!(names(lcm_counts) %in% c("Chr", "Start", "End", "Strand", "Length"))]
names(lcm_counts) <- c("Gene", basename(dirname(names(lcm_counts)[-1])))
lcm_mtx <- as.matrix(lcm_counts[-1])
rownames(lcm_mtx) <- lcm_counts$Gene
lcm_mtx <- lcm_mtx[ rowSums(lcm_mtx) > 1, ]
```

## 准备细胞类型特征矩阵

```{r}
hfile <- SeuratDisk::Connect("../output/obj_annotated.h5seurat")
obj <- SeuratDisk::LoadH5Seurat(
  "../output/obj_annotated.h5seurat",
  assays = list(RNA="counts"),
  reductions = FALSE
)

sc_counts <- Seurat::GetAssayData(obj, slot = "counts")
```

### 标志基因

这个标志基因是通过 `Seurat::FindAllMarkers` 找出的：

```{r}
markers_df <- readRDS(Sys.glob("../output/int_cluster_markers_*.rds"))
markergenes <- unique(markers_df$gene)
```

但这里我没有找到 (Baccin, Chiara et al. 2020) 文献中提到的寻找特征基因的算法。

### 特征矩阵

对 counts 求平均值，以

```{r}
sc_mtx <- vapply(
  X = levels(Seurat::Idents(obj)),
  FUN = function(i) {
    Matrix::rowMeans(
      sc_counts[markergenes, Seurat::WhichCells(obj, ident = i)])
  },
  FUN.VALUE = numeric(length = length(markergenes))
)

sc_mtx <- sc_mtx[ rowSums(sc_mtx) > 1, ]
```


## 反卷积

### 运行算法

```{r}
LCM_design <- c(
  "VC_1" = "vascular",
  "VC_2" = "vascular",
  "VC_3" = "vascular",
  "CC_1" = "cortex",
  "CC_2" = "cortex",
  "CC_3" = "cortex",
  "MC_1" = "mesophyll",
  "MC_2" = "mesophyll",
  "MC_3" = "mesophyll"
)

CIBER <- RNAMagnet::runCIBERSORT(
  exprs = lcm_mtx,
  base = sc_mtx,
  design = LCM_design,
  markergenes = intersect(rownames(lcm_mtx), rownames(sc_mtx)),
  mc.cores = 3
)
```

### 绘图

值得注意的是 CIBERSORT 适合评估 niches 之间的细胞类型比例差异，但不适合评估单个 niches 中的细胞比例。

> The algorithm excels at comparing relative cell type abundancies between niches (i.e. ‘cell type X localizes to niche A over niche B and niche C’), but performs only moderately at estimating cell type proportions within a single niche (i.e. it cannot draw statements like ‘niche A consists to 70% of cell type X and 30% of cell type Y’)

```{r fig.height=7, fig.width=8, dev='svglite'}
ggplot2::ggplot(
    data = CIBER,
    mapping = ggplot2::aes(x = SampleClass, y= Fraction, color = CellType)
  ) +
  ggplot2::geom_point(stat = "summary", fun = mean) +
  ggplot2::geom_errorbar(
    stat="summary",
    fun.min = function(x) mean(x)+sd(x)/sqrt(length(x)),
    fun.max = function(x) mean(x)-sd(x)/sqrt(length(x)),
    width = 0.2
  ) +
  ggplot2::scale_y_continuous(labels = scales::percent) +
  ggplot2::facet_wrap(~ CellType, scales = "free_y") +
  ggplot2::theme_bw(base_size=12) +
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(angle = 90),
    panel.grid = ggplot2::element_blank()
  ) +
  ggplot2::guides(color = "none") +
  ggplot2::ylab("CIBERSORT estimate (a.u.)") +
  ggplot2::xlab("Niche")
```

```{r}
mtx_df <- CIBER %>%
  dplyr::group_by(SampleClass, CellType) %>%
  dplyr::summarise(
    MeanFrac = mean(Fraction)
  ) %>%
  dplyr::select(CellType, SampleClass, MeanFrac) %>%
  tidyr::pivot_wider(names_from = SampleClass, values_from = MeanFrac)

mtx <- as.matrix(mtx_df[, -1])
rownames(mtx) <- mtx_df$CellType
mtx
```

```{r fig.height=5, fig.width=4}
ComplexHeatmap::Heatmap(
  matrix = mtx,
  col = c("white", "blue", "red"),
  row_dend_width = grid::unit(4, "cm"),
  heatmap_legend_param = list(
    title = "Fraction",
    #title_position = "leftcenter",
    legend_direction = "vertical",
    legend_height = grid::unit(4, "cm")
  )
)
```


