---
title: "Differential Abundance Testing by miloR"
author: "Altair Wei"
date: '2022-07-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(SingleCellExperiment)
```

## 算法探索

一些提示：

-   miloR 考虑到了样本间细胞数量的差异，也可以自己检查这个数量差异是否会误导差异丰度分析：<https://github.com/MarioniLab/miloR/issues/208>
-   关于多元变量的比较：<https://github.com/MarioniLab/miloR/issues/158>

### 数据准备

```{r read-seurat-obj}
obj <- readRDS(Sys.glob(
  "../results/ObjectCache/IntegrationAndAnnotation/obj_annotated_*.rds"))
```
 
```{r}
obj_1dpi <- subset(obj, subset = group %in% c("3DPI-MOCK", "3DPI-PNR2"))
sce <- Seurat::as.SingleCellExperiment(obj_1dpi)
sce$sample <- make.names(sce$sample)
sce$group <- make.names(sce$group)
```

```{r}
scater::plotReducedDim(sce, dimred = "TSNE", colour_by = "group", point_size = .2) +
  ggplot2::guides(color = ggplot2::guide_legend(override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

### 创建对象

```{r}
milo <- miloR::Milo(sce)
milo
```

### 构建 KNN 网络

关于 `k` 参数：This affects the power of DA testing, since we need to have enough cells from each sample represented in a neighbourhood to estimate the variance between replicates. On the other side, increasing k too much might lead to over-smoothing. We suggest to start by using the same value for k used for KNN graph building for clustering and UMAP visualization. We will later use some heuristics to evaluate whether the value of k should be increased.

```{r}
milo <- miloR::buildGraph(milo, k = 10, d = 30, reduced.dim = "HARMONY")
```

### 定义代表性领域

关于 `prop` 参数：The proportion of cells to randomly sample to start with. We suggest using `prop=0.1` for datasets of less than 30k cells. For bigger datasets using `prop=0.05` should be sufficient (and makes computation faster).

```{r}
milo <- miloR::makeNhoods(
  milo, prop = 0.1, k = 10, d = 30,
  refined = TRUE, reduced_dims = "HARMONY")
```

我们需要评估每个领域包含的细胞数量。根据经验，我们希望平均邻域大小超过 5 x N_samples

```{r}
miloR::plotNhoodSizeHist(milo)
```

我们的数据集有 20 个样本，所以平均领域尺寸应该大于 5×20=100

```{r}
summary(Matrix::colSums(miloR::nhoods(milo)))
```

### 计算领域中各样本细胞数量

```{r}
milo <- miloR::countCells(milo, meta.data = as.data.frame(colData(milo)), sample = "sample")
head(miloR::nhoodCounts(milo))
```

### 定义实验设计

```{r}
design <- data.frame(colData(milo))[, c("sample", "treatment")]
design <- dplyr::distinct(design)
rownames(design) <- design$sample
design
```

```{r}
## Reorder rownames to match columns of nhoodCounts(milo)
design <- design[colnames(miloR::nhoodCounts(milo)), , drop=FALSE]
design
```

### 定义领域连接性

> Milo uses an adaptation of the Spatial FDR correction introduced by `cydar`, which accounts for the overlap between neighbourhoods. Specifically, each hypothesis test P-value is weighted by the reciprocal of the kth nearest neighbour distance. To use this statistic we first need to store the distances between nearest neighbors in the Milo object.

```{r}
milo <- miloR::calcNhoodDistance(milo, d = 20, reduced.dim = "HARMONY")
```

### 差异丰度测试

```{r}
da_results <- miloR::testNhoods(
  milo, design = ~ treatment,
  design.df = design, reduced.dim = "HARMONY", fdr.weighting = "graph-overlap")
```

几个主要的统计量：

-   `logFC`: indicates the log-Fold change in cell numbers between samples from stimulus and samples from control

-   `PValue`: reports P-values before FDR correction

-   `SpatialFDR`: reports P-values corrected for multiple testing accounting for overlap between neighbourhoods


```{r}
da_results %>%
  dplyr::arrange(SpatialFDR) %>%
  head()
```

### 可视化结果

```{r}
milo <- miloR::buildNhoodGraph(milo)
```

```{r fig.height=6, fig.width=14}
p1 <- scater::plotReducedDim(sce, dimred = "TSNE", colour_by = "group", point_size = .2) +
  ggplot2::guides(color = ggplot2::guide_legend(override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
p2 <- miloR::plotNhoodGraphDA(milo, da_results, layout = "TSNE", alpha = 0.1) +
  ggplot2::coord_fixed()
p1 + p2
```

```{r}
ggplot2::ggplot(da_results, ggplot2::aes(logFC, -log10(SpatialFDR))) + 
  ggplot2::geom_point() +
  ggplot2::geom_hline(yintercept = 1)
```


