---
title: "Known Pathways in DEGs of Each Cel Types"
author: "Altair Wei"
date: "2023-03-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Data

### Metadata

```{r}
cell_types <- c(
  "Stomata",
  "Epidermis",
  "Chlorenchyma",
  "Parenchyma",
  "Outer sheath",
  "Inner sheath",
  "Phloem",
  "Procambium"
)
```

### Gene Lists

```{r}
filelist <- Sys.glob("../data/pathways/*.txt")
names(filelist) <- basename(filelist) |> stringr::str_remove(stringr::fixed(".txt"))

pathways <- lapply(filelist, function(filename) {
  genes <- stringr::str_trim(readLines(filename))
  genes[genes != ""]
})
```

```{r}
summary(pathways)[, 1, drop = FALSE]
```


### DEGs of Cell Types

```{r}
ds_res <- readRDS(Sys.glob(
  "../results/ObjectCache/DifferentialState/ds_res_*.rds"))
```

```{r}
logfc_df <- ds_res$DESeq2 |>
  rhapsodykit::diff_state_significant(fdr_limit = 0.05, logfc_limit = 0) |>
  rhapsodykit::diff_state_format() |>
  dplyr::filter(!stringr::str_detect(contrast, "X\\dDPI\\.PNR2-X\\dDPI\\.TR4")) |>
  dplyr::select(gene, cluster_id, contrast, logFC) |>
  dplyr::mutate(
    time = stringr::str_extract(contrast, "\\dDPI"),
    contrast = stringr::str_replace(contrast,
      "X(\\dDPI)\\.(.*)-X(\\dDPI)\\.(.*)", "\\2-\\4")) |>
  tidyr::pivot_wider(names_from = "time", values_from = "logFC", values_fill = NA) |>
  tidyr::pivot_longer(cols = dplyr::ends_with("DPI"), names_to = "time", values_to = "logFC")
```

```{r}
logfc_mtx <- logfc_df |>
  tidyr::pivot_wider(names_from = c(cluster_id, contrast, time),
                     names_sep = "#",
                     values_from = logFC) |>
  tibble::column_to_rownames("gene") |>
  as.matrix()

str_coldata <- strsplit(colnames(logfc_mtx), "#")

mtx_coldata <- data.frame(
  colidx = seq_along(colnames(logfc_mtx)),
  cell_types = factor(sapply(str_coldata, function(x) x[[1]]),
                      levels = cell_types),
  contrast = factor(sapply(str_coldata, function(x) x[[2]]),
                    #levels = c("PNR2-MOCK", "TR4-MOCK", "PNR2-TR4")),
                    levels = c("PNR2-MOCK", "TR4-MOCK")),
  times = factor(sapply(str_coldata, function(x) x[[3]]),
                 levels = c("1DPI", "2DPI", "3DPI"))
)

mtx_coldata <- mtx_coldata[with(mtx_coldata, order(cell_types, contrast, times)),]

logfc_mtx <- logfc_mtx[, mtx_coldata$colidx]
```

```{r}
ha = ComplexHeatmap::HeatmapAnnotation(
      which = "column",
      df = mtx_coldata[-1],
      col = list(
        cell_types = structure(
          ggthemes::tableau_color_pal()(
            length(levels(mtx_coldata$cell_types))),
          names = levels(mtx_coldata$cell_types)
        ),
        contrast = structure(
          c("#E41A1C", "#377EB8"),
          names = levels(mtx_coldata$contrast)
        ),
        times = structure(
          RColorBrewer::brewer.pal(3, "YlGn"),
          names = levels(mtx_coldata$times)
        )
      ))
```

### Calculate GSVA

```{r}
pseudobulk_cpm_mean <- readRDS(Sys.glob(
  "../results/ObjectCache/Pseudobulk/pseudobulk_cpm_mean_*.rds"))
```

```{r}
expr_gsva <- GSVA::gsva(
  pseudobulk_cpm_mean, pathways,
  method = "gsva", kcdf = "Gaussian",
  verbose = FALSE)

gsva_df <- as.data.frame(expr_gsva) |>
    tibble::rownames_to_column("Pathway") |>
    tidyr::pivot_longer(-Pathway, names_to = "bulk", values_to = "ES") |>
    tidyr::separate(bulk, sep = "\\.", into = c("treatment", "time", "cellType"))
```

```{r}
gsva_df$cellType <- factor(gsva_df$cellType, levels = cell_types)
```

### Functions

```{r}
plotRadarGSVA <- function(gsva, pathw) {
  dplyr::filter(gsva,
      Pathway == pathw) |>
    ggplot2::ggplot(
      ggplot2::aes(
          x = cellType,
          y = ES,
          group = treatment,
          color = treatment,
          fill = treatment
        )) +
    ggplot2::geom_polygon(linewidth = 1, alpha = .1) +
    ggplot2::geom_point() +
    see::coord_radar() +
    ggplot2::facet_wrap(~ time, nrow = 1) +
    see::theme_radar() +
    NULL
}
```

```{r}
plotDEGsHeatmap <- function(mtx, genes, colAnno, max = 5, ...) {
  detected <- intersect(rownames(mtx), genes)
  mtx <- mtx[detected,]

  # undetected <- setdiff(genes, detected)
  # mtx <- rbind(mtx, matrix(
  #   NA, nrow = length(undetected), ncol = ncol(mtx), dimnames = list(undetected)))
  # mtx <- mtx[genes,]
  if (length(detected) != length(genes))
    warning(sprintf(
      "Only %i out of %i genes were detected in DEGs\n",
      length(detected), length(genes)))

  outliers_idx <- which(abs(mtx) > max, arr.ind = TRUE)
  if (length(outliers_idx) != 0)
    warning(sprintf(paste0("%i gene%s with abs(logFC) greater than %i were shown",
                    " in maximum color and marked with white asterisk.\n"),
                    nrow(outliers_idx),
                    if (nrow(outliers_idx) > 1) "s" else "",
                    max))

  dend_mtx <- mtx
  dend_mtx[is.na(dend_mtx)] <- 0
  dend_row_hc <- dendsort::dendsort(
    fastcluster::hclust(dist(dend_mtx)),
    isReverse = TRUE)

  p <- ComplexHeatmap::Heatmap(
    matrix = mtx,
    name = "Log2 Fold Change",
    col = circlize::colorRamp2(c(-max, 0, max), c("blue", "white", "red")),
    use_raster = TRUE,
    na_col = "grey",
    cluster_rows = dend_row_hc,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    show_row_names = TRUE,
    top_annotation = colAnno,
    heatmap_legend_param = list(
      title_position = "leftcenter-rot",
      legend_height  = grid::unit(4, "cm")
    ),
    cell_fun = function(j, i, x, y, width, height, fill) {
      val <- abs(mtx[i, j])
      if (!is.na(val) && val > max)
        grid::grid.text("★", x, y, gp = grid::gpar(
          fontsize = 8, col = "white", fontface = "bold"))
    },
    ...)

  ComplexHeatmap::draw(p, merge_legend = TRUE)
}
```

## Analysis Based on Known Pathways

TODO: perform ORA for each pathway.

### Phytosynthesis related {.tabset}

#### PhotosynAnt

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "PhotosynAnt")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["PhotosynAnt"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### PhotosynCfix

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "PhotosynCfix")
```

```{r fig.height=14, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["PhotosynCfix"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### PhotosynPSEA

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "PhotosynPSEA")
```

```{r fig.height=14, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["PhotosynPSEA"]],
  ha, column_split = mtx_coldata$cell_types)
```

### Cell wall related {.tabset}

#### Aromatic amino acid

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "Aromatic amino acid")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["Aromatic amino acid"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### General phenylpropanoid

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "General phenylpropanoid")
```

```{r fig.height=8, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["General phenylpropanoid"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### Lignin biosynthesis

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "Lignin biosynthesis")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["Lignin biosynthesis"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### Flavanones

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "Flavanones")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["Flavanones"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### HCAAs

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "HCAAs")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["HCAAs"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### Suberin

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "Suberin")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["Suberin"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### CalloseSyn

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "CalloseSyn")
```

```{r fig.height=5, fig.width=14}
detected <- intersect(rownames(logfc_mtx), pathways[["CalloseSyn"]])
print(detected)
logfc_mtx[detected,][!is.na(logfc_mtx[detected,])]
```

#### Polyamine

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "Polyamine")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["Polyamine"]],
  ha, column_split = mtx_coldata$cell_types)
```

### Immune related {.tabset}

#### CCNBLRR

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "CCNBLRR")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["CCNBLRR"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### NBLRR

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "NBLRR")
```

```{r fig.height=10, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["NBLRR"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### TIR

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "TIR")
```

```{r fig.height=5, fig.width=14}
detected <- intersect(rownames(logfc_mtx), pathways[["TIR"]])
print(detected)
logfc_mtx[detected,][!is.na(logfc_mtx[detected,])]
```

#### PR

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "PR")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["PR"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### RLK

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "RLK")
```

```{r fig.height=16, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["RLK"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### TaNPR

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "TaNPR")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["TaNPR"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### RxNonly

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "RxNonly")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["RxNonly"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### GOSAR

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "GOSAR")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["GOSAR"]],
  ha, column_split = mtx_coldata$cell_types)
```

### Metabolism related {.tabset}

#### gluconeogenesis

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "gluconeogenesis")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["gluconeogenesis"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### glycolysis

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "glycolysis")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["glycolysis"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### OxiPhos

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "OxiPhos")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["OxiPhos"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### Sweet

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "Sweet")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["Sweet"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### TaSUT

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "TaSUT")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["TaSUT"]],
  ha, column_split = mtx_coldata$cell_types)
```

### Signaling {.tabset}

#### GOCaSignaling

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "GOCaSignaling")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["GOCaSignaling"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### MAPK

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "MAPK")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["MAPK"]],
  ha, column_split = mtx_coldata$cell_types)
```

### Transcription Factors {.tabset}

#### Myb

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "Myb")
```

```{r fig.height=18, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["Myb"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### NAC

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "NAC")
```

```{r fig.height=14, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["NAC"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### WRKY

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "WRKY")
```

```{r fig.height=14, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["WRKY"]],
  ha, column_split = mtx_coldata$cell_types)
```

## Analysis of Other RNA-seq results

### fusaoctaxin A {.tabset}

#### FA3FB3up

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "FA3FB3up")
```

```{r fig.height=14, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["FA3FB3up"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### FA3hUp

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "FA3hUp")
```

```{r fig.height=26, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["FA3hUp"]],
  ha, column_split = mtx_coldata$cell_types)
```

#### FAFB-3hUp

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "FAFB-3hUp")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["FAFB-3hUp"]],
  ha, column_split = mtx_coldata$cell_types)
```

### TIRH2Y

```{r fig.height=5, fig.width=14}
plotRadarGSVA(gsva_df, "TIRH2Y")
```

```{r fig.height=5, fig.width=14}
plotDEGsHeatmap(
  logfc_mtx, pathways[["TIRH2Y"]],
  ha, column_split = mtx_coldata$cell_types)
```
