---
title: "Analysis of Intersecting DEG Sets"
author: "Altair Wei"
date: "2023-04-25"
output: html_document
---

```{r setup, include=FALSE}
library(magrittr)
library(patchwork)
source("../scripts/LoadUtils.R", chdir = TRUE)
```

## Load Data

### Cell Annotation

```{r}
cell_types <- c(
  "Stomata",
  "Epidermis",
  "Chlorenchyma",
  "Parenchyma",
  "Outer sheath",
  "Inner sheath",
  "Phloem",
  "Procambium"
)
```

### DS Results

```{r}
ds_res <- readRDS(Sys.glob("../results/ObjectCache/DifferentialState/ds_res_*.rds"))
```

```{r}
ds_sig <- ds_res$DESeq2 %>%
  rhapsodykit::diff_state_significant(fdr_limit = 0.05, logfc_limit = 0)
```

```{r}
df_degs <- rhapsodykit::diff_state_format(ds_sig)

matched <- stringr::str_match(
  df_degs$contrast, "X(\\dDPI)\\.(MOCK|PNR2|TR4)-X(\\dDPI)\\.(MOCK|PNR2|TR4)")

df_degs$time <- factor(matched[, 2], levels = c("0DPI", "1DPI", "2DPI", "3DPI"))
df_degs$contrast <- paste(matched[, 3], matched[, 5], sep = " vs. ")
df_degs$regu_type <- factor(ifelse(df_degs$logFC > 0, "UP", "DOWN"), levels = c("UP", "DOWN"))

df_degs <- df_degs |>
  dplyr::select(time, contrast, cluster_id, regu_type, gene) |>
  dplyr::filter(contrast != "PNR2 vs. TR4")
```

### Functions

```{r}
upset_contrasts_list <- function(input, ...) {
  sets_list <- input$table %>%
    purrr::map(function(x) {
      purrr::map(x, "gene") %>%
        purrr::compact()
    })

  sets_list <- Filter(\(x) length(x) > 0, sets_list)

  sets_metadata <- sets_list %>%
    purrr::imap(function(contrast, key) {
      cbind(
        sets = sprintf(
          "%s [%s]",
          stringr::str_replace(key,
            "X(\\dDPI)\\.(.*)-X.*", "\\1-\\2"),
          names(contrast)),
        contrast = key,
        cluster = names(contrast)
      )
    }) %>%
    do.call(rbind, .) %>%
    as.data.frame()

  sets_to_plot <- sets_list %>%
    unlist(recursive = FALSE)
  
  names(sets_to_plot) <- sets_metadata$sets

  sets_to_plot
}
```

```{r}
plot_updown_contrast <- function(ds, contrast, clusters = NULL) {
  p1 <- ds %>%
    rhapsodykit::diff_state_subset(
      contrast = contrast,
      clusters = clusters) %>%
    upset_contrasts_list() %>%
    rhapsodykit::from_list_to_upset() %>%
    rhapsodykit::upset_dataframe() %>%
    rhapsodykit::upset_plot(
      order_by = "degree",
      n_intersections = 100)

  up_list <- ds %>%
    rhapsodykit::diff_state_subset(
      contrast = contrast,
      clusters = clusters) %>%
    rhapsodykit::diff_state_filter(logFC > 0) %>%
    upset_contrasts_list()

  down_list <- ds %>%
    rhapsodykit::diff_state_subset(
      contrast = contrast,
      clusters = clusters) %>%
    rhapsodykit::diff_state_filter(logFC < 0) %>%
    upset_contrasts_list()

  p2 <- rhapsodykit::upset_updown_regulated(
    up_list, down_list,
    order_by = "degree", n_intersections = 100)

  p1 / p2
}
```

因为有不同的 contrast ，所以无法显示上调与下调的基因？也对，某个基因可能在一个集合（cellType或者contrast）里面上调，但在另一个集合里面下调，这种基因会被漏掉。这个困局对任何在 UpsetPlot 上显示上下调基因的绘图来说都是存在的。

## Visualization of Intersections

### UpSet app to explore

TODO: 利用 BioMart 获取基因的属性，包括所属染色体，Description，但不要包括 GO 注释等。

Link: [UpSet App for All DEG Sets](upset-app/index.html)

```{r}
df_upset <- df_degs |>
  dplyr::group_by(time, contrast, cluster_id) |>
  dplyr::summarise(
    genes = list(gene),
    .groups = "drop") |>
  dplyr::mutate(contrast = stringr::str_remove(contrast, " vs\\. .*"))
```

#### All DEG sets

```{r}
writeUpSetData <- function(df, name) {
  data <- df |>
    tibble::deframe() |>
    rhapsodykit::from_list_to_upset() |>
    tibble::rownames_to_column("Gene")

  readr::write_csv(
    x = data,
    file = sprintf(
      "../docs/upset-app/data/wheatscrnaseq/%s.csv", name))

  meta <- jsonlite::toJSON(
    x = list(
      file = sprintf("data/wheatscrnaseq/%s.csv", name),
      name = name,
      header =  0L,
      separator = ",",
      skip = 0L,
      meta = list(
        list(type = "id", index = 0L, name = "Gene")
      ),
      sets = list(
        list(format = "binary", start = 1L, end = nrow(df))
      )
    ),
    auto_unbox = TRUE, pretty = TRUE)

  cat(meta, file = sprintf(
    "../docs/upset-app/data/wheatscrnaseq/%s.json", name))
}
```

```{r}
df_upset |>
  dplyr::transmute(
    sets = paste(time, contrast, cluster_id, sep = "-"),
    genes = genes) |>
  writeUpSetData("DEGs all")
```

#### DEG sets by time

```{r}
for (ti in paste0(1:3, "DPI"))
  df_upset |>
    dplyr::filter(time == ti) |>
    dplyr::transmute(
      sets = paste(contrast, cluster_id, sep = "-"),
      genes = genes) |>
    writeUpSetData(paste("DEGs", ti))
```

#### DEG sets by contrast

```{r}
for (co in c("PNR2", "TR4"))
  df_upset |>
    dplyr::filter(contrast == co) |>
    dplyr::transmute(
      sets = paste(time, cluster_id, sep = "-"),
      genes = genes) |>
    writeUpSetData(paste("DEGs", co))
```

#### DEG sets by cell type

```{r}
for (ct in cell_types)
  df_upset |>
    dplyr::filter(cluster_id == ct) |>
    dplyr::transmute(
      sets = paste(time, contrast, sep = "-"),
      genes = genes) |>
    writeUpSetData(paste("DEGs", ct))
```

#### Add datasets to UpSet

```{r}
jsonlite::toJSON(c(
      "data/wheatscrnaseq/DEGs all.json",
      sprintf("data/wheatscrnaseq/DEGs %s.json", paste0(1:3, "DPI")),
      sprintf("data/wheatscrnaseq/DEGs %s.json", c("PNR2", "TR4")),
      sprintf("data/wheatscrnaseq/DEGs %s.json", cell_types)
    ), pretty = TRUE) |>
  cat(file = "../docs/upset-app/datasets.json")
```

### Intersections between cell types

```{r}
plot_updown <- function(ds, contrast) {
  df <- ds %>%
    rhapsodykit::diff_state_pull(contrast, seq_along(ds$data), "gene") %>%
    rhapsodykit::from_list_to_upset() %>%
    rhapsodykit::upset_dataframe()

  sets_names <- df %>%
    dplyr::select(clusters) %>%
    tidyr::unnest(cols = clusters) %>%
    dplyr::count(clusters) %>%
    dplyr::mutate(n = sprintf("%s (%s)", clusters, n)) %>%
    tibble::deframe()

  p1 <- df |>
    tidyr::unnest(cols = clusters) |>
    dplyr::mutate(clusters = dplyr::recode(clusters, !!!sets_names)) |>
    dplyr::group_by(gene) |>
    dplyr::summarise(clusters = list(clusters)) %>%
    rhapsodykit::upset_plot(n_intersections = 30)
  
  up_list <- ds %>%
    rhapsodykit::diff_state_filter(logFC > 0) %>%
    rhapsodykit::diff_state_pull(contrast, seq_along(ds$data), "gene")

  down_list <- ds %>%
    rhapsodykit::diff_state_filter(logFC < 0) %>%
    rhapsodykit::diff_state_pull(contrast, seq_along(ds$data), "gene")

  p2 <- rhapsodykit::upset_updown_regulated(
    up_list, down_list, n_intersections = 30)

  p1 / p2
}
```

#### All possible intersections {.tabset}

```{r results='asis', fig.height=10, fig.width=10}
for (dp in c("1DPI", "2DPI", "3DPI")) {
  cat("#####", dp, "{.tabset}\n\n")
  for (tr in list(c("PNR2", "MOCK"),
                  c("TR4", "MOCK"),
                  c("PNR2", "TR4"))) {

    cont = sprintf("X%s.%s-X%s.%s", dp, tr[[1]], dp, tr[[2]])
    title = sprintf("%s-%s vs. %s-%s", dp, tr[[1]], dp, tr[[2]])
    cat("######", title, "\n\n")

    p <- plot_updown(ds_sig, cont) &
      ggplot2::ggtitle(title)
    print(p)

    cat("\n\n")
  }
  cat("\n\n")
}
```

#### Uniqueness of cell type responses

Calculate cell type-only DEGs for each `time` and `contrast` combination.

```{r}
cluster_only <- df_degs |>
  dplyr::group_by(time, contrast) |>
  tidyr::nest() |>
  dplyr::mutate(only = lapply(data, function(x) {
    # Calculate unique DEGs for each cell type
    x |>
      dplyr::group_by(cluster_id) |>
      dplyr::summarise(genes = list(gene)) |>
      dplyr::rowwise() |>
      dplyr::mutate(total = length(genes)) |>
      dplyr::ungroup() |>
      dplyr::mutate(diff = vapply(
        X = seq_along(genes),
        FUN = function(idx) length(setdiff(
          genes[[idx]], unique(unlist(genes[-idx])))),
        FUN.VALUE = integer(1))) |>
      dplyr::select(cluster_id, diff, total)
  })) |>
  dplyr::ungroup() |>
  dplyr::select(time, contrast, only) |>
  tidyr::unnest(cols = only) |>
  dplyr::mutate(pct = 100 * (diff / total))

cluster_only |>
  dplyr::mutate(pct = format(pct, digits = 2)) |>
  download_table("cluster_only")
```

```{r fig.height=7, fig.width=10}
p <- ggplot2::ggplot(
    data = cluster_only,
    mapping = ggplot2::aes(
      x = diff, y = pct, color = cluster_id)) +
  ggplot2::geom_point() +
  ggthemes::scale_color_tableau() +
  ggrepel::geom_text_repel(
    data = dplyr::filter(cluster_only, diff > 40, pct > 50),
    mapping = ggplot2::aes(
      label = sprintf("%s;%s%%", diff, format(pct, digits = 1))),
    size = 3, color = "#4d4d4d",
    show.legend = FALSE) +
  ggplot2::facet_grid(time ~ contrast) +
  ggplot2::theme_bw()

p
```

#### Similarity of cell type responses {.tabset}

```{r}
df_jacmtx <- df_degs |>
  dplyr::mutate(cluster_id = factor(
    cluster_id, levels = cell_types)) |>
  dplyr::group_by(time, contrast) |>
  tidyr::nest() |>
  dplyr::mutate(jacmtx = lapply(data, function(x) {
    dplyr::group_by(x, cluster_id) |>
    dplyr::summarise(genes = list(gene)) |>
    tidyr::complete(cluster_id) |>
    dplyr::ungroup() |>
    dplyr::mutate(data = lapply(
      X = seq_along(genes),
      FUN = function(idx) {
        data.frame(
          cluster_ta = cluster_id,
          jaccard = vapply(
            X = genes,
            FUN = function(gs)
              bayesbio::jaccardSets(genes[[idx]], gs),
            FUN.VALUE = numeric(1))
        )
      })) |>
    dplyr::select(cluster_id, data) |>
    tidyr::unnest(cols = data) |>
    tidyr::pivot_wider(names_from = cluster_ta, values_from = jaccard) |>
    tibble::column_to_rownames("cluster_id") |>
    as.matrix()
  })) |>
  dplyr::ungroup() |>
  dplyr::select(time, contrast, jacmtx)
```

```{r results='asis'}
df_jacmtx |>
  purrr::pwalk(function(time, contrast, jacmtx) {
    title = paste(time, contrast, sep = " - ")
    cat("#####", title, "\n\n")
    dendmtx <- jacmtx
    dendmtx[is.na(dendmtx)] <- 0
    rowdend <- hclust(as.dist(1 - dendmtx))
    coldend <- hclust(as.dist(1 - t(dendmtx)))

    ComplexHeatmap::Heatmap(
      matrix = jacmtx,
      name = "Jaccard's coef",
      column_title = title,
      col = c("white", "red"),
      na_col = "white",
      cluster_rows = rowdend,
      cluster_columns = coldend,
      cell_fun = function(j, i, x, y, width, height, fill) {
        value <- jacmtx[i, j]
        if (!is.na(value))
          grid::grid.text(
            sprintf("%.0f%%", value * 100),
            x, y, gp = grid::gpar(fontsize = 10)
          )
      }) |> print()
    cat("\n\n")
  })
```

### Comparison of all DEG list

#### All possible intersections

```{r fig.height=20, fig.width=20}
plot_updown_contrast(
  ds_sig,
  contrast = c(
    "X1DPI.PNR2-X1DPI.MOCK",
    "X1DPI.TR4-X1DPI.MOCK",
    "X2DPI.PNR2-X2DPI.MOCK",
    "X2DPI.TR4-X2DPI.MOCK",
    "X3DPI.PNR2-X3DPI.MOCK",
    "X3DPI.TR4-X3DPI.MOCK")
)
```

#### Similarity of DEGs list

```{r fig.height=6, fig.width=8}
df <- df_degs |>
  dplyr::select(-regu_type) |>
  dplyr::mutate(
    time = forcats::fct_drop(time),
    cluster_id = factor(cluster_id, levels = cell_types),
    contrast = factor(contrast, levels = c("PNR2 vs. MOCK", "TR4 vs. MOCK"))) |>
  dplyr::group_by(time, contrast, cluster_id) |>
  dplyr::summarise(
    genes = list(gene),
    .groups = "drop") |>
  dplyr::mutate(setid = paste(time, contrast, cluster_id, sep = " - "))

df_anno <- dplyr::select(df, cluster_id, contrast, time)

jacmtx <- dplyr::select(df, setid, genes) |>
  dplyr::mutate(data = lapply(
    X = seq_along(genes),
    FUN = function(idx) {
      data.frame(
        setid_ta = setid,
        jaccard = vapply(
          X = genes,
          FUN = function(gs)
            bayesbio::jaccardSets(genes[[idx]], gs),
          FUN.VALUE = numeric(1))
      )
    })) |>
  dplyr::select(setid, data) |>
  tidyr::unnest(cols = data) |>
  tidyr::pivot_wider(names_from = setid_ta, values_from = jaccard) |>
  tibble::column_to_rownames("setid") |>
  as.matrix()

hac <- ComplexHeatmap::HeatmapAnnotation(
  which = "column",
  df = as.data.frame(df_anno),
  col = list(
    cluster_id = structure(
      ggthemes::tableau_color_pal()(
        length(levels(df_anno$cluster_id))),
      names = levels(df_anno$cluster_id)
    ),
    contrast = structure(
      c("#E41A1C", "#377EB8"),
      names = levels(df_anno$contrast)
    ),
    time = structure(
      RColorBrewer::brewer.pal(3, "YlGn"),
      names = levels(df_anno$time)
    )
  ))

har <- ComplexHeatmap::HeatmapAnnotation(
    which = "row",
    df = as.data.frame(df_anno),
    col = list(
      cluster_id = structure(
        ggthemes::tableau_color_pal()(
          length(levels(df_anno$cluster_id))),
        names = levels(df_anno$cluster_id)
      ),
      contrast = structure(
        c("#E41A1C", "#377EB8"),
        names = levels(df_anno$contrast)
      ),
      time = structure(
        RColorBrewer::brewer.pal(3, "YlGn"),
        names = levels(df_anno$time)
      )
    ),
    show_legend = FALSE,
    show_annotation_name = FALSE
  )

p <- ComplexHeatmap::Heatmap(
  matrix = jacmtx,
  name = "Jaccard's coef",
  col = circlize::colorRamp2(c(0, 0.5), c("white", "red")),
  na_col = "white",
  show_column_names = FALSE,
  show_row_names = FALSE,
  top_annotation = hac,
  left_annotation = har,
  # `jacmtx` is the similarity matrix, so distance matrix can
  # be calculated by `1 - jacmtx`
  clustering_distance_rows = \(x) as.dist(1 - x),
  clustering_distance_columns = \(x) as.dist(1 - x))

ComplexHeatmap::draw(p, merge_legend = TRUE)
```

### The differences between treatments

Split DEG lists by time.

#### Possible intersections {.tabset}

```{r fig.height=14, fig.width=18, results='asis'}
for (dp in c("1DPI", "2DPI", "3DPI")) {
  cat("#####", dp, "\n\n")
  contrast <- sprintf(c(
    "X%s.PNR2-X%s.MOCK",
    "X%s.TR4-X%s.MOCK"), dp, dp)
  p <- plot_updown_contrast(
    ds_sig, contrast = contrast)
  print(p)
  cat("\n\n")
}
```

#### Similarity of treatments {.tabset}

```{r results='hide'}
df_jacmtx <- df_degs |>
  dplyr::select(-regu_type) |>
  dplyr::mutate(
    time = forcats::fct_drop(time),
    cluster_id = factor(cluster_id, levels = cell_types),
    contrast = factor(contrast, levels = c("PNR2 vs. MOCK", "TR4 vs. MOCK"))) |>
  dplyr::group_by(time) |>
  tidyr::nest() |>
  dplyr::mutate(jacmtx = lapply(data, function(x) {
    dplyr::group_by(x, contrast, cluster_id) |>
      dplyr::summarise(genes = list(gene), .groups = "drop") |>
      dplyr::rename(contrast.1 = contrast, cluster_id.1 = cluster_id) |>
      dplyr::mutate(data = lapply(
        X = seq_along(genes),
        FUN = function(idx) {
          data.frame(
            contrast.2 = contrast.1,
            cluster_id.2 = cluster_id.1,
            jaccard = vapply(
              X = genes,
              FUN = function(gs)
                bayesbio::jaccardSets(genes[[idx]], gs),
              FUN.VALUE = numeric(1))
          )
        })) |>
      dplyr::select(contrast.1, cluster_id.1, data) |>
      tidyr::unnest(cols = data)
  })) |>
  dplyr::select(time, jacmtx)

df_jacmtx |>
  tidyr::unnest(cols = jacmtx) |>
  download_table("df_jacmtx")
```

```{r fig.height=8, fig.width=9, results='asis'}
df_jacmtx |>
  purrr::pwalk(function(time, jacmtx) {
    cat("#####", as.character(time), "\n\n")
    anno <- jacmtx |>
      dplyr::select(cluster_id.1, contrast.1) |>
      dplyr::rename(cluster_id = cluster_id.1, contrast = contrast.1) |>
      dplyr::distinct()

    mtx <- jacmtx |>
      dplyr::mutate(
        set.1 = paste(contrast.1, cluster_id.1, sep = " - "),
        set.2 = paste(contrast.2, cluster_id.2, sep = " - ")) |>
      dplyr::select(set.1, set.2, jaccard) |>
      tidyr::pivot_wider(names_from = set.2, values_from = jaccard) |>
      tibble::column_to_rownames("set.1") |>
      as.matrix()

    stopifnot(identical(
      colnames(mtx),
      paste(anno$contrast, anno$cluster_id, sep = " - ")))
    
    hac = ComplexHeatmap::HeatmapAnnotation(
        which = "column",
        df = as.data.frame(anno),
        col = list(
          cluster_id = structure(
            ggthemes::tableau_color_pal()(
              length(levels(anno$cluster_id))),
            names = levels(anno$cluster_id)
          ),
          contrast = structure(
            c("#E41A1C", "#377EB8"),
            names = levels(anno$contrast)
          )
        ))

    har <- ComplexHeatmap::HeatmapAnnotation(
        which = "row",
        df = as.data.frame(anno),
        col = list(
          cluster_id = structure(
            ggthemes::tableau_color_pal()(
              length(levels(anno$cluster_id))),
            names = levels(anno$cluster_id)
          ),
          contrast = structure(
            c("#E41A1C", "#377EB8"),
            names = levels(anno$contrast)
          )
        ),
        show_legend = FALSE,
        show_annotation_name = FALSE
      )

    p <- ComplexHeatmap::Heatmap(
      matrix = mtx,
      name = "Jaccard's coef",
      col = circlize::colorRamp2(c(0, 0.5), c("white", "red")),
      na_col = "white",
      column_title = as.character(time),
      show_column_names = FALSE,
      show_row_names = FALSE,
      clustering_distance_rows = \(x) as.dist(1 - x),
      clustering_distance_columns = \(x) as.dist(1 - x),
      top_annotation = hac,
      left_annotation = har,
      cell_fun = function(j, i, x, y, width, height, fill) {
        value <- mtx[i, j]
        if (!is.na(value))
          grid::grid.text(
            sprintf("%.0f%%", value * 100),
            x, y, gp = grid::gpar(fontsize = 10)
          )
      })
    
    ComplexHeatmap::draw(p, merge_legend = TRUE)
    cat("\n\n")
  })
```

### The concordance within cell type

Split DEG lists by cell type.

#### Possible intersections {.tabset}

```{r fig.height=14, fig.width=14, results='asis'}
for (cls in cell_types) {
  cat("#####", cls, "\n\n")
  p <- plot_updown_contrast(
    ds_sig,
    contrast = c(
      "X1DPI.PNR2-X1DPI.MOCK",
      "X1DPI.TR4-X1DPI.MOCK",
      "X2DPI.PNR2-X2DPI.MOCK",
      "X2DPI.TR4-X2DPI.MOCK",
      "X3DPI.PNR2-X3DPI.MOCK",
      "X3DPI.TR4-X3DPI.MOCK"),
    clusters = cls)
  print(p)
  cat("\n\n")
}
```

#### Euler diagrams of sets {.tabset}

Euler diagrams are Venn diagrams without the requirement that all set interactions be present (whether they are empty or not).

```{r results='asis'}
for (cls in cell_types) {
  cat("#####", cls, "\n\n")
  suppressMessages(
    ds_sig %>%
      rhapsodykit::diff_state_subset(
        contrast = c(
          "X1DPI.PNR2-X1DPI.MOCK",
          "X2DPI.PNR2-X2DPI.MOCK",
          "X3DPI.PNR2-X3DPI.MOCK",
          "X1DPI.TR4-X1DPI.MOCK",
          "X2DPI.TR4-X2DPI.MOCK",
          "X3DPI.TR4-X3DPI.MOCK"),
        clusters = cls) %>%
      upset_contrasts_list() %>%
      setNames(stringr::str_remove(names(.), " \\[.*\\]")) %>%
      eulerr::euler(shape = "ellipse") %>%
      plot(
        main = cls,
        lty = rep(1:2, each = 3),
        quantities = list(cex = 0.5),
        legend = list(side = "right")) %>%
      print()
  )

  cat("\n\n")
}
```

#### Similarity within cell type {.tabset}

```{r results='hide'}
df_jacmtx <- df_degs |>
  dplyr::select(-regu_type) |>
  dplyr::mutate(
    time = forcats::fct_drop(time),
    cluster_id = factor(cluster_id, levels = cell_types),
    contrast = factor(contrast, levels = c("PNR2 vs. MOCK", "TR4 vs. MOCK"))) |>
  dplyr::group_by(cluster_id) |>
  tidyr::nest() |>
  dplyr::mutate(jacmtx = lapply(data, function(x) {
    dplyr::group_by(x, time, contrast) |>
      dplyr::summarise(genes = list(gene), .groups = "drop") |>
      dplyr::rename(time.1 = time, contrast.1 = contrast) |>
      dplyr::mutate(data = lapply(
        X = seq_along(genes),
        FUN = function(idx) {
          data.frame(
            time.2 = time.1,
            contrast.2 = contrast.1,
            jaccard = vapply(
              X = genes,
              FUN = function(gs)
                bayesbio::jaccardSets(genes[[idx]], gs),
              FUN.VALUE = numeric(1))
          )
        })) |>
      dplyr::select(time.1, contrast.1, data) |>
      tidyr::unnest(cols = data)
  })) |>
  dplyr::select(cluster_id, jacmtx)

df_jacmtx |>
  tidyr::unnest(cols = jacmtx) |>
  download_table("df_jacmtx")
```

```{r results='asis'}
df_jacmtx |>
  purrr::pwalk(function(cluster_id, jacmtx) {
    cat("#####", as.character(cluster_id), "\n\n")
    anno <- jacmtx |>
      dplyr::select(time.1, contrast.1) |>
      dplyr::rename(time = time.1, contrast = contrast.1) |>
      dplyr::distinct()

    mtx <- jacmtx |>
      dplyr::mutate(
        set.1 = paste(contrast.1, time.1, sep = " - "),
        set.2 = paste(contrast.2, time.2, sep = " - ")) |>
      dplyr::select(set.1, set.2, jaccard) |>
      tidyr::pivot_wider(names_from = set.2, values_from = jaccard) |>
      tibble::column_to_rownames("set.1") |>
      as.matrix()

    stopifnot(identical(
      colnames(mtx),
      paste(anno$contrast, anno$time, sep = " - ")))

    hac <- ComplexHeatmap::HeatmapAnnotation(
        which = "column",
        df = as.data.frame(anno),
        col = list(
          contrast = structure(
            c("#E41A1C", "#377EB8"),
            names = levels(anno$contrast)
          ),
          time = structure(
            RColorBrewer::brewer.pal(3, "YlGn"),
            names = levels(anno$time)
          )
        ))

    har <- ComplexHeatmap::HeatmapAnnotation(
        which = "row",
        df = as.data.frame(anno),
        col = list(
          contrast = structure(
            c("#E41A1C", "#377EB8"),
            names = levels(anno$contrast)
          ),
          time = structure(
            RColorBrewer::brewer.pal(3, "YlGn"),
            names = levels(anno$time)
          )
        ),
        show_legend = FALSE,
        show_annotation_name = FALSE
      )
    
    p <- ComplexHeatmap::Heatmap(
      matrix = mtx,
      name = "Jaccard's coef",
      col = circlize::colorRamp2(c(0, 0.5), c("white", "red")),
      na_col = "white",
      column_title = as.character(cluster_id),
      show_column_names = FALSE,
      show_row_names = FALSE,
      clustering_distance_rows = \(x) as.dist(1 - x),
      clustering_distance_columns = \(x) as.dist(1 - x),
      top_annotation = hac,
      left_annotation = har,
      cell_fun = function(j, i, x, y, width, height, fill) {
        value <- mtx[i, j]
        if (!is.na(value))
          grid::grid.text(
            sprintf("%.0f%%", value * 100),
            x, y, gp = grid::gpar(fontsize = 10)
          )
      })
    
    ComplexHeatmap::draw(p, merge_legend = TRUE)
    cat("\n\n")
  })
```