---
title: "Transcription Factor Regulation Network of DEGs"
author: "Altair Wei"
date: "2023-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../scripts/UtilityFunctions.R")
```


## Load Data

### Metadata

```{r}
cell_types <- c(
  "Stomata",
  "Epidermis",
  "Chlorenchyma",
  "Parenchyma",
  "Outer sheath",
  "Inner sheath",
  "Phloem",
  "Procambium"
)
```

### DEGs of Cell Types

```{r}
ds_res <- readRDS(Sys.glob(
  "../results/ObjectCache/DifferentialState/ds_res_*.rds"))

ds_df <- ds_res$DESeq2 |>
  rhapsodykit::diff_state_significant(fdr_limit = 0.05, logfc_limit = 0) |>
  rhapsodykit::diff_state_format()
```

```{r}
marker_df <- readRDS(Sys.glob(
  "../results/ObjectCache/DifferentialState/ds_identity_genes_*.rds")) |>
  dplyr::filter(
    # Min in group fraction
    pct.1 > 0.2,
    # Max out group fraction
    pct.2 < 0.2) |>
  dplyr::group_by(time, cluster) |>
  dplyr::top_n(n = 300, wt = avg_log2FC)
```

### Functions

```{r}
enrichTFs <- function(genes, timeout = 60, browse = FALSE) {
  job_id = as.character(as.numeric(Sys.time()) * 1000000000)

  RCurl::postForm(
    uri = "http://bioinfo.sibs.ac.cn/Wheat-RegNet/wait.php",
    style = "httppost",
    .params = list(
      method = "gene_list_anno_tf",
      job_id = job_id,
      gene_list_id = "",
      upload_gene_file = RCurl::fileUpload(
        filename = "query_genes.txt",
        contents = paste0(genes, collapse = "\n"),
        contentType = "text/plain"))
  )

  message("Running task...")
  RCurl::postForm(
    uri = "http://bioinfo.sibs.ac.cn/Wheat-RegNet/run_web_task.php",
    style = "post",
    .opts = list(timeout = timeout),
    .params = list(
      method = "gene_list_anno_tf",
      job_id = job_id,
      input_file = "gene_list_id_txt"
    )
  )

  result_url <- paste0("http://bioinfo.sibs.ac.cn/Wheat-RegNet/result.php?job_id=", job_id)

  if (browse)
    utils::browseURL(result_url)

  df <- rvest::read_html(result_url) |>
    rvest::html_element("#tf_list_table") |>
    rvest::html_table() |>
    dplyr::select(-Checkbox)

  df
}
```

```{r}
expandToMatrix <- function(df, name_cols, value_col, rownames_col) {
  df <- df |>
    tidyr::pivot_wider(
      names_from = {{name_cols}},
      names_sep = "#",
      names_expand = TRUE,
      values_from = {{value_col}},
      values_fill = NA) |>
    tidyr::pivot_longer(
      cols = -TF,
      names_to = {{name_cols}},
      names_sep = "#",
      values_to = {{value_col}})

  df |>
    tidyr::pivot_wider(names_from = {{name_cols}},
                       names_sep = "#",
                       values_from = {{value_col}}) |>
    tibble::column_to_rownames(rownames_col) |>
    as.matrix()
}
```

```{r}
plotTFsEnrichHeatmap <- function(mtx, max = 10, ...) {
  outliers_idx <- which(mtx > max, arr.ind = TRUE)
  if (length(outliers_idx) != 0)
    warning(sprintf(paste0("%i TF%s with FoldChange greater than %i were shown",
                    " in maximum color and marked with white asterisk.\n"),
                    nrow(outliers_idx),
                    if (nrow(outliers_idx) > 1) "s" else "",
                    max))

  dend_mtx <- mtx
  dend_mtx[is.na(dend_mtx)] <- 0
  dend_row_hc <- dendsort::dendsort(
    fastcluster::hclust(dist(dend_mtx)),
    isReverse = TRUE)

  p <- ComplexHeatmap::Heatmap(
    matrix = mtx,
    name = "Ratio Fold Change",
    col = circlize::colorRamp2(c(0, max), c("blue", "red")),
    use_raster = TRUE,
    na_col = "grey",
    cluster_rows = dend_row_hc,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    show_row_names = TRUE,
    heatmap_legend_param = list(
      title_position = "leftcenter-rot",
      legend_height  = grid::unit(4, "cm")
    ),
    cell_fun = function(j, i, x, y, width, height, fill) {
      val <- abs(mtx[i, j])
      if (!is.na(val) && val > max)
        grid::grid.text("★", x, y, gp = grid::gpar(
          fontsize = 8, col = "white", fontface = "bold"))
    },
    ...)

  ComplexHeatmap::draw(p, merge_legend = TRUE)
}
```

## Overlapping of Itentity and Immunity

### Combine DEG list

```{r}
df_imm <- ds_df |>
  dplyr::filter(!stringr::str_detect(contrast, "X\\dDPI\\.PNR2-X\\dDPI\\.TR4")) |>
  dplyr::select(gene, cluster_id, contrast, logFC) |>
  dplyr::mutate(
    time = stringr::str_extract(contrast, "\\dDPI"),
    contrast = stringr::str_replace(contrast,
      "X(\\dDPI)\\.(.*)-X(\\dDPI)\\.(.*)", "\\2-\\4")) |>
  dplyr::select(time, cluster_id, contrast, gene) |>
  dplyr::rename(type = contrast)
```

```{r}
df_ide <- marker_df |>
  dplyr::mutate(type = "identity") |>
  dplyr::select(time, cluster, type, gene) |>
  dplyr::rename(cluster_id = cluster)
```

```{r}
df_deg <- dplyr::bind_rows(df_imm, df_ide)
```

```{r}
str(df_deg)
```


### Venn plot of DEG list {.tabset}

```{r}
plotVennOverlap <- function(df, cluster) {
  venn_list <- df |>
    dplyr::filter(cluster_id == cluster, time != "0DPI") |>
    dplyr::select(time, type, gene) |>
    dplyr::group_by(time, type) |>
    dplyr::summarise(
      genes = list(gene),
      .groups = "drop") |>
    dplyr::group_by(time) |>
    tidyr::nest() |>
    dplyr::rowwise() |>
    dplyr::summarise(
      data = list(tibble::deframe(data)),
      .groups = "drop") |>
    tibble::deframe()


  plots <- mapply(
    venn_list, names(venn_list),
    SIMPLIFY = FALSE,
    USE.NAMES = FALSE,
    FUN = function(lst, name) {
      if (length(lst) < 2) {
        warning("Less than 2 gene sets.")
        return(NULL)
      }
      gridExtra::arrangeGrob(
        plot(eulerr::venn(lst)), top = name)
    })

  gridExtra::grid.arrange(grobs = plots, nrow = 1)
}
```

```{r fig.height=5, fig.width=14, results='asis'}
for (ct in cell_types) {
  cat("####", ct, "\n\n")
  suppressWarnings(plotVennOverlap(df_deg, ct))
  cat("\n\n")
}
```

## Enriched TFs of Marker Genes

### TFs enrich for identity genes

```{r}
TFs_enrich_identity <- xfun::cache_rds(
  file = "ds_identity_TFs_enrich_df.rds",
  dir = "../results/ObjectCache/DifferentialState/",
  expr = marker_df |>
    dplyr::group_by(time, cluster) |>
    dplyr::summarise(DEGs = list(gene)) |>
    dplyr::rowwise() |>
    dplyr::filter(length(DEGs) > 50) |>
    dplyr::mutate(DEGs = list(convertV11ToV10(DEGs))) |>
    dplyr::mutate(TFs = list(enrichTFs(DEGs)))
)
```

```{r}
head(TFs_enrich_identity)
```

```{r}
df_to_plot <- TFs_enrich_identity |>
  dplyr::select(-DEGs) |>
  tidyr::unnest(TFs)

df_to_plot
```

### Heatmap of enriched TFs

```{r}
plotIdentityTFsHeatmap <- function(df, max = 10, ...) {
  tf_fc_df <- tibble::tibble(
    time = factor(df$time, levels = c("0DPI", "1DPI", "2DPI", "3DPI")),
    cluster_id = factor(df$cluster, levels = cell_types),
    TF = df$TF,
    FoldChange = df$FoldChange
  )

  tf_fc_mtx <- expandToMatrix(
    tf_fc_df, c("cluster_id", "time"), "FoldChange", "TF")

  str_coldata <- strsplit(colnames(tf_fc_mtx), "#")

  mtx_coldata <- data.frame(
    colidx = seq_along(colnames(tf_fc_mtx)),
    cell_types = factor(sapply(str_coldata, function(x) x[[1]]),
                        levels = cell_types),
    times = factor(sapply(str_coldata, function(x) x[[2]]),
                   levels = c("0DPI", "1DPI", "2DPI", "3DPI"))
  )

  mtx_coldata <- mtx_coldata[with(mtx_coldata, order(cell_types, times)),]

  mtx <- tf_fc_mtx[, mtx_coldata$colidx]

  ha = ComplexHeatmap::HeatmapAnnotation(
        which = "column",
        df = mtx_coldata[-1],
        col = list(
          cell_types = structure(
            ggthemes::tableau_color_pal()(
              length(levels(mtx_coldata$cell_types))),
            names = levels(mtx_coldata$cell_types)
          ),
          times = structure(
            RColorBrewer::brewer.pal(4, "YlGn"),
            names = levels(mtx_coldata$times)
          )
        ))

  plotTFsEnrichHeatmap(
    mtx,
    top_annotation = ha,
    column_split = mtx_coldata$cell_types,
    ...)
}
```

```{r fig.height=10, fig.width=14}
plotIdentityTFsHeatmap(df_to_plot)
```


## Enriched TFs of Immunity DEGs

### Generate DEG lists

```{r}
logfc_df <- ds_res$DESeq2 |>
  rhapsodykit::diff_state_significant(fdr_limit = 0.05, logfc_limit = 0) |>
  rhapsodykit::diff_state_format() |>
  dplyr::filter(!stringr::str_detect(contrast, "X\\dDPI\\.PNR2-X\\dDPI\\.TR4")) |>
  dplyr::select(gene, cluster_id, contrast, logFC) |>
  dplyr::mutate(
    time = stringr::str_extract(contrast, "\\dDPI"),
    contrast = stringr::str_replace(contrast,
      "X(\\dDPI)\\.(.*)-X(\\dDPI)\\.(.*)", "\\2-\\4")) |>
  tidyr::pivot_wider(names_from = "time", values_from = "logFC", values_fill = NA) |>
  tidyr::pivot_longer(cols = dplyr::ends_with("DPI"), names_to = "time", values_to = "logFC")

str(logfc_df)
```

```{r}
DEGs_list_df <- logfc_df |>
  dplyr::filter(!is.na(logFC)) |>
  dplyr::mutate(regu_type = ifelse(logFC > 0, "UP", "DOWN")) |>
  dplyr::select(-logFC) |>
  dplyr::group_by(time, contrast, cluster_id, regu_type) |>
  dplyr::summarise(DEGs = list(gene), .groups = "drop")
```

### TFs enrich for DEG lists

FoldChange 是指 GeneRatio/BgRatio

只有 DEGs 数量大于 50 的列表才用于 TFs 富集分析。

```{r}
TFs_enrich_df <- xfun::cache_rds(
  file = "ds_DEGs_TFs_enrich_df.rds",
  dir = "../results/ObjectCache/DifferentialState/",
  expr = DEGs_list_df |>
    dplyr::rowwise() |>
    dplyr::filter(length(DEGs) > 50) |>
    dplyr::mutate(DEGs = list(convertV11ToV10(DEGs))) |>
    dplyr::mutate(TFs = list(enrichTFs(DEGs)))
)
```

```{r}
head(TFs_enrich_df)
```

```{r}
df_to_plot <- TFs_enrich_df |>
  dplyr::select(-DEGs) |>
  tidyr::unnest(TFs)

df_to_plot
```

### Bubble Chart {.tabset}

#### Up-regulated DEGs

```{r fig.height=7, fig.width=14}
df_to_plot |>
  dplyr::filter(regu_type == "UP") |>
  ggplot2::ggplot(
    mapping = ggplot2::aes(
      x = cluster_id,
      y = TF, color = Pvalue, size = `Ratio(%)`
    )) +
  ggplot2::geom_point() +
  ggplot2::scale_color_continuous(
    low=  "red", high = "blue", name = "P-value",
    guide = ggplot2::guide_colorbar(reverse = TRUE)) +
  ggplot2::facet_wrap(ggplot2::vars(time, contrast), scale = "free_x", nrow = 1) +
  rotate_x_labels()
```

#### Down-regulated DEGs

```{r fig.height=7, fig.width=14}
df_to_plot |>
  dplyr::filter(regu_type == "DOWN") |>
  ggplot2::ggplot(
    mapping = ggplot2::aes(
      x = cluster_id,
      y = TF, color = Pvalue, size = `Ratio(%)`
    )) +
  ggplot2::geom_point() +
  ggplot2::scale_color_continuous(
    low=  "red", high = "blue", name = "P-value",
    guide = ggplot2::guide_colorbar(reverse = TRUE)) +
  ggplot2::facet_wrap(ggplot2::vars(time, contrast), scale = "free_x", nrow = 1) +
  rotate_x_labels()
```

### Heatmap {.tabset}

```{r}
plotTFsImmunityHeatmap <- function(df, ...) {
  tf_fc_df <- tibble::tibble(
    time = factor(df$time, levels = c("1DPI", "2DPI", "3DPI")),
    contrast = factor(df$contrast, levels = c("PNR2-MOCK", "TR4-MOCK")),
    cluster_id = factor(df$cluster_id, levels = cell_types),
    TF = df$TF,
    FoldChange = df$FoldChange
  )

  tf_fc_mtx <- expandToMatrix(
    tf_fc_df, c("cluster_id", "contrast", "time"), "FoldChange", "TF")

  str_coldata <- strsplit(colnames(tf_fc_mtx), "#")

  mtx_coldata <- data.frame(
    colidx = seq_along(colnames(tf_fc_mtx)),
    cell_types = factor(sapply(str_coldata, function(x) x[[1]]),
                        levels = cell_types),
    contrast = factor(sapply(str_coldata, function(x) x[[2]]),
                      #levels = c("PNR2-MOCK", "TR4-MOCK", "PNR2-TR4")),
                      levels = c("PNR2-MOCK", "TR4-MOCK")),
    times = factor(sapply(str_coldata, function(x) x[[3]]),
                   levels = c("1DPI", "2DPI", "3DPI"))
  )

  mtx_coldata <- mtx_coldata[with(mtx_coldata, order(cell_types, contrast, times)),]

  mtx <- tf_fc_mtx[, mtx_coldata$colidx]

  ha = ComplexHeatmap::HeatmapAnnotation(
        which = "column",
        df = mtx_coldata[-1],
        col = list(
          cell_types = structure(
            ggthemes::tableau_color_pal()(
              length(levels(mtx_coldata$cell_types))),
            names = levels(mtx_coldata$cell_types)
          ),
          contrast = structure(
            c("#E41A1C", "#377EB8"),
            names = levels(mtx_coldata$contrast)
          ),
          times = structure(
            RColorBrewer::brewer.pal(3, "YlGn"),
            names = levels(mtx_coldata$times)
          )
        ))

  plotTFsEnrichHeatmap(
    mtx,
    top_annotation = ha,
    column_split = mtx_coldata$cell_types,
    ...)
}
```

#### Up-regulated DEGs

```{r fig.height=7, fig.width=14}
df_to_plot |>
  dplyr::filter(regu_type == "UP") |>
  plotTFsImmunityHeatmap()
```

#### Down-regulated DEGs

```{r fig.height=10, fig.width=14}
df_to_plot |>
  dplyr::filter(regu_type == "DOWN") |>
  plotTFsImmunityHeatmap()
```

## Interaction of Two Networks

### Combine TF list

```{r}
df_id <- TFs_enrich_identity |>
  dplyr::filter(time != "0DPI") |>
  dplyr::mutate(type = "identity", cluster_id = cluster) |>
  dplyr::select(time, type, cluster_id, TFs) |>
  tidyr::unnest(TFs)
```

```{r}
df_up <- TFs_enrich_df |>
  dplyr::filter(regu_type == "UP") |>
  dplyr::mutate(type = contrast) |>
  dplyr::select(time, type, cluster_id, TFs) |>
  tidyr::unnest(TFs)
```

```{r}
df_down <- TFs_enrich_df |>
  dplyr::filter(regu_type == "DOWN") |>
  dplyr::mutate(type = contrast) |>
  dplyr::select(time, type, cluster_id, TFs) |>
  tidyr::unnest(TFs)
```

### Heatmap of overlap TFs {.tabset}

```{r}
plotTFsOverlap <- function(df, ...) {
  tf_fc_df <- tibble::tibble(
    time = factor(df$time, levels = c("1DPI", "2DPI", "3DPI")),
    type = factor(df$type, levels = c("identity", "PNR2-MOCK", "TR4-MOCK")),
    cluster_id = factor(df$cluster_id, levels = cell_types),
    TF = df$TF,
    FoldChange = df$FoldChange
  )

  tf_fc_mtx <- expandToMatrix(
    tf_fc_df, c("cluster_id", "type", "time"), "FoldChange", "TF")

  str_coldata <- strsplit(colnames(tf_fc_mtx), "#")

  mtx_coldata <- data.frame(
    colidx = seq_along(colnames(tf_fc_mtx)),
    cell_types = factor(sapply(str_coldata, function(x) x[[1]]),
                        levels = cell_types),
    type = factor(sapply(str_coldata, function(x) x[[2]]),
                      levels = c("identity", "PNR2-MOCK", "TR4-MOCK")),
    times = factor(sapply(str_coldata, function(x) x[[3]]),
                   levels = c("1DPI", "2DPI", "3DPI"))
  )

  mtx_coldata <- mtx_coldata[with(mtx_coldata, order(cell_types, type, times)),]

  mtx <- tf_fc_mtx[, mtx_coldata$colidx]

  ha = ComplexHeatmap::HeatmapAnnotation(
        which = "column",
        df = mtx_coldata[-1],
        col = list(
          cell_types = structure(
            ggthemes::tableau_color_pal()(
              length(levels(mtx_coldata$cell_types))),
            names = levels(mtx_coldata$cell_types)
          ),
          type = structure(
            c("#8856a7", "#E41A1C", "#377EB8"),
            names = levels(mtx_coldata$type)
          ),
          times = structure(
            RColorBrewer::brewer.pal(3, "YlGn"),
            names = levels(mtx_coldata$times)
          )
        ))

  plotTFsEnrichHeatmap(
    mtx,
    top_annotation = ha,
    column_split = mtx_coldata$cell_types,
    ...)

}
```

#### Up-regulated

```{r fig.height=8, fig.width=18}
plotTFsOverlap(
  df = dplyr::bind_rows(df_id, df_up))
```

#### Down-regulated

```{r fig.height=10, fig.width=18}
plotTFsOverlap(
  df = dplyr::bind_rows(df_id, df_down)
)
```
