---
title: "Evaluating Cluster Stability of MOCK Samples"
author: "Altair Wei"
date: "2023-05-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../scripts/LoadUtils.R", chdir = TRUE)
library(ggtree, include.only = "%<+%")
```

## Load Data

### Seurat Object

npcs = 20, k = 20, resolution = 0.6

```{r}
obj_mock <- readRDS(Sys.glob(
  "../results/ObjectCache/IntegrationAndAnnotation/obj_mock_annotated_*.rds"))
```

### Pipeline Outputs

```{r}
subsample_idents <- readRDS("../results/ClustEval/gather_subsample.rds")
fullsample_idents <- readRDS("../results/ClustEval/gather_fullsample.rds")
```

```{r}
subsample_2nd <- readRDS("../results/ClustEval-2nd/gather_subsample.rds")
fullsample_2nd <- readRDS("../results/ClustEval-2nd/gather_fullsample.rds")
```

### Functions

```{r}
selectParamSet <- function(fullsample, ...) {
  fullsample |>
    dplyr::filter(...) |>
    dplyr::pull(original_ident_full) |>
    purrr::pluck(1)
}

selectParamSetRow <- function(fullsample, pc, k, res) {
  list(
    ident1 = fullsample |>
      dplyr::filter(
        pc == !!pc,
        k_param == k,
        resolution == res) |>
      dplyr::pull(original_ident_full) |>
      purrr::pluck(1),
    row_title = sprintf("cluster id w/ pc=%s k=%s res=%s", pc, k, res),
    row_title_side = "right"
  )
}

selectParamSetColumn <- function(fullsample, pc, k, res) {
  list(
    ident2 = fullsample |>
      dplyr::filter(
        pc == !!pc,
        k_param == k,
        resolution == res) |>
      dplyr::pull(original_ident_full) |>
      purrr::pluck(1),
    title = sprintf("cluster id w/ pc=%s k=%s res=%s", pc, k, res),
    column_title_side = "bottom"
  )
}

calcClusteringDistance <- function(ident_list) {
  n <- length(ident_list)
  w <- matrix(0, nrow = n, ncol = n)
  colnames(w) <- rownames(w) <- names(ident_list)
  for (i in seq_len(n - 1)) {
    for (j in (i + 1):n) {
      w[i, j] <- mcclust::vi.dist(
        ident_list[[i]], ident_list[[j]])
    }
  }

  w2 <- w + t(w)
  for (i in seq_len(nrow(w2)))
      w2[i, i] <- 1

  as.dist(w2)
}

calcClusteringDistance2 <- function(
    ident_list, BPPARAM = BiocParallel::SerialParam()) {

  n <- length(ident_list)
  w <- matrix(0, nrow = n, ncol = n)
  colnames(w) <- rownames(w) <- names(ident_list)
  arrIndList <- which(upper.tri(w), arr.ind = TRUE, useNames = TRUE) |>
    apply(1, \(x) x, simplify = FALSE)

  distList <- BiocParallel::bplapply(
    X = arrIndList,
    FUN = function(ind) mcclust::vi.dist(
        ident_list[[ind["row"]]], ident_list[[ind["col"]]]),
    BPPARAM = BPPARAM
  )

  for (i in seq_along(arrIndList))
    w[arrIndList[[i]][["row"]],
      arrIndList[[i]][["col"]]] <- distList[[i]]

  w2 <- w + t(w)
  for (i in seq_len(nrow(w2)))
      w2[i, i] <- 1

  as.dist(w2)
}
```

## Visualization of First Run

We want to maximize the number of clusters but also want the majority of the cells to be in stable clusters.

### Scatter for All Parameter Sets

TODO: 使用 clustree::clustree 来探索某个参数的增加对细胞分群的影响。

使用 0.8 作为阈值有点太高了，试试 0.6 吧。

> As a rule of thumb, clusters with a mean/median stability score less than 0.6 should be considered unstable. scores between 0.6 and 0.75 indicate that the cluster is measuring a pattern in the data. clusters with stability score greater than 0.85 are highly stable (Zumel and Mount 2014).
>
> We observed for some datasets, the jaccard index follows a bimodal distribution, so the mean or median may not be representative. As an alternative, we also calculate the percentage of subsampling with a jaccard greater than a cutoff (e.g. 0.85), which can be used to check stability assessments.

```{r}
subsample_idents_list <- subsample_idents |>
  dplyr::group_by(pc, resolution, k_param) |> 
  tidyr::nest()
```

```{r}
stable_clusters <- xfun::cache_rds(
  file = "df_mock_stable_clusters.rds",
  dir = "../results/ObjectCache/IntegrationAndAnnotation/",
  expr = subsample_idents_list |>
    dplyr::mutate(
      stable_cluster = purrr::map(
        data,
        ~ scclusteval::AssignStableCluster(
          .x$original_ident,
          .x$recluster_ident,
          jaccard_cutoff = 0.6,
          method = "jaccard_percent", 
          percent_cutoff = 0.8)
      )
    )
)
```

#### Number of clusters {.tabset}

##### x-axis = resolution

```{r fig.height=14, fig.width=14}
p <- scclusteval::ParameterSetScatterPlot(
    stable_clusters = stable_clusters,
    fullsample_idents = fullsample_idents,
    x_var = "resolution",
    y_var = "number",
    facet_rows = "pc",
    facet_cols = "k_param") +
  ggplot2::ggtitle("Number of stable/total clusters")

p
```

##### x-axis = k param

```{r fig.height=14, fig.width=14}
p <- scclusteval::ParameterSetScatterPlot(
    stable_clusters = stable_clusters,
    fullsample_idents = fullsample_idents,
    x_var = "k_param",
    y_var = "number",
    facet_rows = "pc",
    facet_cols = "resolution") +
  ggplot2::ggtitle("Number of stable/total clusters")

p
```

##### x-axis = PCs

```{r fig.height=14, fig.width=14}
p <- scclusteval::ParameterSetScatterPlot(
    stable_clusters = stable_clusters,
    fullsample_idents = fullsample_idents,
    x_var = "pc",
    y_var = "number",
    facet_rows = "k_param",
    facet_cols = "resolution") +
  ggplot2::ggtitle("Number of stable/total clusters")

p
```

#### Percentage of cells {.tabset}

##### x-axis = resolution

```{r fig.height=14, fig.width=14}
p <- scclusteval::ParameterSetScatterPlot(
    stable_clusters = stable_clusters,
    fullsample_idents = fullsample_idents,
    x_var = "resolution",
    y_var = "percentage",
    facet_rows = "pc",
    facet_cols = "k_param") +
  ggplot2::ggtitle("Percentage of cells in stable clusters")

p
```

##### x-axis = k param

```{r fig.height=14, fig.width=14}
p <- scclusteval::ParameterSetScatterPlot(
    stable_clusters = stable_clusters,
    fullsample_idents = fullsample_idents,
    x_var = "k_param",
    y_var = "percentage",
    facet_rows = "pc",
    facet_cols = "resolution") +
  ggplot2::ggtitle("Percentage of cells in stable clusters")

p
```

##### x-axis = PCs

```{r fig.height=14, fig.width=14}
p <- scclusteval::ParameterSetScatterPlot(
    stable_clusters = stable_clusters,
    fullsample_idents = fullsample_idents,
    x_var = "pc",
    y_var = "percentage",
    facet_rows = "k_param",
    facet_cols = "resolution") +
  ggplot2::ggtitle("Percentage of cells in stable clusters")

p
```

#### Overview Scatter

```{r}
df_perf <- dplyr::left_join(stable_clusters, fullsample_idents) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(total = purrr::map_dbl(stable_cluster, ~ length(.x$stable_cluster))) %>%
  dplyr::mutate(stable = purrr::map_dbl(stable_cluster, ~ .x$number_of_stable_cluster)) %>%
  dplyr::mutate(percentage = purrr::map2_dbl(
    original_ident_full, stable_cluster,
    function(x, y) scclusteval::CalculatePercentCellInStable(x, y$stable_cluster))) %>%
  dplyr::select(-data, -stable_cluster, -original_ident_full) %>%
  dplyr::mutate_if(is.character, function(x) as.factor(as.numeric(x))) %>%
  tidyr::gather(total:stable , key = "category", value = "number")
```

```{r fig.height=7, fig.width=14}
p <- df_perf |>
  dplyr::filter(category == "total") |>
  ggplot2::ggplot(ggplot2::aes(
    x = number, y = percentage,
    color = resolution, size = k_param)) +
  ggplot2::geom_point(alpha = 0.3) +
  ggplot2::facet_grid(~ pc) +
  ggplot2::xlab("Total number of clusters") +
  ggplot2::ylab("Percentage of cells in stable clusters")

p
```

```{r ig.height=7, fig.width=9}
plotly::ggplotly(p)
```

### Similarity of Clusterings

#### Cladogram

```{r}
clust_dist <- xfun::cache_rds(
  file = "dist_mock_fullsample_paramsets.rds",
  dir = "../results/ObjectCache/IntegrationAndAnnotation/",
  expr = fullsample_idents |>
    dplyr::mutate(label = sprintf(
      "pc=%s, k=%s, res=%s", pc, k_param, resolution)) |>
    dplyr::select(label, original_ident_full) |>
    tibble::deframe() |>
    calcClusteringDistance2(BPPARAM = BiocParallel::MulticoreParam(8))
)
```

```{r}
df_anno <- df_perf |>
  tidyr::pivot_wider(names_from = "category", values_from = "number") |>
  dplyr::mutate(label = sprintf(
      "pc=%s, k=%s, res=%s", pc, k_param, resolution)) |>
  tibble::column_to_rownames("label")
```

```{r fig.height=12, fig.width=14}
fit <- hclust(clust_dist, method = "complete")
clus <- cutree(fit, 7)
g <- split(names(clus), clus)

p <- ggtree::ggtree(fit, show.legend = FALSE, layout = "circular")
clades <- sapply(g, \(n) ggtree::MRCA(p, n))
p <- ggtree::groupClade(p, clades, group_name = "subtree") +
  ggplot2::aes(color = subtree) +
  ggnewscale::new_scale_color()

pdata <- tibble::rownames_to_column(
  df_anno[names(clus),], "label")

p <- p %<+% pdata +
  ggtree::geom_tippoint(
    mapping = ggplot2::aes(size = total, color = percentage)) +
  ggplot2::scale_color_distiller(palette = "BuGn", direction = 1) +
  ggnewscale::new_scale_color() +
  ggtree::geom_tiplab(
    mapping = ggplot2::aes(color = percentage > 0.8),
    size = 3, hjust = -0.1) +
  ggplot2::scale_color_manual(values = c("black", "red")) +
  NULL


p
```

#### Check tree

Two distinct clades:

```{r}
rlang::inject(
  scclusteval::PairWiseJaccardSetsHeatmap(
    !!!selectParamSetRow(fullsample_idents,
      pc = 35, k = 100, res = 1),
    !!!selectParamSetColumn(fullsample_idents,
      pc = 35, k = 8, res = 0.2),
    show_row_dend = F, show_column_dend = F,
    cluster_row = F, cluster_column = F)
)
```

Two closed clades:

```{r}
rlang::inject(
  scclusteval::PairWiseJaccardSetsHeatmap(
    !!!selectParamSetRow(fullsample_idents,
      pc = 35, k = 100, res = 1),
    !!!selectParamSetColumn(fullsample_idents,
      pc = 30, k = 80, res = 0.8),
    show_row_dend = F, show_column_dend = F,
    cluster_row = F, cluster_column = F)
)
```

Two nearly identical results:

```{r}
rlang::inject(
  scclusteval::PairWiseJaccardSetsHeatmap(
    !!!selectParamSetRow(fullsample_idents,
      pc = 20, k = 50, res = 0.4),
    !!!selectParamSetColumn(fullsample_idents,
      pc = 20, k = 30, res = 0.4),
    show_row_dend = F, show_column_dend = F,
    cluster_row = F, cluster_column = F)
)
```

#### Existed param sets

```{r}
rlang::inject(
  scclusteval::PairWiseJaccardSetsHeatmap(
    !!!selectParamSetRow(fullsample_idents,
      pc = 20, k = 50, res = 0.4),
    !!!selectParamSetColumn(fullsample_idents,
      pc = 20, k = 20, res = 0.6),
    show_row_dend = F, show_column_dend = F,
    cluster_row = F, cluster_column = F)
)
```

```{r}
suppressWarnings(rlang::inject(
  scclusteval::PairWiseJaccardSetsHeatmap(
    !!!selectParamSetRow(
      fullsample_idents, pc = 20, k = 30, res = 0.4),
    Seurat::Idents(obj_mock),
    title = "cluster id w/ pc=20 k=20 res=0.6",
    column_title_side = "bottom",
    show_row_dend = F, show_column_dend = F,
    cluster_row = F, cluster_column = F)
))
```

### Jaccard Raincloud for Subsamples

How to interpret?

> With resolution of 0.05, Seurat finds 5 clusters which is optimal given that 5 cancer cell lines were mixed together. The jaccard indices are close to 1 for all clusters after subsampling and reclustering suggesting all 5 clusters are quite stable.
>
> From the Jaccard raincloud plot, cluster 0,5 (res = 0.1) have dropped jaccard similarity index. cluster 4,6 (res = 0.1) also have dropped jaccard similarity index, suggesting that the orginal cluster 0 and cluster 4 (res - 0.05) should not be further splitted.
>
> Increasing resolution will always give more clusters and whenever we observe a bimodal distribution of jaccard in a cluster when increasing the resolution, it indicates this cluster can be merged with a different cluster in the resampling and reclustering procedure

```{r}
subsample_idents_list2 <- subsample_idents_list |>
  dplyr::rowwise() |>
  dplyr::mutate(
    plot = list(
      scclusteval::JaccardRainCloudPlot(
        data$original_ident, data$recluster_ident) +
      ggplot2::geom_hline(yintercept = c(0.6, 0.75), linetype = 2) +
      ggplot2::ggtitle(sprintf(
        "cluster id w/ pc=%s k=%s res=%s", pc, k_param, resolution))
    )
  )
```

#### Two param sets under pc=10

```{r fig.height=14, fig.width=10}
subsample_idents_list2 |>
  dplyr::filter(pc == 10, (k_param == 80 & resolution == 0.4) | (k_param == 8 & resolution == 0.2)) |>
  dplyr::pull(plot) |>
  patchwork::wrap_plots(ncol = 1)
```

```{r}
scclusteval::PairWiseJaccardSetsHeatmap(
  selectParamSet(fullsample_idents,
    pc == 15, k_param == 50, resolution == 0.4),
  selectParamSet(fullsample_idents,
    pc == 15, k_param == 80, resolution == 0.4),
  show_row_dend = F, show_column_dend = F,
  cluster_row = F, cluster_column =F)
```

```{r}
scclusteval::ClusterIdentityChordPlot(
  selectParamSet(fullsample_idents,
    pc == 10, k_param == 80, resolution == 0.4),
  selectParamSet(fullsample_idents,
    pc == 20, k_param == 100, resolution == 0.4)
)
```

```{r include=FALSE}
rm(subsample_idents, fullsample_idents)
gc()
```

## Visualization of Second Run

### Calculating

```{r}
clust_dist_2nd <- xfun::cache_rds(
  file = "dist_mock_fullsample_paramsets_2nd.rds",
  dir = "../results/ObjectCache/IntegrationAndAnnotation/",
  expr = fullsample_2nd |>
    dplyr::mutate(label = sprintf(
      "pc=%s, k=%s, res=%s", pc, k_param, resolution)) |>
    dplyr::select(label, original_ident_full) |>
    tibble::deframe() |>
    calcClusteringDistance2(BPPARAM = BiocParallel::MulticoreParam(8))
)
```

```{r}
subsample_idents_list <- subsample_2nd |>
  dplyr::group_by(pc, resolution, k_param) |> 
  tidyr::nest()
```

```{r}
stable_clusters <- xfun::cache_rds(
  file = "df_mock_stable_clusters_2nd.rds",
  dir = "../results/ObjectCache/IntegrationAndAnnotation/",
  expr = subsample_idents_list |>
    dplyr::mutate(
      stable_cluster = purrr::map(
        data,
        ~ scclusteval::AssignStableCluster(
          .x$original_ident,
          .x$recluster_ident,
          jaccard_cutoff = 0.6,
          method = "jaccard_percent", 
          percent_cutoff = 0.8)
      )
    )
)

gc()
```

```{r}
df_perf <- dplyr::left_join(stable_clusters, fullsample_2nd) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(total = purrr::map_dbl(stable_cluster, ~ length(.x$stable_cluster))) %>%
  dplyr::mutate(stable = purrr::map_dbl(stable_cluster, ~ .x$number_of_stable_cluster)) %>%
  dplyr::mutate(percentage = purrr::map2_dbl(
    original_ident_full, stable_cluster,
    function(x, y) scclusteval::CalculatePercentCellInStable(x, y$stable_cluster))) %>%
  dplyr::select(-data, -stable_cluster, -original_ident_full) %>%
  dplyr::mutate_if(is.character, function(x) as.factor(as.numeric(x))) %>%
  tidyr::gather(total:stable , key = "category", value = "number")
```

```{r}
df_anno <- df_perf |>
  tidyr::pivot_wider(names_from = "category", values_from = "number") |>
  dplyr::mutate(label = sprintf(
      "pc=%s, k=%s, res=%s", pc, k_param, resolution)) |>
  tibble::column_to_rownames("label")
```

### Scatter for All Parameter Sets

#### Number of clusters {.tabset}

##### x-axis = resolution

```{r fig.height=14, fig.width=14}
p <- scclusteval::ParameterSetScatterPlot(
    stable_clusters = stable_clusters,
    fullsample_idents = fullsample_2nd,
    x_var = "resolution",
    y_var = "number",
    facet_rows = "pc",
    facet_cols = "k_param") +
  ggplot2::ggtitle("Number of stable/total clusters")

p
```

##### x-axis = k param

```{r fig.height=14, fig.width=14}
p <- scclusteval::ParameterSetScatterPlot(
    stable_clusters = stable_clusters,
    fullsample_idents = fullsample_2nd,
    x_var = "k_param",
    y_var = "number",
    facet_rows = "pc",
    facet_cols = "resolution") +
  ggplot2::ggtitle("Number of stable/total clusters")

p
```

##### x-axis = PCs

```{r fig.height=14, fig.width=14}
p <- scclusteval::ParameterSetScatterPlot(
    stable_clusters = stable_clusters,
    fullsample_idents = fullsample_2nd,
    x_var = "pc",
    y_var = "number",
    facet_rows = "k_param",
    facet_cols = "resolution") +
  ggplot2::ggtitle("Number of stable/total clusters")

p
```

#### Percentage of cells {.tabset}

##### x-axis = resolution

```{r fig.height=14, fig.width=14}
p <- scclusteval::ParameterSetScatterPlot(
    stable_clusters = stable_clusters,
    fullsample_idents = fullsample_2nd,
    x_var = "resolution",
    y_var = "percentage",
    facet_rows = "pc",
    facet_cols = "k_param") +
  ggplot2::ggtitle("Percentage of cells in stable clusters")

p
```

##### x-axis = k param

```{r fig.height=14, fig.width=14}
p <- scclusteval::ParameterSetScatterPlot(
    stable_clusters = stable_clusters,
    fullsample_idents = fullsample_2nd,
    x_var = "k_param",
    y_var = "percentage",
    facet_rows = "pc",
    facet_cols = "resolution") +
  ggplot2::ggtitle("Percentage of cells in stable clusters")

p
```

##### x-axis = PCs

```{r fig.height=14, fig.width=14}
p <- scclusteval::ParameterSetScatterPlot(
    stable_clusters = stable_clusters,
    fullsample_idents = fullsample_2nd,
    x_var = "pc",
    y_var = "percentage",
    facet_rows = "k_param",
    facet_cols = "resolution") +
  ggplot2::ggtitle("Percentage of cells in stable clusters")

p
```

#### Overview Scatter

```{r fig.height=7, fig.width=14}
p <- df_perf |>
  dplyr::filter(category == "total") |>
  ggplot2::ggplot(ggplot2::aes(
    x = number, y = percentage,
    color = resolution, size = k_param)) +
  ggplot2::geom_point(alpha = 0.3) +
  ggplot2::facet_grid(~ pc) +
  ggplot2::xlab("Total number of clusters") +
  ggplot2::ylab("Percentage of cells in stable clusters")

p
```

```{r ig.height=7, fig.width=9}
plotly::ggplotly(p)
```

### Similarity of Clusterings

#### Cladogram

```{r fig.height=12, fig.width=14}
fit <- hclust(clust_dist_2nd, method = "complete")
clus <- cutree(fit, 7)
g <- split(names(clus), clus)

p <- ggtree::ggtree(fit, show.legend = FALSE, layout = "circular")
clades <- sapply(g, \(n) ggtree::MRCA(p, n))
p <- ggtree::groupClade(p, clades, group_name = "subtree") +
  ggplot2::aes(color = subtree) +
  ggnewscale::new_scale_color()

pdata <- tibble::rownames_to_column(
  df_anno[names(clus),], "label")

p <- p %<+% pdata +
  ggtree::geom_tippoint(
    mapping = ggplot2::aes(size = total, color = percentage)) +
  ggplot2::scale_color_distiller(palette = "BuGn", direction = 1) +
  ggnewscale::new_scale_color() +
  ggtree::geom_tiplab(
    mapping = ggplot2::aes(color = percentage > 0.8),
    size = 3, hjust = -0.1) +
  ggplot2::scale_color_manual(values = c("black", "red")) +
  NULL

p
```

```{r include=FALSE, results='hide'}
rm(list = ls())
gc()
```

## Downstream Analysis of Optimum

### Load Data

```{r}
samples <- c(
  "0DPI-MOCK-1",
  "0DPI-MOCK-2",
  "1DPI-MOCK-1",
  "1DPI-MOCK-2",
  "2DPI-MOCK-1",
  "2DPI-MOCK-2",
  "3DPI-MOCK-1",
  "3DPI-MOCK-2"
)

names(samples) <- samples
```

```{r}
obj_mock <- lapply(samples, function(sample) readRDS(
  Sys.glob(paste0(
    "../results/ObjectCache/QualityControl/",
    "obj_strained_", sample, "_*.rds"))
  )
)
```

### Data Integration

#### Optimum parameters

```{r}
npcs <- 20
k <- 40
res <- 0.3
```

#### Merge objects

```{r}
obj_mock_combined <- merge(
  x = obj_mock[[1]],
  y = obj_mock[-1],
  add.cell.ids = names(obj_mock)
)
```

```{r}
obj_mock_combined <- obj_mock_combined %>%
  Seurat::NormalizeData(verbose = FALSE) %>%
  Seurat::FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>%
  Seurat::ScaleData(verbose = FALSE)
```

```{r}
obj_mock_combined <- obj_mock_combined %>%
  Seurat::RunPCA(features = NULL, npcs = npcs, verbose = FALSE)
```

#### Run Harmony

```{r fig.height=2.5, fig.width=6}
obj_mock_combined <- obj_mock_combined %>% 
    harmony::RunHarmony(
      group.by.vars = c("time", "sample"),
      theta = c(4, 2),
      kmeans_init_nstart=20, kmeans_init_iter_max=100,
      epsilon.cluster=-Inf, epsilon.harmony=-Inf,
      plot_convergence = TRUE)
```

### Dim Reduction and Clustering

```{r}
obj_mock_combined <- obj_mock_combined %>%
    Seurat::RunUMAP(reduction = "harmony", dims = seq_len(npcs), verbose = FALSE, seed.use = 230816) 
```

```{r}
obj_mock_combined <- obj_mock_combined %>%
    Seurat::FindNeighbors(reduction = "harmony", dims = seq_len(npcs), k.param = k, verbose = FALSE) %>%
    Seurat::FindClusters(resolution = res, verbose = FALSE, random.seed = 230816)
```

### Visualize dim reduction {.tabset}

```{r fig.height=7, fig.width=8}
ident_cols <- Seurat::DiscretePalette(length(levels(Seurat::Idents(obj_mock_combined))), palette = NULL)

Seurat::DimPlot(
        obj_mock_combined, reduction = "umap", cols = ident_cols,
        label = TRUE, label.box = TRUE, label.color = "black", label.size = 2) +
  ggplot2::scale_fill_manual(values = rep("white", length(ident_cols))) +
  ggplot2::coord_fixed() +
  theme_dimred() +
  ggplot2::ggtitle(sprintf("pc=%s, k=%s, res=%s", npcs, k, res)) +
  center_plot_title()
```

### Deconvolution of LCM-seq Samples

```{r}
marker_cosg <- COSG::cosg(
 obj_mock_combined,
 groups = 'all',
 assay = 'RNA',
 slot = 'data',
 mu = 1,
 n_genes_user = 300
)
```

```{r}
LCM_design <- c(
  "VC_1" = "vb.",
  "VC_2" = "vb.",
  "VC_3" = "vb.",
  "CC_1" = "bp.",
  "CC_2" = "bp.",
  "CC_3" = "bp.",
  "MC3_1" = "dp.",
  "MC3_2" = "dp.",
  "MC3_3" = "dp.",
  "MC_1" = "me.",
  "MC_2" = "me.",
  "MC_3" = "me.",
  "EC_1" = "ep.",
  "EC_2" = "ep.",
  "EC_3" = "ep."
)

CIBER <- deconvLCM(
  seurat = obj_mock_combined,
  lcm_file = "../results/LCMSeq/gene_quanti/counts/all.featureCounts",
  markers = unique(unlist(marker_cosg$names)),
  design = LCM_design,
  mc.cores = 4
)
```

```{r fig.height=7, fig.width=7, dev='svg'}
deconvScatter(CIBER, LCM_design, ncol = 3)
```

```{r fig.height=6, fig.width=14}
p2 <- Seurat::DotPlot(obj_mock_combined, features = KNOWN_MARKERS) +
  ggplot2::theme(
    axis.text.x = ggplot2::element_blank(),
    legend.position = "bottom",
    legend.title = ggplot2::element_text(vjust = 1)
  )
p2
```
