---
title: "Figure 5 Specific pathways"
author: "Altair Wei"
date: "2023-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(ComplexHeatmap, include.only = "%v%")
source("../scripts/LoadUtils.R", chdir = TRUE)

options(ggrastr.default.dpi=300)
PT = 1/2.13
FT = 6/17.1
```

## Load Data Objects

### DS Results

```{r}
logfc_data <- readRDS(Sys.glob(
  "../results/ObjectCache/DifferentialState/ds_deg_logfc_data_*.rds"))

ds_res <- readRDS(Sys.glob("../results/ObjectCache/DifferentialState/ds_res_*.rds"))

ds_sig <- ds_res$DESeq2 %>%
  rhapsodykit::diff_state_significant(fdr_limit = 0.05, logfc_limit = 0)
```

### Pseudobulk CPM

```{r}
pseudobulk_cpm_mean <- readRDS(Sys.glob(
  "../results/ObjectCache/Pseudobulk/pseudobulk_cpm_mean_*.rds"))
```

```{r}
cpm_se <- local({
  cpm_df <- as.data.frame(pseudobulk_cpm_mean) |>
      tibble::rownames_to_column("Gene") |>
      tidyr::pivot_longer(-Gene, names_to = "bulk", values_to = "cpm") |>
      tidyr::separate(bulk, sep = "\\.", into = c("treatment", "time", "cellType"))
  
  cpm_df_0 <- dplyr::filter(cpm_df, time == "0DPI")
  
  cpm_df <- cpm_df |>
    dplyr::add_row(dplyr::mutate(cpm_df_0, treatment = "PNR2")) |>
    dplyr::add_row(dplyr::mutate(cpm_df_0, treatment = "TR4"))
  
  cpm_df$cellType <- factor(cpm_df$cellType, levels = TISSUE_TYPES)
  cpm_df$treatment <- factor(cpm_df$treatment, levels = c("MOCK", "PNR2", "TR4"))
  cpm_df$time <- factor(cpm_df$time, levels = c("0DPI", "1DPI", "2DPI", "3DPI"))
  
  cpm_df <- cpm_df |>
    dplyr::select(cellType, treatment, time, Gene, cpm) |>
    dplyr::arrange(cellType, treatment, time)
  
  cpm_coldata <- cpm_df |>
    dplyr::group_by(cellType, treatment, time) |>
    dplyr::summarise(.groups = "drop")
  
  cpm_mtx <- cpm_df |>
    tidyr::pivot_wider(names_from = c(cellType, treatment, time),
                       values_from = cpm, names_sep = "#") |>
    tibble::column_to_rownames("Gene") |>
    as.matrix()
  
  SummarizedExperiment(
    assays = list(
      raw = cpm_mtx,
      scaled = t(scale(t(cpm_mtx)))),
    colData = cpm_coldata)
})
```

## Prepare for Figure 5

### Make A Common Colorbar

```{r}
gene_ids <- Sys.glob("../data/immunity/Module*.xlsx") |>
  lapply(function(xlsx_file) {
    lapply(
      openxlsx::getSheetNames(xlsx_file),
      FUN = function(sheet)
        readxl::read_excel(xlsx_file, sheet) |>
          dplyr::select(Gene)) |>
      dplyr::bind_rows() |>
      dplyr::distinct()
  }) |>
  dplyr::bind_rows() |>
  dplyr::distinct() |>
  dplyr::pull(Gene)

col_fun <- makeColorFunction(
  SummarizedExperiment::assay(cpm_se, "scaled")[gene_ids, ],
  function(n) colorRampPalette(
    c("darkgrey", "grey", "white", "yellow", "red", "darkred"))(n))
```

### Prepare Row Annotation

```{r}
df_rowanno <- ds_sig |>
  rhapsodykit::diff_state_format() |>
  dplyr::mutate(
    time = stringr::str_extract(contrast, "\\dDPI"),
    comparison = stringr::str_replace(contrast,
      "X(\\dDPI)\\.(.*)-X(\\dDPI)\\.(.*)", "\\2 vs. \\4")) |>
  dplyr::filter(comparison != "PNR2 vs. TR4")  |>
  dplyr::group_by(gene) |>
  dplyr::summarise(hits = dplyr::n(), regu_score = sum(logFC / abs(logFC)))

head(df_rowanno)
```

### Heatmap of Core Immunity

```{r}
df_selected <- openxlsx::getSheetNames("../data/immunity/Module01-09.xlsx") |>
  lapply(function(x)
    readxl::read_xlsx("../data/immunity/Module01-09.xlsx", x) |>
      dplyr::mutate(Module = x, Weights = as.numeric(factor(Category)))) |>
  dplyr::bind_rows() |>
  dplyr::rename(gene = Gene)
```

```{r fig.height=20, fig.width=12}
pFun <- local({
  df_genes <- df_selected |>
    dplyr::select(gene, RankName, Module, Category, Label, Weights) |>
    dplyr::distinct() |>
    dplyr::left_join(df_rowanno, by = "gene")

  assay_use = "scaled"
  # Prepare expression matrix
  cpm_sesub <- cpm_se[df_genes$gene, ]
  rowData(cpm_sesub) <- df_genes
  mtx <- SummarizedExperiment::assay(cpm_sesub, assay_use)
  colHA <- makeColumnAnnotation(colData(cpm_sesub))
  rowHA <- makeRowAnnotation(rowData(cpm_sesub))

  # Create heatmap
  p <- ComplexHeatmap::Heatmap(
    matrix = mtx,
    col = col_fun,
    use_raster = TRUE,
    border = TRUE,
    na_col = "grey",

    row_labels = rowData(cpm_sesub)[["Label"]],
    show_row_names = TRUE,
    row_names_gp = grid::gpar(fontsize = 7),
    cluster_row_slices = FALSE,
    cluster_row = TRUE,
    clustering_method_rows = "average",
    row_title_rot = 90,
    row_title_side = "left",
    row_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),
    row_split = rowData(cpm_sesub)[["Module"]],

    top_annotation = colHA,
    right_annotation = rowHA,
    column_split = colData(cpm_sesub)[["cellType"]],
    column_title = c("ST", "EP", "CH", "PA", "VOS", "VIS", "PH", "PR"),
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),

    heatmap_legend_param = list(
      title = switch(assay_use, raw = "CPM", scaled = "Z-score"),
      direction = "horizontal",
      title_position = "topleft",
      title_gp = grid::gpar(fontsize = 20,
                            fontfamily = "Arial"),
      labels_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),
      legend_width = unit(8, "cm"))
  )

  # Wrap drawing function
  function(...) {
    suppressWarnings(
      ComplexHeatmap::draw(p,
        merge_legend = TRUE,
        heatmap_legend_side = "top", 
        annotation_legend_side = "top",
        legend_gap = grid::unit(1.5, "cm"),
        align_heatmap_legend = "heatmap_left",
        align_annotation_legend = "heatmap_left",
        ...
      )
    )
  }
})

p1 <- pFun(show_heatmap_legend = TRUE)
```

```{r include=FALSE, eval=FALSE}
plotpowerpoint(p, "../data/Figure.pptx",
  fun = \() pFun(show_heatmap_legend = FALSE),
  location = grid::unit(c(2.5, 12.82, 24.28, 43.56), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

### Heatmap of Sugar Transport

```{r}
df_selected <- readxl::read_xlsx("../data/immunity/Module10-SugarTransport.xlsx") |>
  dplyr::rename(gene = Gene)
```

```{r fig.height=3, fig.width=12}
pFun <- local({
  df_genes <- df_selected |>
    dplyr::select(gene, RankName, Label) |>
    dplyr::distinct() |>
    dplyr::left_join(df_rowanno, by = "gene")

  assay_use = "scaled"
  # Prepare expression matrix
  cpm_sesub <- cpm_se[df_genes$gene, ]
  rowData(cpm_sesub) <- df_genes
  mtx <- SummarizedExperiment::assay(cpm_sesub, assay_use)
  colHA <- makeColumnAnnotation(colData(cpm_sesub))
  rowHA <- makeRowAnnotation(rowData(cpm_sesub))

  # Create heatmap
  p <- ComplexHeatmap::Heatmap(
    matrix = mtx,
    col = col_fun,
    use_raster = TRUE,
    border = TRUE,
    na_col = "grey",

    row_labels = rowData(cpm_sesub)[["Label"]],
    show_row_names = TRUE,
    row_names_gp = grid::gpar(fontsize = 7),
    cluster_row_slices = FALSE,
    cluster_row = TRUE,
    row_title_rot = 90,
    row_title_side = "left",
    row_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),
    row_title = "Sugar Transport",

    top_annotation = colHA,
    right_annotation = rowHA,
    column_split = colData(cpm_sesub)[["cellType"]],
    column_title = c("ST", "EP", "CH", "PA", "VOS", "VIS", "PH", "PR"),
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),

    heatmap_legend_param = list(
      title = switch(assay_use, raw = "CPM", scaled = "Z-score"),
      direction = "horizontal",
      title_position = "topleft",
      title_gp = grid::gpar(fontsize = 20,
                            fontfamily = "Arial"),
      labels_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),
      legend_width = unit(8, "cm"))
  )

  # Wrap drawing function
  function(...) {
    suppressWarnings(
      ComplexHeatmap::draw(p,
        merge_legend = TRUE,
        heatmap_legend_side = "top", 
        annotation_legend_side = "top",
        legend_gap = grid::unit(1.5, "cm"),
        align_heatmap_legend = "heatmap_left",
        align_annotation_legend = "heatmap_left",
        ...
      )
    )
  }
})

p2 <- pFun(show_heatmap_legend = FALSE)
```

```{r include=FALSE, eval=FALSE}
plotpowerpoint(p, "../data/Figure.pptx", fun = \() pFun(show_heatmap_legend = FALSE),
  location = grid::unit(c(27.22, 4.51, 24.28, 6.07), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

### Heatmap of Benzoxazinoid

```{r}
df_selected <- readxl::read_xlsx("../data/immunity/Module11-Benzoxa.xlsx") |>
  dplyr::rename(gene = Gene)
```

```{r fig.height=4, fig.width=12}
pFun <- local({
  df_genes <- df_selected |>
    dplyr::select(gene, RankName, Label) |>
    dplyr::distinct() |>
    dplyr::left_join(df_rowanno, by = "gene")

  assay_use = "scaled"
  # Prepare expression matrix
  cpm_sesub <- cpm_se[df_genes$gene, ]
  rowData(cpm_sesub) <- df_genes
  mtx <- SummarizedExperiment::assay(cpm_sesub, assay_use)
  colHA <- makeColumnAnnotation(colData(cpm_sesub))
  rowHA <- makeRowAnnotation(rowData(cpm_sesub))

  # Create heatmap
  p <- ComplexHeatmap::Heatmap(
    matrix = mtx,
    col = col_fun,
    use_raster = TRUE,
    border = TRUE,
    na_col = "grey",

    row_labels = rowData(cpm_sesub)[["Label"]],
    show_row_names = TRUE,
    row_names_gp = grid::gpar(fontsize = 7),
    cluster_row_slices = FALSE,
    cluster_row = TRUE,
    row_title_rot = 90,
    row_title_side = "left",
    row_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),
    row_title = "Benzoxazinoid",

    top_annotation = colHA,
    right_annotation = rowHA,
    column_split = colData(cpm_sesub)[["cellType"]],
    column_title = c("ST", "EP", "CH", "PA", "VOS", "VIS", "PH", "PR"),
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),

    heatmap_legend_param = list(
      title = switch(assay_use, raw = "CPM", scaled = "Z-score"),
      direction = "horizontal",
      title_position = "topleft",
      title_gp = grid::gpar(fontsize = 20,
                            fontfamily = "Arial"),
      labels_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),
      legend_width = unit(8, "cm"))
  )

  # Wrap drawing function
  function(...) {
    suppressWarnings(
      ComplexHeatmap::draw(p,
        merge_legend = TRUE,
        heatmap_legend_side = "top", 
        annotation_legend_side = "top",
        legend_gap = grid::unit(1.5, "cm"),
        align_heatmap_legend = "heatmap_left",
        align_annotation_legend = "heatmap_left",
        ...
      )
    )
  }
})

p3 <- pFun(show_heatmap_legend = FALSE)
```

```{r eval=FALSE, include=FALSE}
plotpowerpoint(p, "../data/Figure.pptx", fun = \() pFun(show_heatmap_legend = FALSE),
  location = grid::unit(c(27.22, 4.51, 24.28, 10), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

### Heatmap of Phytohormone

```{r}
df_selected <- openxlsx::getSheetNames("../data/immunity/Module12-Hormone.xlsx") |>
  lapply(function(x)
    readxl::read_xlsx("../data/immunity/Module12-Hormone.xlsx", x) |>
      dplyr::mutate(Module = x)) |>
  dplyr::bind_rows() |>
  dplyr::rename(gene = Gene)
```

```{r fig.height=7, fig.width=12}
pFun <- local({
  df_genes <- df_selected |>
    dplyr::select(gene, RankName, Module, Label) |>
    dplyr::distinct() |>
    dplyr::left_join(df_rowanno, by = "gene")

  assay_use = "scaled"
  # Prepare expression matrix
  cpm_sesub <- cpm_se[df_genes$gene, ]
  rowData(cpm_sesub) <- df_genes
  mtx <- SummarizedExperiment::assay(cpm_sesub, assay_use)
  colHA <- makeColumnAnnotation(colData(cpm_sesub))
  rowHA <- makeRowAnnotation(rowData(cpm_sesub))

  # Create heatmap
  p <- ComplexHeatmap::Heatmap(
    matrix = mtx,
    col = col_fun,
    use_raster = TRUE,
    border = TRUE,
    na_col = "grey",

    row_labels = rowData(cpm_sesub)[["Label"]],
    show_row_names = TRUE,
    row_names_gp = grid::gpar(fontsize = 7),
    cluster_row_slices = FALSE,
    cluster_row = TRUE,
    row_title_rot = 90,
    row_title_side = "left",
    row_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),
    row_split = rowData(cpm_sesub)[["Module"]],

    top_annotation = colHA,
    right_annotation = rowHA,
    column_split = colData(cpm_sesub)[["cellType"]],
    column_title = c("ST", "EP", "CH", "PA", "VOS", "VIS", "PH", "PR"),
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),

    heatmap_legend_param = list(
      title = switch(assay_use, raw = "CPM", scaled = "Z-score"),
      direction = "horizontal",
      title_position = "topleft",
      title_gp = grid::gpar(fontsize = 20,
                            fontfamily = "Arial"),
      labels_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),
      legend_width = unit(8, "cm"))
  )

  # Wrap drawing function
  function(...) {
    suppressWarnings(
      ComplexHeatmap::draw(p,
        merge_legend = TRUE,
        heatmap_legend_side = "top", 
        annotation_legend_side = "top",
        legend_gap = grid::unit(1.5, "cm"),
        align_heatmap_legend = "heatmap_left",
        align_annotation_legend = "heatmap_left",
        ...
      )
    )
  }
})

p4 <- pFun(show_heatmap_legend = FALSE)
```

```{r  eval=FALSE, include=FALSE}
plotpowerpoint(p, "../data/Figure.pptx", fun = \() pFun(show_heatmap_legend = FALSE),
  location = grid::unit(c(27.22, 4.51, 24.28, 14.2), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

### Heatmap of Redox

```{r}
df_selected <- openxlsx::getSheetNames("../data/immunity/Module13-Redox.xlsx") |>
  lapply(function(x)
    readxl::read_xlsx("../data/immunity/Module13-Redox.xlsx", x) |>
      dplyr::mutate(Module = x)) |>
  dplyr::bind_rows() |>
  dplyr::rename(gene = Gene)
```

```{r fig.height=7, fig.width=12}
pFun <- local({
  df_genes <- df_selected |>
    dplyr::filter(Module %in% c("Redox", "Polyamine")) |>
    dplyr::select(gene, RankName, Module, Label) |>
    dplyr::distinct() |>
    dplyr::left_join(df_rowanno, by = "gene")

  assay_use = "scaled"
  # Prepare expression matrix
  cpm_sesub <- cpm_se[df_genes$gene, ]
  rowData(cpm_sesub) <- df_genes
  mtx <- SummarizedExperiment::assay(cpm_sesub, assay_use)
  colHA <- makeColumnAnnotation(colData(cpm_sesub))
  rowHA <- makeRowAnnotation(rowData(cpm_sesub))

  # Create heatmap
  p <- ComplexHeatmap::Heatmap(
    matrix = mtx,
    col = col_fun,
    use_raster = TRUE,
    border = TRUE,
    na_col = "grey",

    row_labels = rowData(cpm_sesub)[["Label"]],
    show_row_names = TRUE,
    row_names_gp = grid::gpar(fontsize = 7),
    cluster_row_slices = FALSE,
    cluster_row = TRUE,
    row_title_rot = 90,
    row_title_side = "left",
    row_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),
    row_split = rowData(cpm_sesub)[["Module"]],

    top_annotation = colHA,
    right_annotation = rowHA,
    column_split = colData(cpm_sesub)[["cellType"]],
    column_title = c("ST", "EP", "CH", "PA", "VOS", "VIS", "PH", "PR"),
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),

    heatmap_legend_param = list(
      title = switch(assay_use, raw = "CPM", scaled = "Z-score"),
      direction = "horizontal",
      title_position = "topleft",
      title_gp = grid::gpar(fontsize = 20,
                            fontfamily = "Arial"),
      labels_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),
      legend_width = unit(8, "cm"))
  )

  # Wrap drawing function
  function(...) {
    suppressWarnings(
      ComplexHeatmap::draw(p,
        merge_legend = TRUE,
        heatmap_legend_side = "top", 
        annotation_legend_side = "top",
        legend_gap = grid::unit(1.5, "cm"),
        align_heatmap_legend = "heatmap_left",
        align_annotation_legend = "heatmap_left",
        ...
      )
    )
  }
})

p5 <- pFun(show_heatmap_legend = FALSE)
```

```{r eval=FALSE, include=FALSE}
plotpowerpoint(p, "../data/Figure.pptx", fun = \() pFun(show_heatmap_legend = FALSE),
  location = grid::unit(c(27.22, 4.51, 24.28, 14.2), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

```{r fig.height=4, fig.width=12}
pFun <- local({
  df_genes <- df_selected |>
    dplyr::filter(Module %in% c("Peroxidase", "GST")) |>
    dplyr::select(gene, RankName, Module, Label) |>
    dplyr::distinct() |>
    dplyr::left_join(df_rowanno, by = "gene")

  assay_use = "scaled"
  # Prepare expression matrix
  cpm_sesub <- cpm_se[df_genes$gene, ]
  rowData(cpm_sesub) <- df_genes
  mtx <- SummarizedExperiment::assay(cpm_sesub, assay_use)
  colHA <- makeColumnAnnotation(colData(cpm_sesub))
  rowHA <- makeRowAnnotation(rowData(cpm_sesub))

  # Create heatmap
  p <- ComplexHeatmap::Heatmap(
    matrix = mtx,
    col = col_fun,
    use_raster = TRUE,
    border = TRUE,
    na_col = "grey",

    row_labels = rowData(cpm_sesub)[["Label"]],
    show_row_names = TRUE,
    row_names_gp = grid::gpar(fontsize = 7),
    cluster_row_slices = FALSE,
    cluster_row = TRUE,
    row_title_rot = 90,
    row_title_side = "left",
    row_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),
    row_split = rowData(cpm_sesub)[["Module"]],

    top_annotation = colHA,
    right_annotation = rowHA,
    column_split = colData(cpm_sesub)[["cellType"]],
    column_title = c("ST", "EP", "CH", "PA", "VOS", "VIS", "PH", "PR"),
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),

    heatmap_legend_param = list(
      title = switch(assay_use, raw = "CPM", scaled = "Z-score"),
      direction = "horizontal",
      title_position = "topleft",
      title_gp = grid::gpar(fontsize = 20,
                            fontfamily = "Arial"),
      labels_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),
      legend_width = unit(8, "cm"))
  )

  # Wrap drawing function
  function(...) {
    suppressWarnings(
      ComplexHeatmap::draw(p,
        merge_legend = TRUE,
        heatmap_legend_side = "top", 
        annotation_legend_side = "top",
        legend_gap = grid::unit(1.5, "cm"),
        align_heatmap_legend = "heatmap_left",
        align_annotation_legend = "heatmap_left",
        ...
      )
    )
  }
})

p6 <- pFun(show_heatmap_legend = TRUE)
```

```{r eval=FALSE, include=FALSE}
plotpowerpoint(p, "../data/Figure.pptx", fun = \() pFun(show_heatmap_legend = FALSE),
  location = grid::unit(c(27.22, 4.51, 24.28, 8), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

### Heatmap of Trascription Factors

```{r}
df_selected <- readxl::read_xlsx("../data/immunity/Module14-TF.xlsx") |>
  dplyr::rename(gene = Gene)
```

```{r fig.height=3, fig.width=12}
pFun <- local({
  df_genes <- df_selected |>
    dplyr::select(gene, RankName, Category, Label) |>
    dplyr::distinct() |>
    dplyr::left_join(df_rowanno, by = "gene")

  assay_use = "scaled"
  # Prepare expression matrix
  cpm_sesub <- cpm_se[df_genes$gene, ]
  rowData(cpm_sesub) <- df_genes
  mtx <- SummarizedExperiment::assay(cpm_sesub, assay_use)
  colHA <- makeColumnAnnotation(colData(cpm_sesub))
  rowHA <- makeRowAnnotation(rowData(cpm_sesub))

  # Create heatmap
  p <- ComplexHeatmap::Heatmap(
    matrix = mtx,
    col = col_fun,
    use_raster = TRUE,
    border = TRUE,
    na_col = "grey",

    row_labels = rowData(cpm_sesub)[["Label"]],
    show_row_names = TRUE,
    row_names_gp = grid::gpar(fontsize = 7),
    cluster_row_slices = FALSE,
    cluster_row = TRUE,
    row_title_rot = 90,
    row_title_side = "left",
    row_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),
    row_split = rowData(cpm_sesub)[["Category"]],

    top_annotation = colHA,
    right_annotation = rowHA,
    column_split = colData(cpm_sesub)[["cellType"]],
    column_title = c("ST", "EP", "CH", "PA", "VOS", "VIS", "PH", "PR"),
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),

    heatmap_legend_param = list(
      title = switch(assay_use, raw = "CPM", scaled = "Z-score"),
      direction = "horizontal",
      title_position = "topleft",
      title_gp = grid::gpar(fontsize = 20,
                            fontfamily = "Arial"),
      labels_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),
      legend_width = unit(8, "cm"))
  )

  # Wrap drawing function
  function(...) {
    suppressWarnings(
      ComplexHeatmap::draw(p,
        merge_legend = TRUE,
        heatmap_legend_side = "top", 
        annotation_legend_side = "top",
        legend_gap = grid::unit(1.5, "cm"),
        align_heatmap_legend = "heatmap_left",
        align_annotation_legend = "heatmap_left",
        ...
      )
    )
  }
})

p7 <- pFun(show_heatmap_legend = FALSE)
```

```{r eval=FALSE, include=FALSE}
plotpowerpoint(p, "../data/Figure.pptx", fun = \() pFun(show_heatmap_legend = FALSE),
  location = grid::unit(c(27.22, 4.51, 24.28, 6.07), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

## Figure 5 Specific Pathways

### Heatmap of Splited Category

```{r}
df_selected <- openxlsx::getSheetNames("../data/immunity/Module01-09.xlsx") |>
  lapply(function(x)
    readxl::read_xlsx("../data/immunity/Module01-09.xlsx", x) |>
      dplyr::mutate(Module = x, Weights = as.numeric(factor(Category)))) |>
  dplyr::bind_rows() |>
  dplyr::rename(gene = Gene)
```

```{r fig.height=20, fig.width=12}
genHeatFun <- function(df, module, width = NULL, show_heads = TRUE, show_row_names = TRUE, ...) {
  df_genes <- df |>
    dplyr::filter(Module == module) |>
    dplyr::select(gene, RankName, Module, Category, Label, Weights) |>
    dplyr::distinct() |>
    dplyr::left_join(df_rowanno, by = "gene")

  assay_use = "scaled"
  # Prepare expression matrix
  cpm_sesub <- cpm_se[df_genes$gene, ]
  rowData(cpm_sesub) <- df_genes
  mtx <- SummarizedExperiment::assay(cpm_sesub, assay_use)
  colHA <- makeColumnAnnotation(colData(cpm_sesub), show_annotation_name = FALSE)
  rowHA <- makeRowAnnotation(rowData(cpm_sesub))

  # Create heatmap
  p <- ComplexHeatmap::Heatmap(
    matrix = mtx,
    col = col_fun,
    use_raster = TRUE,
    border = TRUE,
    na_col = "grey",

    width = width,
    
    row_labels = rowData(cpm_sesub)[["Label"]],
    show_row_names = show_row_names,
    row_names_gp = grid::gpar(fontsize = 7),
    cluster_row_slices = TRUE,
    cluster_row = TRUE,
    clustering_method_rows = "average",
    row_title = module,
    row_title_rot = 90,
    row_title_side = "left",
    row_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),
    row_split = rowData(cpm_sesub)[["Category"]],

    top_annotation = if (show_heads) colHA else NULL,
    right_annotation = rowHA,
    column_split = colData(cpm_sesub)[["cellType"]],
    column_title = if (show_heads)
      c("ST", "EP", "CH", "PA", "VOS", "VIS", "PH", "PR") else NULL,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),

    heatmap_legend_param = list(
      title = switch(assay_use, raw = "CPM", scaled = "Z-score"),
      direction = "horizontal",
      title_position = "topleft",
      title_gp = grid::gpar(fontsize = 20,
                            fontfamily = "Arial"),
      labels_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),
      legend_width = unit(8, "cm")),
    ...
  )

  # Wrap drawing function
  function(...) {
    suppressWarnings(
      ComplexHeatmap::draw(p,
        merge_legend = TRUE,
        heatmap_legend_side = "top", 
        annotation_legend_side = "top",
        legend_gap = grid::unit(1.5, "cm"),
        align_heatmap_legend = "heatmap_left",
        align_annotation_legend = "heatmap_left",
        show_heatmap_legend = FALSE,
        ...
      )
    )
  }
}
```

ComplexHeatmap Rasterize 阐述的图片会以 slice 命名，这就导致绘制在同一个 PPT Slides 上的 heatmap 会出错。所以必须 `use_raster=FALSE` 才能正常绘图。或者将 heatmap 绘制到不同 slides 上，然后手动合并。

```{r eval=FALSE, include=FALSE}
local({
  modules <- unique(df_selected$Module)

  plotFuns <- mapply(
    modules, seq_along(modules),
    FUN = function(module, index) {
      genHeatFun(df_selected, module, show_heads = index == 1,
                 #width = grid::unit(0.9, "npc"),
                 row_names_max_width = unit(0.005, "npc"),
                 row_dend_width = unit(0.05, "npc"),
                 show_row_names = TRUE)
    }
  )

  heights <- mapply(
    modules, seq_along(modules),
    FUN = function(module, index) {
      b <- nrow(dplyr::filter(df_selected, Module == module)) * 0.22
      if (index == 1) b <- b + 1.95
      b
  })

  GAP <- 0.1
  TOP <- 4.79
  tops <- cumsum(heights + GAP) - heights + TOP - GAP

  target.file <- paste0(tempfile(), ".pptx")
  file.copy("../data/Figure.pptx", target.file, overwrite = TRUE)

  doc <- officer::read_pptx(path = target.file)

  for (i in seq_along(modules)) {
    location <- grid::unit(c(2.8, tops[[i]], 22, heights[[i]]), "cm")
    location <- grid::convertUnit(location, "inches")
    phloc <- officer::ph_location(
      left = location[[1]], top = location[[2]],
      width = location[[3]], height = location[[4]]
    )

    doc <- officer::add_slide(doc, layout = "Blank", master = "Office Theme")

    doc <- officer::ph_with(
      x = doc,
      value = rvg::dml(
        code = plotFuns[[i]](),
        fonts = list(sans = "Arial", serif = "Arial", mono = "Courier")),
      location = phloc)
  }

  print(doc, target = target.file)

  utils::browseURL(target.file)
})
```

## Supplementary Figure 6

```{r}
TaWRKY <- readxl::read_xlsx("../data/immunity/TaWRKY.xlsx")
```

```{r fig.height=10, fig.width=12}
pFun <- local({
  df_genes <- TaWRKY |>
    dplyr::left_join(df_rowanno, by = c("Gene" = "gene")) |>
    dplyr::filter(!is.na(hits))

  assay_use = "scaled"
  # Prepare expression matrix
  cpm_sesub <- cpm_se[df_genes$Gene, ]
  rowData(cpm_sesub) <- df_genes
  mtx <- SummarizedExperiment::assay(cpm_sesub, assay_use)
  colHA <- makeColumnAnnotation(colData(cpm_sesub))
  rowHA <- makeRowAnnotation(rowData(cpm_sesub))

  # Create heatmap
  p <- ComplexHeatmap::Heatmap(
    matrix = mtx,
    col = col_fun,
    use_raster = TRUE,
    border = TRUE,
    na_col = "grey",

    row_labels = rowData(cpm_sesub)[["Name"]],
    show_row_names = TRUE,
    row_names_gp = grid::gpar(fontsize = 16, fontface = "italic"),
    cluster_row_slices = FALSE,
    cluster_row = TRUE,
    row_title_rot = 90,
    row_title_side = "left",
    row_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),

    top_annotation = colHA,
    right_annotation = rowHA,
    column_split = colData(cpm_sesub)[["cellType"]],
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_title_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),

    heatmap_legend_param = list(
      title = switch(assay_use, raw = "CPM", scaled = "Z-score"),
      direction = "horizontal",
      title_position = "topleft",
      title_gp = grid::gpar(fontsize = 20,
                            fontfamily = "Arial"),
      labels_gp = grid::gpar(fontsize = 16, fontfamily = "Arial"),
      legend_width = unit(8, "cm"))
  )

  # Wrap drawing function
  function(...) {
    suppressWarnings(
      ComplexHeatmap::draw(p,
        merge_legend = TRUE,
        heatmap_legend_side = "top", 
        annotation_legend_side = "top",
        legend_gap = grid::unit(1.5, "cm"),
        align_heatmap_legend = "heatmap_left",
        align_annotation_legend = "heatmap_left",
        ...
      )
    )
  }
})

p7 <- pFun(show_heatmap_legend = TRUE)
```

```{r eval=FALSE, include=FALSE}
plotpowerpoint(p, "../data/Figure.pptx", fun = \() pFun(show_heatmap_legend = TRUE),
  location = grid::unit(c(2.5, 4.6, 49.01, 40.84), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

## Supplementary Table 4

### Generate a Curated Gene List

```{r}
df_deg <- ds_sig |>
  rhapsodykit::diff_state_format() |>
  dplyr::mutate(
    time = stringr::str_extract(contrast, "\\dDPI"),
    comparison = stringr::str_replace(contrast,
      "X(\\dDPI)\\.(.*)-X(\\dDPI)\\.(.*)", "\\2 vs. \\4")) |>
  dplyr::filter(comparison != "PNR2 vs. TR4")
```

```{r eval=FALSE}
ranknames <- Sys.glob("../data/heatmap/*.xlsx") |>
  lapply(function(xlsx_file) {
    category <- stringr::str_extract(
      basename(xlsx_file), ".*(?=-for)")
    lapply(
      openxlsx::getSheetNames(xlsx_file),
      FUN = function(sheet) {
        readxl::read_excel(xlsx_file, sheet) |>
          dplyr::select(Gene, RankName)
      }) |>
      dplyr::bind_rows() |>
      dplyr::distinct() |>
      dplyr::mutate(category = category)
  }) |>
  dplyr::bind_rows() |>
  dplyr::distinct() |>
  dplyr::filter(category != "FAup") |>
  dplyr::group_by(Gene) |>
  dplyr::summarise(RankName = paste(unique(RankName), collapse = " | "),
                   Category = sample(category, 1))

ranknames
```

```{r}
df_annolist <- dplyr::left_join(df_deg, ranknames, by = c("gene" = "Gene")) |>
  dplyr::filter(!is.na(RankName)) |>
  dplyr::select(gene, RankName, Category, time, comparison, cluster_id, baseMean, logFC)
```

```{r}
dplyr::sample_n(df_annolist, 20)
```

```{r eval=FALSE}
openxlsx::write.xlsx(df_annolist, "../results/DataPlots/SupplementaryTable_4_Curated.xlsx")
```

### Query Information

```{r}
df_annolist |>
  dplyr::filter(Category == "CaSignaling")
```

