---
title: "Figure 3 Cell Annotation"
author: "Altair Wei"
date: "2023-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(ggtree, include.only = "%<+%")
source("../scripts/LoadUtils.R", chdir = TRUE)

options(ggrastr.default.dpi=300)
PT = 1/2.13
FT = 6/17.1
```

## Load Data Objects

### Seurat Objects

#### MOCK samples

```{r}
obj_mock <- readRDS(Sys.glob("../results/ObjectCache/IntegrationAndAnnotation/obj_mock_annotated_*.rds"))
sce_mock <- Seurat::as.SingleCellExperiment(obj_mock, assay = "RNA")
sce_mock <- muscat::prepSCE(sce_mock,
  kid = "ident", # subpopulation assignments
  gid = "group", # group IDs (ctrl/stim)
  sid = "sample", # sample IDs (ctrl/stim.1234)
  drop = FALSE
)
```

```{r}
mock_celltypes <- c(
  "Gu", "Ep_α", "Ep_β",
  "Me_α", "Me_β", "Me_ε", "Me_γ", "Va_δ", "Me_δ",
  "Va_β", "BS", "CC", "MPV", "Va_α", "Va_γ")

mock_celltype_colors <- ggthemes::tableau_color_pal("Tableau 20")(length(mock_celltypes))
names(mock_celltype_colors) <- mock_celltypes
```

```{r}
mock_body_layers <- c(
  "Gu"   = "L1",
  "Ep_α" = "L1",
  "Ep_β" = "L1",
  "Me_α" = "L2",
  "Me_β" = "L2",
  "Me_γ" = "L2",
  "Me_δ" = "L2",
  "Me_ε" = "L2",
  "BS"   = "L3",
  "CC"   = "L3",
  "Va_α" = "L3",
  "Va_β" = "L3",
  "Va_γ" = "L3",
  "Va_δ" = "L3",
  "MPV"  = "L3"
)


mock_tissues <- c(
  Gu   = "Stomata",
  Ep_α = "Epidermis",
  Ep_β = "Epidermis",
  Me_α = "Chlorenchyma",
  Me_β = "Chlorenchyma",
  Me_ε = "Chlorenchyma",
  Me_γ = "Parenchyma",
  Me_δ = "Undefined mesophyll",
  Va_δ = "Undefined mesophyll",
  Va_β = "Outer sheath",
  BS   = "Inner sheath",
  CC   = "Phloem",
  MPV  = "Procambium",
  Va_α = "Undefined vascular cells",
  Va_γ = "Undefined vascular cells"
)

mock_tissues_colors <- structure(
  names = c("Stomata",      "Epidermis",    "Chlorenchyma", "Parenchyma", "Undefined mesophyll",
            "Outer sheath", "Inner sheath", "Phloem",       "Procambium", "Undefined vascular cells"),
  .Data = c("#4e79a7",      "#f28e2b",      "#8cd17d",      "#b6992d",    "#a6a6a6",
            "#86bcb6",      "#e15759",      "#ff9d9a",      "#79706e",    "#a6a6a6")
)
```

```{r}
sce_mock$cluster_id <- forcats::fct_relevel(sce_mock$cluster_id, mock_celltypes)
sce_mock$body_layer <- do.call(
  dplyr::recode, c(list(.x = sce_mock$cluster_id), as.list(mock_body_layers)))
sce_mock$tissue <- dplyr::recode(sce_mock$cluster_id, !!!mock_tissues)
Seurat::Idents(obj_mock) <- forcats::fct_relevel(Seurat::Idents(obj_mock), mock_celltypes)
```

#### All samples

```{r}
obj <- readRDS(Sys.glob("../results/ObjectCache/IntegrationAndAnnotation/obj_annotated_*.rds"))
sce <- Seurat::as.SingleCellExperiment(obj, assay = "RNA")
sce <- muscat::prepSCE(sce,
  kid = "ident", # subpopulation assignments
  gid = "group", # group IDs (ctrl/stim)
  sid = "sample", # sample IDs (ctrl/stim.1234)
  drop = FALSE
)
```

```{r}
all_celltypes <- c(
   "Gu", "Ep_1", "Ep_2",
  "Me_1", "Me_2", "Me_3", "Me_4", "Me_5", "Me_6",
  "Va_1", "Va_2", "Va_3", "Va_4",
  "BS", "CC", "MPV_1", "MPV_2")

all_celltype_colors <- ggthemes::tableau_color_pal("Tableau 20")(length(all_celltypes))
#mock_celltype_colors <- scater:::.get_palette("tableau20")[seq_along(mock_celltypes)]
names(all_celltype_colors) <- all_celltypes
scales::show_col(all_celltype_colors)
```

```{r}
all_body_layers <- c(
  "Gu"    = "L1",
  "Ep_1"  = "L1",
  "Ep_2"  = "L1",
  "Me_1"  = "L2",
  "Me_2"  = "L2",
  "Me_3"  = "L2",
  "Me_4"  = "L2",
  "Me_5"  = "L2",
  "Me_6"  = "L2",
  "Va_1"  = "L2",
  "Va_2"  = "L2",
  "Va_3"  = "L2",
  "Va_4"  = "L2",
  "BS"    = "L3",
  "CC"    = "L3",
  "MPV_1" = "L3",
  "MPV_2" = "L3"
)

all_tissues <- c(
  Gu   = "Stomata",
  Ep_1 = "Epidermis",
  Ep_2 = "Epidermis",
  Me_1 = "Chlorenchyma",
  Me_2 = "Chlorenchyma",
  Me_3 = "Chlorenchyma",
  Me_4 = "Parenchyma",
  Me_5 = "Parenchyma",
  Me_6 = "Parenchyma",
  Va_1 = "Chlorenchyma",
  Va_2 = "Outer sheath",
  Va_3 = "Outer sheath",
  Va_4 = "Outer sheath",
  BS = "Inner sheath",
  CC = "Phloem",
  MPV_1 = "Procambium",
  MPV_2 = "Procambium"
)

all_tissues_colors <- structure(
  names = c("Stomata",      "Epidermis",    "Chlorenchyma", "Parenchyma",
            "Outer sheath", "Inner sheath", "Phloem",       "Procambium"),
  .Data = c("#4e79a7",      "#f28e2b",      "#8cd17d",      "#b6992d",
            "#86bcb6",      "#e15759",      "#ff9d9a",      "#79706e")
)


obj$tissue <- do.call(
  dplyr::recode, c(list(.x = Seurat::Idents(obj)), as.list(all_tissues)))
```

```{r}
sce$cluster_id <- forcats::fct_relevel(sce$cluster_id, all_celltypes)
Seurat::Idents(obj) <- forcats::fct_relevel(Seurat::Idents(obj), all_celltypes)
sce$body_layer <- do.call(
  dplyr::recode, c(list(.x = sce$cluster_id), as.list(all_body_layers)))
sce$tissue <- do.call(
  dplyr::recode, c(list(.x = sce$cluster_id), as.list(all_tissues)))
```

## Figure 3 Cell Annotation

### Dim Reduction

```{r fig.height=8, fig.width=10}
df_to_plot <- Seurat::FetchData(obj_mock, vars = c("UMAP_1", "UMAP_2", "ident", "RNA_snn_res.0.6"))

df_to_label <- df_to_plot |>
  dplyr::group_by(ident) |>
  dplyr::summarise(
    UMAP_1 = median(UMAP_1),
    UMAP_2 = median(UMAP_2),
    text = unique(RNA_snn_res.0.6))

p1 <- ggplot2::ggplot(df_to_plot, ggplot2::aes(x = UMAP_1, y = UMAP_2, color = ident)) +
  ggplot2::geom_point(size = 0.2) +
  ggplot2::scale_color_manual(values = mock_celltype_colors) +
  ggplot2::geom_point(
    data = df_to_label,
    mapping = ggplot2::aes(x = UMAP_1, y = UMAP_2),
    shape = 21, color = "black", size = 11, fill = "white", alpha = 0.5) +
  ggplot2::geom_text(
    data = df_to_label,
    mapping = ggplot2::aes(x = UMAP_1, y = UMAP_2, label = text),
    color = "black", size = 16 * FT, hjust = 0.5, vjust = 0.5, nudge_y = -0.05) +
  ggplot2::guides(
    color = ggplot2::guide_legend(
      override.aes = list(size = 4),
      title = "cluster")) +
  ggplot2::coord_fixed() +
  cowplot::theme_cowplot(font_size = 22, font_family = "Arial") +
  theme_dimred(linesize = 1.5 * PT) +
  remove_legend()


p1 <- ggrastr::rasterise(p1)

p1
```

```{r eval=FALSE, include=FALSE}
plotpowerpoint(p1, "../data/Figure.pptx",
  location = grid::unit(c(2.5, 4.49,  25.68, 20.32), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

### Known Marker Expression

```{r fig.height=10, fig.width=7.5}
cluster_num <- data.frame(
    names = as.character(Seurat::Idents(obj_mock)),
    values = as.character(obj_mock$RNA_snn_res.0.6)) |>
  dplyr::distinct() |>
  tibble::deframe()

p2a <- Seurat::DotPlot(obj_mock, features = KNOWN_MARKERS) +
  ggplot2::coord_flip() +
  ggplot2::facet_grid(
    facets = feature.groups ~ .,
    scales = "free_y",
    space = "free_y",
    switch = "y") +
  ggplot2::ylab("cluster") +
  ggplot2::scale_y_discrete(labels = \(x) dplyr::recode(x, !!!cluster_num)) +
  ggplot2::guides(
    size = ggplot2::guide_legend(title.position = "left"),
    color = ggplot2::guide_colorbar(title.position = "left")) +
  cowplot::theme_cowplot(font_size = 22, font_family = "Arial") +
  ggplot2::theme(
    axis.title.y = ggplot2::element_blank(),
    axis.title.x = ggplot2::element_blank(),
    axis.text.y = ggplot2::element_blank(),
    axis.text.x = ggplot2::element_blank(),
    axis.ticks = ggplot2::element_line(linewidth = 1.5/2.13),
    axis.line = ggplot2::element_line(linewidth = 1.5/2.13),
    strip.background = ggplot2::element_blank(),
    strip.text.y.left = ggplot2::element_text(
      size = 18, angle = 0, face = "italic", hjust = 1),
    strip.placement = "outside",
    #panel.spacing.y = grid::unit(0.5/100, "npc"),
    panel.grid = ggplot2::element_blank(),
    legend.title = ggplot2::element_text(angle = 90, hjust = 0.5),
    plot.margin = ggplot2::margin(t = 9, r = 9, b = 0, l = 9, unit = "pt")
  ) + remove_legend()

p2b <- ggplot2::ggplot(
    data = data.frame(
      cluster = factor(mock_celltypes, level = mock_celltypes),
      label = dplyr::recode(mock_celltypes, !!!cluster_num)),
    mapping = ggplot2::aes(x = cluster, y = 1, label = label, fill = cluster)) +
  ggplot2::geom_point(shape = 21, size = 11) +
  ggplot2::scale_fill_manual(values = mock_celltype_colors) +
  ggplot2::geom_text(size = 16 * FT, hjust = 0.5, vjust = 0.6) +
  ggplot2::theme_void(base_size = 18) +
  ggplot2::theme(legend.position = "none",
                 plot.margin = ggplot2::margin(t = 0, r = 9, b = 9, l = 9, unit = "pt"))

(p2 <- p2a / p2b + patchwork::plot_layout(heights = c(18, 1)))
```

```{r eval=FALSE, include=FALSE}
plotpowerpoint(p2, "../data/Figure.pptx",
  location = grid::unit(c(30.87, 4.49, 20.63, 26.15), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

### Deconvolution of LCM-seq

```{r}
CIBER_raw <- readr::read_tsv("../results/Deconv/CIBERSORTx_Adjusted.txt")
CIBER_frac <- dplyr::select(CIBER_raw, Sample = Mixture, !c(`P-value`, Correlation, RMSE)) |>
  tidyr::pivot_longer(!Sample, names_to = "CellType", values_to = "Fraction")
```

```{r}
LCM_design <- c(
  "VC_1" = "vb.",
  "VC_2" = "vb.",
  "VC_3" = "vb.",
  "CC_1" = "bp.",
  "CC_2" = "bp.",
  "CC_3" = "bp.",
  "MC3_1" = "dp.",
  "MC3_2" = "dp.",
  "MC3_3" = "dp.",
  "MC_1" = "me.",
  "MC_2" = "me.",
  "MC_3" = "me.",
  "EC_1" = "ep.",
  "EC_2" = "ep.",
  "EC_3" = "ep."
)

CIBER_frac$SampleClass <- dplyr::recode(CIBER_frac$Sample, !!!LCM_design)
CIBER_frac$CellType <- dplyr::recode(CIBER_frac$CellType, !!!cluster_num)
CIBER_frac$CellType <- factor(CIBER_frac$CellType, levels = cluster_num[mock_celltypes])

# Put stomata after epidermis
CIBER_frac$CellType <- forcats::fct_relevel(CIBER_frac$CellType, "11", "13", "14")
```

```{r fig.height=9.2, fig.width=7}
mock_cluster_colors <- setNames(
  mock_celltype_colors, cluster_num[names(mock_celltype_colors)])

p3 <- ggplot2::ggplot(
      data = CIBER_frac,
      mapping = ggplot2::aes(
        x = factor(SampleClass, unique(LCM_design)),
        y= Fraction, color = as.character(CellType))
    ) +
    ggplot2::geom_point(stat = "summary", fun = mean, size = 4) +
    ggplot2::geom_errorbar(
      stat="summary",
      fun.min = function(x) mean(x)+sd(x)/sqrt(length(x)),
      fun.max = function(x) mean(x)-sd(x)/sqrt(length(x)),
      width = 0.2, linewidth = 1.5) +
    ggplot2::scale_y_continuous(labels = scales::percent) +
    ggplot2::facet_wrap(~ CellType, scales = "free_y", ncol = 3,
      labeller = ggplot2::labeller(CellType = \(x) paste("cluster:", x))) +
    ggplot2::guides(color = "none") +
    ggplot2::ylab("CIBERSORTx estimate (a.u.)") +
    ggplot2::xlab("LCM Samples") +
    ggplot2::scale_color_manual(values = mock_cluster_colors) +
    ggplot2::theme_bw(base_size = 22) +
    ggplot2::theme(
      panel.grid = ggplot2::element_blank(),
      strip.text = ggplot2::element_text(
        size = 18, vjust = 0.2, margin = ggplot2::margin(t = 4, b = 4)),
      axis.text.x = ggplot2::element_text(
        size = 16, angle = 0, hjust = 0.5, vjust = 0.5, color = "black"),
      axis.text.y = ggplot2::element_blank(),
      axis.ticks = ggplot2::element_line(linewidth = 1.5 * PT),
      strip.background = ggplot2::element_rect(linewidth = 1.5 * PT),
      panel.border = ggplot2::element_rect(linewidth = 1.5 * PT))

p3
```

```{r eval=FALSE, include=FALSE}
plotpowerpoint(p3, "../data/Figure.pptx",
  location = grid::unit(c(15.34, 32.96, 19.24, 23.22), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

```{r}
knitr::knit_exit()
```

## Supplementary Figure 2

### Resolution tree

```{r fig.height=10, fig.width=7}
p1 <- local({
  library(clustree)
  clustree::clustree(
      obj_mock, prefix = "RNA_snn_res.",
      edge_width = 1,
      node_text_size = 16 * FT) +
    ggplot2::scale_size_continuous(range = c(4, 15), breaks = c(500, 1000, 5000)) +
    legend_override("color", list(size = 4), title = "resolution") +
    ggplot2::theme(
      legend.title = ggplot2::element_text(size = 20),
      legend.text = ggplot2::element_text(size = 18)
    )
})

p1
```

```{r eval=FALSE, include=FALSE}
plotpowerpoint(p1, "../data/Figure.pptx",
  location = grid::unit(c(2.5, 4.49, 24, 24.5), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

### Cell Cluster Tree

#### Based on dim reduc

```{r}
obj_mock <- Seurat::BuildClusterTree(obj_mock, dims = 1:20, reduction = "harmony")
data.tree <- Seurat::Tool(obj_mock, slot = "Seurat::BuildClusterTree")
ape::plot.phylo(x = data.tree, direction = "rightwards", underscore = TRUE)
```

#### Based on euclidean

Based euclidean distance of non-log averaged expression of variable features:

```{r}
obj_mock <- Seurat::BuildClusterTree(obj_mock)
data.tree <- Seurat::Tool(obj_mock, slot = "Seurat::BuildClusterTree")
ape::plot.phylo(x = data.tree, direction = "rightwards", underscore = TRUE)
```

#### Based on correlation

Based on dissimilarity of log averaged expression of all features:

```{r}
avg_expr <- log1p(Seurat::AverageExpression(obj_mock, slot = "data")$RNA)
avg_expr_cor <- cor(avg_expr)
cor_dist <- as.dist(sqrt(1 - avg_expr_cor^2))

cor_dist |>
  hclust() |>
  ape::as.phylo() |>
  ape::plot.phylo(direction = "rightwards", underscore = TRUE)
title("All features")
```

Based on dissimilarity of log averaged expression of variable features:

```{r}
avg_expr <- log1p(Seurat::AverageExpression(
  obj_mock, features = Seurat::VariableFeatures(obj_mock), slot = "data")$RNA)
avg_expr_cor <- cor(avg_expr)
cor_dist <- as.dist(sqrt(1 - avg_expr_cor^2))

cor_dist |>
  hclust() |>
  ape::as.phylo() |>
  ape::plot.phylo(direction = "rightwards", underscore = TRUE)
title("Variable features")
```

#### Correlation Heatmap

```{r fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
avg_expr <- log1p(Seurat::AverageExpression(
  obj_mock, features = Seurat::VariableFeatures(obj_mock),
  slot = "data", group.by = "RNA_snn_res.0.6")$RNA)
avg_expr_cor <- cor(avg_expr)

p2 <- ComplexHeatmap::Heatmap(
  matrix = avg_expr_cor,
  col = viridis::viridis(256),
  column_names_rot = 0,
  column_names_centered = TRUE,
  clustering_distance_rows = \(x) as.dist(sqrt(1 - x^2)),
  clustering_distance_columns = \(x) as.dist(sqrt(1 - x^2)),
  row_dend_reorder = FALSE,
  column_dend_reorder = FALSE,
  row_dend_width = grid::unit(30, "mm"),
  column_dend_height = grid::unit(30, "mm"),
  row_names_gp = grid::gpar(fontsize = 20),
  column_names_gp = grid::gpar(fontsize = 20),
  row_dend_gp = grid::gpar(lwd = 2),
  column_dend_gp = grid::gpar(lwd = 2),

  heatmap_legend_param = list(
    title = "Pearson's R",
    title_gp = grid::gpar(fontsize = 20),
    title_position = "topcenter",
    labels_gp = grid::gpar(fontsize = 16),
    legend_direction = "horizontal",
    legend_width = grid::unit(8, "cm"),
    grid_height = grid::unit(0.6, "cm")
  )
)

ComplexHeatmap::draw(p2, heatmap_legend_side = "top")
```

```{r eval=FALSE, include=FALSE}
plotpowerpoint(p2, "../data/Figure.pptx",
               fun = \() {ComplexHeatmap::draw(p2, heatmap_legend_side = "top")},
  location = grid::unit(c(27.13, 4.49, 24.37, 24.5), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

### Marker Genes Expression

```{r fig.height=7, fig.width=7}
p3 <- local({
  scater::plotReducedDim(
      sce_mock, "UMAP",
      colour_by = "TraesCS1B02G406200",
      order_by = "TraesCS1B02G406200",
      point_alpha = 1, point_size = 0.5,
      rasterise = TRUE) +
    ggplot2::scale_color_gradient(low = "grey", high = "blue") +
    ggplot2::coord_fixed() +
    ggplot2::ggtitle("TraesCS1B02G406200") +
    ggplot2::labs(x = "UMAP_1", y = "UMAP_2") +
    ggplot2::guides(color = ggplot2::guide_colorbar(
      title = "Expression Level", title.position = "top")) +
    cowplot::theme_cowplot(font_size = 22, font_family = "Arial") +
    theme_dimred(linesize = 1.5 * PT) +
    ggplot2::theme(
      axis.title = ggplot2::element_text(size = 22),
      plot.title = ggplot2::element_text(size = 22, face = "plain", hjust = 0.5),
      legend.title = ggplot2::element_text(size = 20),
      legend.text = ggplot2::element_text(size = 18),
      legend.key.width = grid::unit(1.2, "cm"),
      legend.key.height = grid::unit(0.7, "cm"),
      legend.direction = "horizontal",
      legend.title.align = 0.5,
      legend.position = c(0.55, 0))
})

p3
```

```{r eval=FALSE, include=FALSE}
plotpowerpoint(p3, "../data/Figure.pptx",
  location = grid::unit(c(3.75, 29.84, 15.83, 13.64), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

```{r fig.height=7, fig.width=7}
p4 <- local({
  scater::plotReducedDim(
      sce_mock, "UMAP",
      colour_by = "TraesCS2A02G350900",
      order_by = "TraesCS2A02G350900",
      point_alpha = 1, point_size = 0.5,
      rasterise = TRUE) +
    ggplot2::scale_color_gradient(low = "grey", high = "blue") +
    ggplot2::coord_fixed() +
    ggplot2::ggtitle("TraesCS2A02G350900") +
    ggplot2::labs(x = "UMAP_1", y = "UMAP_2") +
    ggplot2::guides(color = ggplot2::guide_colorbar(
      title = "Expression Level", title.position = "top")) +
    cowplot::theme_cowplot(font_size = 22, font_family = "Arial") +
    theme_dimred(linesize = 1.5 * PT) +
    ggplot2::theme(
      axis.title = ggplot2::element_text(size = 22),
      plot.title = ggplot2::element_text(size = 22, face = "plain", hjust = 0.5),
      legend.title = ggplot2::element_text(size = 20),
      legend.text = ggplot2::element_text(size = 18),
      legend.key.width = grid::unit(1.2, "cm"),
      legend.key.height = grid::unit(0.7, "cm"),
      legend.direction = "horizontal",
      legend.title.align = 0.5,
      legend.position = c(0.55, 0))
})

p4
```

```{r eval=FALSE, include=FALSE}
plotpowerpoint(p4, "../data/Figure.pptx",
  location = grid::unit(c(3.75, 44.72, 15.83, 13.64), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

### Cluster stability

```{r}
silhouette_scores <- scclusteval::CalculateSilhouette(
  obj_mock, dims = 1:20, reduction = "harmony")

scclusteval::SilhouetteRainCloudPlot(silhouette_scores) +
  ggplot2::scale_color_manual(values = mock_celltype_colors) +
  ggplot2::scale_fill_manual(values = mock_celltype_colors)
```

```{r}
subsample <- readRDS("../results/ClustEval-2nd/gather_subsample.rds")
fullsample <- readRDS("../results/ClustEval-2nd/gather_fullsample.rds")
```

```{r}
scclusteval::JaccardRainCloudPlot(
    #rep(list(Seurat::Idents(obj_mock)), 100),
  dplyr::filter(subsample, pc == 20, k_param == 20, resolution == 0.6)$original_ident,
    dplyr::filter(subsample, pc == 20, k_param == 20, resolution == 0.6)$recluster_ident) + 
  ggplot2::geom_hline(yintercept = c(0.6, 0.75), linetype = 2) +
  ggplot2::xlab("cluster id w/ pc=20 k=20 res=0.6") +
  # ggplot2::scale_x_discrete(limits = mock_celltypes) +
  # ggplot2::scale_color_manual(values = mock_celltype_colors) +
  # ggplot2::scale_fill_manual(values = mock_celltype_colors) +
  NULL
```

```{r}
scclusteval::JaccardRainCloudPlot(
    rep(list(Seurat::Idents(obj_mock)), 100),
    dplyr::filter(subsample, pc == 20, k_param == 20, resolution == 0.6)$recluster_ident) + 
  ggplot2::geom_hline(yintercept = c(0.6, 0.75), linetype = 2) +
  ggplot2::xlab("cluster id w/ pc=20 k=20 res=0.6") +
  ggplot2::scale_x_discrete(limits = mock_celltypes) +
  ggplot2::scale_color_manual(values = mock_celltype_colors) +
  ggplot2::scale_fill_manual(values = mock_celltype_colors) +
  NULL
```

```{r}
rm(subsample, fullsample)
gc()
```


## Supplementary Figure 3

### Cell Proportion

```{r}
ca_res_mock <- readRDS(Sys.glob(
  "../results/ObjectCache/DifferentialAbundance/ca_res_nonclust_objmock_*.rds"))
ca_res_mock$results$cellTypes <- forcats::fct_relevel(ca_res_mock$results$cellTypes, mock_celltypes)
```

```{r fig.height=10, fig.width=14}
ca_res_mock$results$cellTypes <- dplyr::recode(as.character(ca_res_mock$results$cellTypes), !!!cluster_num)
ca_res_mock$results$cellTypes <- factor(ca_res_mock$results$cellTypes, levels = cluster_num[mock_celltypes])

p1 <- local({
  df_to_plot <- ca_res_mock$results %>%
    tibble::as_tibble() %>%
    dplyr::filter(method == "BCa") %>%
    dplyr::mutate(
      time = sapply(strsplit(cond, split = "-"), "[", 1),
      rep = sapply(strsplit(as.character(subject), split = "-"),
                   function(x) paste(x[2:3], collapse = "-")),
      treatment = sapply(strsplit(cond, split = "-"), "[", 2)
    )

  prop_df <- ca_res_mock$thetastar %>%
    as.data.frame()

  colnames(prop_df) <- paste("BS_", seq_len(ncol(ca_res_mock$thetastar)), sep = "")

  df_to_plot <- dplyr::bind_cols(df_to_plot, prop_df) %>%
    tidyr::pivot_longer(tidyr::starts_with("BS_"), names_to = "bootstrap", values_to = "prop")

  ggplot2::ggplot(df_to_plot, ggplot2::aes(x = time, y = prop, fill = cellTypes)) +
    ggplot2::geom_boxplot(
      width = 0.2, outlier.shape = NA,
      position = ggplot2::position_dodge(width = 0.4)) +
    ggplot2::geom_line(
      mapping = ggplot2::aes(group = treatment),
      stat = "summary", fun = median,
      position = ggplot2::position_dodge(width = 0.4)) +
    ggplot2::scale_y_continuous(labels = scales::percent) +
    ggplot2::facet_wrap(~ cellTypes, scales = "free_y", ncol = 3, drop = FALSE,
      labeller = ggplot2::labeller(cellTypes = \(x) paste("cluster:", x))) +
    ggplot2::scale_fill_manual(values = setNames(mock_celltype_colors, cluster_num[mock_celltypes])) +
    ggplot2::guides(fill = "none") +
    ggplot2::labs(x = "Time", y = "Proportion") +
    ggplot2::theme_bw(base_size = 22) +
    ggplot2::theme(
      panel.grid = ggplot2::element_blank(),
      strip.text = ggplot2::element_text(
        size = 18, vjust = 0.2, margin = ggplot2::margin(t = 4, b = 4)),
      axis.text.x = ggplot2::element_text(
        size = 16, angle = 0, hjust = 0.5, vjust = 1, color = "black"),
      axis.ticks = ggplot2::element_line(linewidth = 1.5 * PT),
      strip.background = ggplot2::element_blank(),
      panel.border = ggplot2::element_rect(linewidth = 1.5 * PT)
    ) + rotate_x_labels()
})

p2 <- rhapsodykit::calculateClusterProportion(
  sce_mock$body_layer, sce_mock$sample_id, sce_mock$group_id) |>
  dplyr::mutate(percent = scales::label_percent(accuracy = 0.1)(frequency)) |>
  ggplot2::ggplot(ggplot2::aes(
    x = sample_id, y = frequency, fill = cluster_id)) +
  ggplot2::facet_wrap(~group_id, nrow = 1, scales = "free_x") +
  ggplot2::geom_bar(
    stat = "identity", col = "white",
    width = 1, size = 0.2, position = "stack") +
  ggplot2::geom_text(
    mapping = ggplot2::aes(label = percent),
    stat = "identity", size = 16 * FT, angle = 90,
    position = ggplot2::position_fill(vjust = 0.5)) +
  ggplot2::scale_y_continuous(breaks = seq(0, 1, 0.2), expand = c(0, 0)) +
  ggplot2::scale_x_discrete(expand = c(0, 0)) +
  ggplot2::scale_fill_manual(values = c(L1 = "#a0cbe8", L2 = "#8cd17d", L3 = "#ff9d9a")) +
  ggplot2::guides(fill = ggplot2::guide_legend(title = "Body Layer")) +
  ggplot2::ylab("Proportion") +
  ggplot2::theme_grey() +
  ggplot2::theme(
    aspect.ratio = NULL,
    axis.title.x = ggplot2::element_blank(),
    axis.title.y = ggplot2::element_text(family = "Arial", size = 20, color = "black"),
    axis.text = ggplot2::element_text(family = "Arial", size = 16, color = "black"),
    axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, hjust = 1),
    axis.text.y = ggplot2::element_blank(),
    axis.ticks.y = ggplot2::element_blank(),
    panel.grid = ggplot2::element_blank(),
    panel.spacing = grid::unit(2, "mm"),
    strip.background = ggplot2::element_blank(),
    strip.text = ggplot2::element_blank(),
    legend.title = ggplot2::element_text(family = "Arial", size = 20),
    legend.text = ggplot2::element_text(family = "Arial", size = 18),
    legend.key.size = grid::unit(0.8, "cm"),
    legend.position = "bottom",
    legend.justification = "left",
    legend.direction = "horizontal",
    plot.margin = ggplot2::margin(l = 24)
  ) +
  rotate_x_labels()

(p12 <- p2 + p1 + patchwork::plot_layout(widths = c(1, 1.5)))
```

```{r eval=FALSE, include=FALSE}
plotpowerpoint(p12, "../data/Figure.pptx",
  location = grid::unit(c(2.69, 4.68, 30, 28.43), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

### CNA Results

```{r}
sce_mock$time_val <- as.numeric(factor(
  sce_mock$time, levels = c("0DPI", "1DPI", "2DPI", "3DPI")))
sce_mock$sample <- as.character(sce_mock$sample_id)

sce_mock <- association.SingleCellExperiment(
  sce_mock, test_var = "time_val", samplem_key = "sample_id", dimred = "HARMONY")
```

```{r fig.height=6, fig.width=14}
p3 <- scater::plotReducedDim(
    sce_mock,
    dimred = "UMAP",
    point_size = 0.5,
    point_alpha = 1,
    colour_by = "cna_ncorrs",
    order_by = I(sample(seq_len(ncol(sce_mock)))),
    theme_size = 14) +
  ggplot2::scale_color_gradientn(colours = rev(ggthemes::ggthemes_data[[
    c("tableau", "color-palettes", "ordered-diverging", "Orange-Blue Diverging")]][["value"]])) +
  ggplot2::guides(color = ggplot2::guide_colorbar(
    title = "Coefficient", title.position = "top")) +
  ggplot2::coord_fixed() +
  center_plot_title() +
  cowplot::theme_cowplot(font_size = 22, font_family = "Arial") +
  theme_dimred(linesize = 1.5 * PT) +
  remove_legend_title() +
  theme(
    legend.title = ggplot2::element_text(size = 20),
    legend.text = ggplot2::element_text(size = 18),
    legend.key.width = grid::unit(1.2, "cm"),
    legend.key.height = grid::unit(0.7, "cm"),
    legend.direction = "horizontal",
    legend.title.align = 0.5,
    legend.position = c(0.55, 0)
  )

p3 <- ggrastr::rasterise(p3)

p4 <- plotNcorrRidges(sce_mock, theme_size = 14, group_by = "tissue",
                      y_expand = ggplot2::expansion(mult = c(0, 0.4))) +
  ggplot2::scale_fill_manual(values = mock_tissues_colors) +
  ggplot2::xlab("Neighborhood coefficient") +
  ggplot2::ylab("Cell cluster") +
  ggplot2::guides(fill = "none") +
  ggplot2::theme(
    axis.text = ggplot2::element_text(size = 16),
    axis.title = ggplot2::element_text(size = 20)
  )

(p <- p3 + p4 +
    patchwork::plot_annotation(tag_levels = "A") &
    font_plot_tag(size = 16))
```

```{r eval=FALSE, include=FALSE}
plotpowerpoint(p3, "../data/Figure.pptx",
  location = grid::unit(c(33.23, 4.68, 18.2, 14.52), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

```{r eval=FALSE, include=FALSE}
plotpowerpoint(p4, "../data/Figure.pptx",
  location = grid::unit(c(33.23, 19.3, 18.2, 11), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```

### Module Scores

```{r}
ncorrs <- as(matrix(metadata(sce_mock)$cnaRes$ncorrs, ncol = 1), "dgCMatrix")
genes <- qlcMatrix::corSparse(ncorrs, t(logcounts(sce_mock)))
genes <- structure(as.numeric(genes), names = rownames(logcounts(sce_mock)))
genes <- sort(genes, decreasing = TRUE)
```

```{r}
gsea <- clusterProfiler::gseGO(
  genes,
  ont = "BP",
  OrgDb = "org.Taestivum.iwgsc.db",
  keyType = "GID"
)
```

```{r fig.height=10, fig.width=10, eval=FALSE}
p <- gsea |>
  enrichplot::pairwise_termsim(showCategory = 600) |>
  enrichplot::treeplot(
    showCategory = 600,
    cluster.params = list(
      n = 12, label_words_n = 0, label_format = 5),
    color = "NES",
    geneClusterPanel = "dotplot") +
  ggplot2::scale_color_gradient2(low = "blue", mid = "white", high = "red", name = "NES") +
  NULL

plotpowerpoint(
  ggobj = p, template = "../data/60x120.pptx")
```

```{r}
GO_DATA <- clusterProfiler:::get_GO_data("org.Taestivum.iwgsc.db", "BP", "GID")

pathways <- GO_DATA$PATHID2EXTID |>
  tibble::enframe(name = "GO", value = "Gene") |>
  tidyr::unnest(Gene) |>
  dplyr::filter(Gene %in% rownames(sce_mock)) |>
  dplyr::group_by(GO) |>
  dplyr::summarise(Gene = list(Gene)) |>
  tibble::deframe()
```

```{r}
negcorr <- c(
  "autophagy",
  "cell wall organization",
  "negative regulation of abscisic acid-activated signaling pathway",
  "positive regulation of post-embryonic development",
  "plant-type cell wall organization",
  "microtubule-based movement",
  "positive regulation of growth",
  "spindle organization",
  "phospholipid transport"
)

names(negcorr) <- names(GO_DATA$PATHID2NAME)[match(negcorr, GO_DATA$PATHID2NAME)]

poscorr <- c(
  "ribosome biogenesis",
  "chlorophyll biosynthetic process",
  "photosynthesis, light reaction",
  "cytoplasmic translation",
  "negative regulation of developmental growth",
  "response to cytokinin",
  "photosynthetic electron transport chain",
  "response to sucrose",
  "positive regulation of translation",
  "abscisic acid biosynthetic process"
)

names(poscorr) <- names(GO_DATA$PATHID2NAME)[match(poscorr, GO_DATA$PATHID2NAME)]
```

```{r}
pathway_summed <- calcPathwayActivity(
  sce_mock, pathways[c(names(negcorr), names(poscorr))],
  method = "seuratModuleScore",
  annotation = gsea@result)
```

```{r}
modcor <- cor(x = metadata(sce_mock)$cnaRes$ncorrs,
    y = t(assay(pathway_summed, "activity")),
    use = "complete.obs", method = "pearson")
modcor <- drop(modcor)

names(modcor) <- c(negcorr, poscorr)[names(modcor)]

sort(modcor)
```

```{r}
selected_go <- data.frame(
  Term = c("positive regulation of post-embryonic development",
           "autophagy", "plant-type cell wall organization",
           "negative regulation of developmental growth",
           "ribosome biogenesis", "abscisic acid biosynthetic process"),
  Type = rep(c("Negative", "Positive"), each = 3)
)

selected_go$ID <- names(c(negcorr, poscorr))[match(selected_go$Term, c(negcorr, poscorr))]
selected_go$R <- modcor[selected_go$Term]
```

```{r fig.height=7, fig.width=14.7}
p5 <- selected_go |>
  purrr::pmap(function(Term, Type, ID, R) {
    p <- plotPathwayActivation(
        pathway_summed, ID,
        dimred = "UMAP", point_size = 0.5,
        theme_size = 12) +
      ggplot2::annotate("text", label = paste("r =", round(R, 2), collapse = ""),
                        x = -10, y = 15, size = 16 * FT) +
      ggplot2::coord_fixed() +
      cowplot::theme_cowplot(font_size = 22, font_family = "Arial") +
      theme_dimred(linesize = 1.5 * PT) +
      remove_legend_title() +
      ggplot2::theme(
        plot.title = ggplot2::element_text(size = 16, face = "plain")
      )
    ggrastr::rasterise(p)
  }) |>
  patchwork::wrap_plots(ncol = 3)

p5
```

```{r eval=FALSE, include=FALSE}
plotpowerpoint(p5, "../data/Figure.pptx",
  location = grid::unit(c(2.57, 30.75, 48.86, 23.32), "cm"),
  fonts = list(sans = "Arial", serif = "Arial", mono = "Courier"))
```
