---
title: "Mesophyll and Vascular Bundle Lineages"
author: "Altair Wei"
date: '2022-11-04'
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
source("../scripts/TrajectoryUtilities.R")
source("../scripts/EnrichmentUtilities.R")
```

```{python}
import scvelo as scv
import matplotlib.pyplot as plt
```

```{r message=FALSE, warning=FALSE, results="hide"}
library(org.Taestivum.iwgsc.db)
orgdb <- org.Taestivum.iwgsc.db
```

```{r}
sce_mock <- readRDS(Sys.glob(
  "../results/ObjectCache/TrajectoryInference/sce_mock_spliced_*.rds"))
```

## Mesophyll Lineages

```{r}
sce_me <- sce_mock[, sce_mock$cluster_id %in% c(paste0("Me_", c("α", "β", "γ", "δ", "ε")), "Va_β")]
```

### 降维可视化 {.tabset}

```{r}
sce_me <- scater::runUMAP(sce_me, dimred = "HARMONY", ncomponents = 20L)
sce_me <- runDiffusionMap(sce_me, dimred = "HARMONY", seed = 1118)
sce_me <- runPHATE(sce_me, dimred = "HARMONY", seed = 1204, knn = 5, t = 15, gamma = 0)
```

#### UMAP

```{r}
scater::plotReducedDim(sce_me, dimred = "UMAP",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

#### t-SNE

```{r}
scater::plotReducedDim(sce_me, dimred = "TSNE",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

#### PHATE

```{r}
scater::plotReducedDim(sce_me, dimred = "PHATE",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

#### DiffusionMap

```{r}
scater::plotReducedDim(sce_me, dimred = "DiffusionMap",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

### 叶肉细胞群的差别

```{r}
obj_me <- Seurat::as.Seurat(sce_me)
Seurat::Idents(obj_me) <- droplevels(sce_me$cluster_id)
```

#### Chlorenchyma

```{r}
de_β <- Seurat::FindMarkers(obj_me, ident.1 = "Me_β", min.pct = 0.25)
```

```{r}
ggplot2::ggplot(de_β, ggplot2::aes(x = avg_log2FC, y = -log10(p_val_adj))) +
  ggplot2::geom_point() +
  ggplot2::theme_minimal()
```

```{r}
de_β_cut <- dplyr::filter(de_β, p_val_adj < 0.01)
geneList <- structure(de_β_cut$avg_log2FC, names = rownames(de_β_cut))
listUpDown <- list(
  upregulated = names(geneList[geneList > 0]),
  downregulated = names(geneList[geneList < 0])
)
ego <- clusterProfiler::compareCluster(
  listUpDown, OrgDb = orgdb,
  fun = "enrichGO", universe = rownames(obj_me),
  ont = "BP", keyType = "GID")
```

```{r}
ego <- clusterProfiler::simplify(ego, cutoff = 0.7)
```

```{r fig.height=7, fig.width=10, dev="svg"}
ego |>
  enrichplot::pairwise_termsim() |>
  enrichplot::treeplot(
    showCategory = 20, nCluster = 12,
    offset = 20, offset_tiplab = 10,
    label_format_cladelab = 5, size = "count",
    geneClusterPanel = "dotplot", nWords = 0) +
  ggplot2::scale_y_continuous(expand = ggplot2::expansion(mult = c(0, 0), add = c(4, 0))) +
  NULL
```

```{r}
printCompareORATable(ego)
```


#### Subepidermal mesophyll

```{r}
de_ε <- Seurat::FindMarkers(obj_me, ident.1 = "Me_ε", min.pct = 0.25)
```

```{r}
ggplot2::ggplot(de_ε, ggplot2::aes(x = avg_log2FC, y = -log10(p_val_adj))) +
  ggplot2::geom_point() +
  ggrepel::geom_text_repel(
    data = data.frame(
      gene = rownames(de_ε)[de_ε$avg_log2FC > 1],
      xpos = with(de_ε, avg_log2FC[avg_log2FC > 1]),
      ypos = with(de_ε, (-log10(p_val_adj))[avg_log2FC > 1])
    ),
    mapping = ggplot2::aes(label = gene, x = xpos, y = ypos)) +
  ggplot2::theme_minimal()
```

```{r}
de_ε_cut <- dplyr::filter(de_ε, p_val_adj < 0.01)
geneList <- structure(de_ε_cut$avg_log2FC, names = rownames(de_ε_cut))

ego <- clusterProfiler::compareCluster(
  geneClusters = list(
    upregulated = names(geneList[geneList > 0]),
    downregulated = names(geneList[geneList < 0])
  ),
  OrgDb = orgdb, fun = "enrichGO",
  universe = rownames(obj_me),
  ont = "BP", keyType = "GID")
```

```{r}
ego <- clusterProfiler::simplify(ego, cutoff = 0.7)
```

```{r fig.height=7, fig.width=10, dev="svg"}
ego |>
  enrichplot::pairwise_termsim() |>
  enrichplot::treeplot(
    showCategory = 20, nCluster = 12,
    offset = 20, offset_tiplab = 10,
    label_format_cladelab = 5, size = "count",
    geneClusterPanel = "dotplot", nWords = 0) +
  ggplot2::scale_y_continuous(expand = ggplot2::expansion(mult = c(0, 0), add = c(4, 0))) +
  NULL
```

```{r}
printCompareORATable(ego)
```

#### Mesophyll precursor

```{r}
de_γ <- Seurat::FindMarkers(obj_me, ident.1 = "Me_γ", min.pct = 0.25)
```

```{r}
ggplot2::ggplot(de_γ, ggplot2::aes(x = avg_log2FC, y = -log10(p_val_adj))) +
  ggplot2::geom_point() +
  ggplot2::theme_minimal()
```

```{r}
de_γ_cut <- dplyr::filter(de_γ, p_val_adj < 0.01)
geneList <- structure(de_γ_cut$avg_log2FC, names = rownames(de_γ_cut))

ego <- clusterProfiler::compareCluster(
  geneClusters = list(
    upregulated = names(geneList[geneList > 0]),
    downregulated = names(geneList[geneList < 0])
  ),
  OrgDb = orgdb, fun = "enrichGO",
  universe = rownames(obj_me),
  ont = "BP", keyType = "GID")
```

```{r}
ego <- clusterProfiler::simplify(ego, cutoff = 0.7)
```

```{r fig.height=7, fig.width=10, dev="svg"}
ego |>
  enrichplot::pairwise_termsim() |>
  enrichplot::treeplot(
    showCategory = 20, nCluster = 12,
    offset = 20, offset_tiplab = 10,
    label_format_cladelab = 5, size = "count",
    geneClusterPanel = "dotplot", nWords = 0) +
  ggplot2::scale_y_continuous(expand = ggplot2::expansion(mult = c(0, 0), add = c(4, 0))) +
  NULL
```

```{r}
printCompareORATable(ego)
```

#### Result summary

```{r}
me_ora_df <- purrr::imap(
  list(
    Me_β = de_β_cut,
    Me_ε = de_ε_cut,
    Me_γ = de_γ_cut),
  function(df, name) data.frame(
    gene = rownames(df),
    cellType = name,
    regulation = ifelse(df$avg_log2FC > 0, "up", "down")
  )) |> dplyr::bind_rows()
```

```{r}
ora <- clusterProfiler::compareCluster(
  gene ~ cellType + regulation, data = me_ora_df,
  OrgDb = orgdb, fun = "enrichGO",
  universe = rownames(obj_me),
  ont = "BP", keyType = "GID")
```

```{r}
ora <- clusterProfiler::simplify(ora, cutoff = 0.5)
```

```{r fig.height=16, fig.width=14}
ora@compareClusterResult$cellType <- factor(
  ora@compareClusterResult$cellType, c("Me_β", "Me_ε", "Me_γ"))
ora@compareClusterResult$regulation <- factor(
  ora@compareClusterResult$regulation, c("up", "down"))

enrichplot::dotplot(
    ora, x = "regulation",
    showCategory = 40, by = "count", label_format = 100) +
  ggplot2::facet_grid(~cellType) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
```


### 发育轨迹推断

```{r}
sce_me <- xfun::cache_rds(
  file = "sce_mock_me_slingshot.rds",
  dir = "../results/ObjectCache/TrajectoryInference/",
  expr = slingshot::slingshot(sce_me,
    clusterLabels = sce_me$cluster_id,
    reducedDim = "UMAP", approx_points = 500,
    start.clus = "Me_γ", end.clus = c("Me_ε", "Me_β", "Va_β"))
)
```

#### 细胞世系 {.tabset}

##### PHATE

```{r}
plotSlingshotCurveOnReduc(sce_me, dimred = "PHATE",
                          path_size = 2, text_by = "cellType") +
  ggplot2::coord_fixed()
```

```{r fig.height=4, fig.width=14}
plotLineagesOnReduc(sce_me, dimred = "PHATE")
```

#### 细胞种群路径

```{r}
psts <- slingshot::slingPseudotime(sce_me$slingshot)
colnames(psts) <- paste0(colnames(psts), "_pseudotime")
cwts <- slingshot::slingCurveWeights(sce_me$slingshot)
colnames(cwts) <- paste0(colnames(cwts), "_curveweights")

curve_data <- as.data.frame(cbind(psts, cwts)) %>%
  dplyr::mutate(
    cells = rownames(.),
    conditions = sce_me$treatment,
    time = sce_me$time,
    cellType = sce_me$cellType) %>%
  tidyr::pivot_longer(dplyr::starts_with("Lineage"),
                      names_to = c("lineages", "type"),
                      names_sep = "_",
                      values_to = "value") %>%
  tidyr::pivot_wider(names_from = type, values_from = value)
```

```{r fig.height=3, fig.width=7}
dplyr::filter(curve_data, !is.na(pseudotime)) |>
ggplot2::ggplot(ggplot2::aes(x = pseudotime, color = cellType)) +
  ggplot2::geom_jitter(ggplot2::aes(
    y = forcats::fct_reorder(cellType, pseudotime, .desc = TRUE),
    alpha = curveweights)) +
  ggplot2::ylab("cellType") +
  ggthemes::scale_color_tableau() +
  ggplot2::facet_wrap(~lineages, scales = "free_x", ncol = 3) +
  ggridges::theme_ridges() +
  ggplot2::theme(
    strip.background = ggplot2::element_rect(fill = "white"))
```

#### 拟时间分布

```{r}
psts <- slingshot::slingPseudotime(sce_me$slingshot) %>%
  as.data.frame() %>%
  dplyr::mutate(
    cells = rownames(.),
    time = sce_me$time) %>%
  tidyr::pivot_longer(dplyr::starts_with("Lineage"),
    values_to = "pseudotime", names_to = "lineages")
```

```{r fig.height=3, fig.width=7}
dplyr::filter(psts, !is.na(pseudotime)) |>
ggplot2::ggplot(ggplot2::aes(x = pseudotime, fill = time)) +
    ggplot2::geom_density(alpha = .5) +
    #ggridges::geom_density_ridges(ggplot2::aes(y = time), scale = 4) +
    ggplot2::scale_fill_brewer(type = "qual") +
    ggplot2::facet_wrap(~lineages, scales = "free_x", ncol = 1) +
    cowplot::theme_cowplot() +
    ggplot2::theme(
      strip.background = ggplot2::element_rect(fill = "white"))
```

```{r fig.height=3, fig.width=7}
dplyr::filter(psts, !is.na(pseudotime)) |>
ggplot2::ggplot(ggplot2::aes(x = pseudotime, fill = time)) +
    #ggplot2::geom_density(alpha = .5) +
    ggridges::geom_density_ridges(ggplot2::aes(y = time), scale = 4) +
    ggplot2::scale_fill_brewer(type = "qual") +
    ggplot2::facet_wrap(~lineages, scales = "free_x", ncol = 1) +
    cowplot::theme_cowplot() +
    ggplot2::theme(
      strip.background = ggplot2::element_rect(fill = "white"))
```

## Outer Sheath Lineage

```{r}
sce_os <- sce_mock[, sce_mock$cluster_id %in% c("Me_γ", "Va_β")]
```

### 降维可视化 {.tabset}

```{r}
sce_os <- scater::runUMAP(sce_os, dimred = "HARMONY", ncomponents = 20L)
sce_os <- runDiffusionMap(sce_os, dimred = "HARMONY", n_eigs = 2L, ncomponents = 2L, seed = 1201)
sce_os <- runPHATE(sce_os, dimred = "HARMONY", seed = 1118)
```

#### PHATE

```{r}
scater::plotReducedDim(sce_os, dimred = "PHATE",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

#### DiffusionMap

```{r}
scater::plotReducedDim(sce_os, dimred = "DiffusionMap",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

### 发育轨迹推断

```{r}
sce_os <- xfun::cache_rds(
  file = "sce_mock_os_slingshot.rds",
  dir = "../results/ObjectCache/TrajectoryInference/",
  expr = slingshot::slingshot(sce_os,
    clusterLabels = sce_os$cluster_id,
    reducedDim = "UMAP", approx_points = 500,
    start.clus = "Me_γ")
)
```

#### 细胞世系 {.tabset}

##### PHATE

```{r}
plotLineagesOnReduc(sce_os, dimred = "PHATE",
                    path_size = 1, point_size = 0.5, text_by = "cellType")
```

##### DiffusionMap

```{r}
plotLineagesOnReduc(sce_os, dimred = "DiffusionMap",
                    path_size = 1, point_size = 0.5, text_by = "cellType")
```

#### 拟时间分布

```{r}
psts <- slingshot::slingPseudotime(sce_os$slingshot) %>%
  as.data.frame() %>%
  dplyr::mutate(
    cells = rownames(.),
    time = sce_os$time) %>%
  tidyr::pivot_longer(dplyr::starts_with("Lineage"),
    values_to = "pseudotime", names_to = "lineages")
```

```{r fig.height=3, fig.width=7}
dplyr::filter(psts, !is.na(pseudotime)) |>
ggplot2::ggplot(ggplot2::aes(x = pseudotime, fill = time)) +
    ggplot2::geom_density(alpha = .5) +
    #ggridges::geom_density_ridges(ggplot2::aes(y = time), scale = 4) +
    ggplot2::scale_fill_brewer(type = "qual") +
    ggplot2::facet_wrap(~lineages, scales = "free_x", ncol = 1) +
    cowplot::theme_cowplot() +
    ggplot2::theme(
      strip.background = ggplot2::element_rect(fill = "white"))
```

```{r fig.height=3, fig.width=7}
dplyr::filter(psts, !is.na(pseudotime)) |>
ggplot2::ggplot(ggplot2::aes(x = pseudotime, fill = time)) +
    #ggplot2::geom_density(alpha = .5) +
    ggridges::geom_density_ridges(ggplot2::aes(y = time), scale = 4) +
    ggplot2::scale_fill_brewer(type = "qual") +
    ggplot2::facet_wrap(~lineages, scales = "free_x", ncol = 1) +
    cowplot::theme_cowplot() +
    ggplot2::theme(
      strip.background = ggplot2::element_rect(fill = "white"))
```

## Xylem Lineage

```{r}
sce_xyl <- sce_mock[, sce_mock$cluster_id %in% c("MPV", "Va_α", "Va_γ", "BS")]
```

### 降维可视化 {.tabset}

```{r}
sce_xyl <- scater::runUMAP(sce_xyl, dimred = "HARMONY", ncomponents = 20L)
sce_xyl <- runDiffusionMap(sce_xyl, dimred = "HARMONY", n_eigs = 2L, ncomponents = 2L, seed = 1201)
sce_xyl <- runPHATE(sce_xyl, dimred = "HARMONY", seed = 1118)
```

#### PHATE

```{r}
scater::plotReducedDim(sce_xyl, dimred = "PHATE",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

#### DiffusionMap

```{r}
scater::plotReducedDim(sce_xyl, dimred = "DiffusionMap",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

### 发育轨迹推断

```{r}
sce_xyl <- xfun::cache_rds(
  file = "sce_mock_xyl_slingshot.rds",
  dir = "../results/ObjectCache/TrajectoryInference/",
  expr = slingshot::slingshot(sce_xyl,
    clusterLabels = sce_xyl$cluster_id,
    reducedDim = "UMAP", approx_points = 500,
    start.clus = "MPV")
)
```

#### 细胞世系 {.tabset}

##### PHATE

```{r}
plotLineagesOnReduc(sce_xyl, dimred = "PHATE",
                    path_size = 1, point_size = 0.5, text_by = "cellType")
```

##### DiffusionMap

```{r}
plotLineagesOnReduc(sce_xyl, dimred = "DiffusionMap",
                    path_size = 1, point_size = 0.5, text_by = "cellType")
```

#### 拟时间分布

```{r}
psts <- slingshot::slingPseudotime(sce_xyl$slingshot) %>%
  as.data.frame() %>%
  dplyr::mutate(
    cells = rownames(.),
    time = sce_xyl$time) %>%
  tidyr::pivot_longer(dplyr::starts_with("Lineage"),
    values_to = "pseudotime", names_to = "lineages")
```

```{r fig.height=3, fig.width=7}
dplyr::filter(psts, !is.na(pseudotime)) |>
ggplot2::ggplot(ggplot2::aes(x = pseudotime, fill = time)) +
    ggplot2::geom_density(alpha = .5) +
    #ggridges::geom_density_ridges(ggplot2::aes(y = time), scale = 4) +
    ggplot2::scale_fill_brewer(type = "qual") +
    ggplot2::facet_wrap(~lineages, scales = "free_x", ncol = 1) +
    cowplot::theme_cowplot() +
    ggplot2::theme(
      strip.background = ggplot2::element_rect(fill = "white"))
```

```{r fig.height=3, fig.width=7}
dplyr::filter(psts, !is.na(pseudotime)) |>
ggplot2::ggplot(ggplot2::aes(x = pseudotime, fill = time)) +
    #ggplot2::geom_density(alpha = .5) +
    ggridges::geom_density_ridges(ggplot2::aes(y = time), scale = 4) +
    ggplot2::scale_fill_brewer(type = "qual") +
    ggplot2::facet_wrap(~lineages, scales = "free_x", ncol = 1) +
    cowplot::theme_cowplot() +
    ggplot2::theme(
      strip.background = ggplot2::element_rect(fill = "white"))
```

## Preprocambium to mesophyll

```{r}
sce_preme <- sce_mock[, sce_mock$cluster_id %in% c("MPV", "Va_δ", "Me_γ")]
```

### 降维可视化 {.tabset}

```{r}
sce_preme <- scater::runUMAP(sce_preme, dimred = "HARMONY", ncomponents = 20L)
sce_preme <- runDiffusionMap(sce_preme, dimred = "HARMONY", n_eigs = 2L, ncomponents = 2L, seed = 1201)
sce_preme <- runPHATE(sce_preme, dimred = "HARMONY", seed = 1118)
```

#### PHATE

```{r}
scater::plotReducedDim(sce_preme, dimred = "PHATE",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

#### DiffusionMap

```{r}
scater::plotReducedDim(sce_preme, dimred = "DiffusionMap",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

### 发育轨迹推断

```{r}
sce_preme <- xfun::cache_rds(
  file = "sce_mock_preme_slingshot.rds",
  dir = "../results/ObjectCache/TrajectoryInference/",
  expr = slingshot::slingshot(sce_preme,
    clusterLabels = sce_preme$cluster_id,
    reducedDim = "UMAP", approx_points = 500,
    start.clus = "MPV")
)
```

#### 细胞世系 {.tabset}

##### PHATE

```{r}
plotLineagesOnReduc(sce_preme, dimred = "PHATE",
                    path_size = 1, point_size = 0.5, text_by = "cellType")
```

##### DiffusionMap

```{r}
plotLineagesOnReduc(sce_preme, dimred = "DiffusionMap",
                    path_size = 1, point_size = 0.5, text_by = "cellType")
```

#### 拟时间分布

```{r}
psts <- slingshot::slingPseudotime(sce_preme$slingshot) %>%
  as.data.frame() %>%
  dplyr::mutate(
    cells = rownames(.),
    time = sce_preme$time) %>%
  tidyr::pivot_longer(dplyr::starts_with("Lineage"),
    values_to = "pseudotime", names_to = "lineages")
```

```{r fig.height=3, fig.width=7}
dplyr::filter(psts, !is.na(pseudotime)) |>
ggplot2::ggplot(ggplot2::aes(x = pseudotime, fill = time)) +
    ggplot2::geom_density(alpha = .5) +
    #ggridges::geom_density_ridges(ggplot2::aes(y = time), scale = 4) +
    ggplot2::scale_fill_brewer(type = "qual") +
    ggplot2::facet_wrap(~lineages, scales = "free_x", ncol = 1) +
    cowplot::theme_cowplot() +
    ggplot2::theme(
      strip.background = ggplot2::element_rect(fill = "white"))
```

```{r fig.height=3, fig.width=7}
dplyr::filter(psts, !is.na(pseudotime)) |>
ggplot2::ggplot(ggplot2::aes(x = pseudotime, fill = time)) +
    #ggplot2::geom_density(alpha = .5) +
    ggridges::geom_density_ridges(ggplot2::aes(y = time), scale = 4) +
    ggplot2::scale_fill_brewer(type = "qual") +
    ggplot2::facet_wrap(~lineages, scales = "free_x", ncol = 1) +
    cowplot::theme_cowplot() +
    ggplot2::theme(
      strip.background = ggplot2::element_rect(fill = "white"))
```
