---
title: "RNA Velocity Estimation"
author: "Altair Wei"
date: '2022-08-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(SingleCellExperiment)
source("../scripts/TrajectoryUtilities.R")
plt <- reticulate::import("matplotlib.pyplot")
scv <- reticulate::import("scvelo")
```

```{python}
import scvelo as scv
import matplotlib.pyplot as plt
```

## 1 数据处理

### 合并 velocyto 数据集

```{r}
samples <- c(
  "0DPI-MOCK-1",
  "0DPI-MOCK-2",

  "1DPI-MOCK-1",
  "1DPI-MOCK-2",
  "1DPI-PNR2-1",
  "1DPI-PNR2-2",
  "1DPI-TR4-1",
  "1DPI-TR4-2",

  "2DPI-MOCK-1",
  "2DPI-MOCK-2",
  "2DPI-PNR2-1",
  "2DPI-PNR2-2",
  "2DPI-TR4-1",
  "2DPI-TR4-2",

  "3DPI-MOCK-1",
  "3DPI-MOCK-2",
  "3DPI-PNR2-1",
  "3DPI-PNR2-3",
  "3DPI-TR4-1",
  "3DPI-TR4-2"
)

data_folders <- structure(
  paste0("../results/RawOutput/", samples), names = samples)

data_list <- xfun::cache_rds(
  file = "velocity_data_list.rds",
  dir = "../results/ObjectCache/TrajectoryInference/",
  expr = lapply(data_folders, function(base_dir) {
    loom_file = paste0(base_dir, "/Velocyto/", basename(base_dir), ".loom")
    SeuratWrappers::ReadVelocity(loom_file)
   })
)
```

```{r}
fixCellNames <- function(mtx_list) {
  lapply(mtx_list, function(mtx) {
    colnames(mtx) <- colnames(mtx) |>
      stringr::str_sub(end = -2L) |>
      stringr::str_replace(stringr::fixed(":"), "_")
    mtx
  })
}

spliced <- xfun::cache_rds(
  file = "velocity_mtx_spliced.rds",
  dir = "../results/ObjectCache/TrajectoryInference/",
  expr = do.call(cbind, fixCellNames(lapply(data_list, "[[", "spliced")))
)

unspliced <- xfun::cache_rds(
  file = "velocity_mtx_unspliced.rds",
  dir = "../results/ObjectCache/TrajectoryInference/",
  expr = do.call(cbind, fixCellNames(lapply(data_list, "[[", "unspliced")))
)

ambiguous <- xfun::cache_rds(
  file = "velocity_mtx_ambiguous.rds",
  dir = "../results/ObjectCache/TrajectoryInference/",
  expr = do.call(cbind, fixCellNames(lapply(data_list, "[[", "ambiguous")))
)
```

```{r}
obj <- readRDS(Sys.glob(
  "../results/ObjectCache/IntegrationAndAnnotation/obj_annotated_*.rds"))
```

### 添加到 SingleCellExperiment 对象

```{r}
sce <- xfun::cache_rds(
  file = "sce_spliced.rds",
  dir = "../results/ObjectCache/TrajectoryInference/",
  expr = local({
    obj$cellType <- Seurat::Idents(obj)
    sce <- Seurat::as.SingleCellExperiment(obj, assay = "RNA")
    sce <- muscat::prepSCE(sce,
      kid = "ident", # subpopulation assignments
      gid = "group", # group IDs (ctrl/stim)
      sid = "sample", # sample IDs (ctrl/stim.1234)
      drop = FALSE
    )

    cell_names <- colnames(sce)
    gene_names <- rownames(sce)
    assay(sce, "spliced") <- spliced[gene_names, cell_names]
    assay(sce, "unspliced") <- unspliced[gene_names, cell_names]

    sce
  })
)
```

```{r}
rm(obj, data_list, spliced, unspliced, ambiguous)
gc()
```

## 2 MOCK 样品分析

### 读取数据

```{r}
sce_mock <- sce[, sce$treatment == "MOCK"]

sce_mock <- sce_mock[, sce_mock$cluster_id %in% c(
  paste0("Me_", 1:6), paste0("Va_", 1:4),
  "MPV_1", "MPV_2", "BS")]

embed_umap <- reducedDim(sce_mock, "UMAP")

sce_mock <- sce_mock[,
  embed_umap[, 1L] < 8
    & embed_umap[, 1L] > -8
    & embed_umap[, 2L] < 8
]

sce_mock$cluster_id <- droplevels(sce_mock$cluster_id)
sce_mock$cellType <- droplevels(sce_mock$cellType)

sce_sub <- sce_mock
```

### 降维可视化 {.tabset}

```{r run-dimred}
sce_sub <- runDiffusionMap(sce_sub, dimred = "HARMONY", seed = 2022,
                           ncomponents = 2, n_eigs = 2)
sce_sub <- runPHATE(sce_sub, dimred = "HARMONY", seed = 2022, knn = 15, t = 30)
```

#### t-SNE

```{r plot-dimred-tsne}
scater::plotReducedDim(
    sce_sub, dimred = "TSNE",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

#### UMAP

```{r plot-dimred-umap}
scater::plotReducedDim(
    sce_sub, dimred = "UMAP",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

#### PHATE

```{r plot-dimred-phate}
scater::plotReducedDim(
    sce_sub, dimred = "PHATE",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

#### DiffusionMap

```{r plot-dimred-dc}
scater::plotReducedDim(
    sce_sub, dimred = "DiffusionMap",
    point_size = 0.5,
    colour_by = "cellType",
    text_by = "cellType",
    text_size = 3) +
  ggplot2::guides(color = ggplot2::guide_legend(
    override.aes = list(size = 2))) +
  ggplot2::coord_fixed()
```

### 三维可视化 {.tabset}

```{r cell-3d-colors}
cellTypeColors <- ggthemes::tableau_color_pal("Tableau 20")(length(levels(sce_sub$cellType)))
```

#### PHATE

```{r dimred-3d-phate}
embed_phate <- calculatePHATE(
  sce_sub, dimred = "HARMONY", seed = 2022,
  knn = 15, t = 30, ncomponents = 3)

df_to_plot <- data.frame(
  Dim_1 = embed_phate[, 1],
  Dim_2 = embed_phate[, 2],
  Dim_3 = embed_phate[, 3],
  cellType = sce_sub$cluster_id
)

plotly::plot_ly(
  df_to_plot,
  x = ~ Dim_1, y = ~ Dim_2, z = ~ Dim_3,
  color = ~ cellType,
  colors = cellTypeColors,
  type = "scatter3d",
  mode = "markers",
  marker = list(size = 4)
)
```

#### DiffusionMap

```{r dimred-3d-dc}
embed_dc <- calculateDiffusionMap(
  sce_sub, dimred = "HARMONY", seed = 2022, ncomponents = 3)

df_to_plot <- data.frame(
  Dim_1 = embed_dc[, 1],
  Dim_2 = embed_dc[, 2],
  Dim_3 = embed_dc[, 3],
  cellType = sce_sub$cluster_id
)

plotly::plot_ly(
  df_to_plot,
  x = ~ Dim_1, y = ~ Dim_2, z = ~ Dim_3,
  color = ~ cellType,
  colors = cellTypeColors,
  type = "scatter3d",
  mode = "markers",
  marker = list(size = 4)
)
```

### 运行算法

```{r run-scvelo}
adata <- cache_pickle(
  file = "unitvelo_adata_mock.pickle",
  dir = "../results/ObjectCache/TrajectoryInference/",
  expr = runUniTVelo(
    sce_mock,
    label_key = "cluster_id",
    use.dimred = "HARMONY",
    assay.X = "logcounts",
    GPU = 0L,
    config = list(
      R2_ADJUST = TRUE,
      IROOT = NULL,
      FIT_OPTION = "1",
      AGENES_R2 = 1L
    )
  )
)
```

```{r update-embeddings}
adata$obsm$update(list(
  X_tsne = reducedDim(sce_sub, "TSNE"),
  X_umap = reducedDim(sce_sub, "UMAP"),
  X_phate = reducedDim(sce_sub, "PHATE"),
  X_dc = reducedDim(sce_sub, "DiffusionMap")
))

adata$obs$cellType <- droplevels(sce_sub$cluster_id)
```

> Here, the proportions of spliced/unspliced counts are displayed. Depending on the protocol used (Drop-Seq, Smart-Seq), we typically have between 10%-25% of unspliced molecules containing intronic sequences. We also advice you to examine the variations on cluster level to verify consistency in splicing efficiency. 

```{python plot-prop, fig.height=4, fig.width=8}
scv.pl.proportions(r.adata, groupby="cellType", figsize = (8, 4))
```

### 速率流线 {.tabset}

#### t-SNE

```{python plot-stream-tsne, fig.height=7, fig.width=7}
ax = scv.pl.velocity_embedding_stream(
  r.adata, basis="tsne", color="cellType", figsize= (7, 7), show=False)
ax.set_aspect('equal', 'box')
plt.show()
```

#### UMAP

```{python plot-stream-umap, fig.height=7, fig.width=7}
ax = scv.pl.velocity_embedding_stream(
  r.adata, basis="umap", color="cellType", figsize= (7, 7), show=False)
ax.set_aspect('equal', 'box')
plt.show()
```

### 速率网格 {.tabset}

#### t-SNE

```{python plot-grid-tsne, fig.height=7, fig.width=7}
ax = scv.pl.velocity_embedding_grid(
  r.adata, basis="tsne", color="cellType",
  arrow_length=3, arrow_size=1, figsize=(7, 7), show=False)
ax.set_aspect('equal', 'box')
plt.show()
```

#### UMAP

```{python plot-grid-umap, fig.height=7, fig.width=7}
ax = scv.pl.velocity_embedding_grid(
  r.adata, basis="umap", color="cellType",
  arrow_length=3, arrow_size=1, figsize=(7, 7), show=False)
ax.set_aspect('equal', 'box')
plt.show()
```

### 速率细胞 {.tabset}

#### t-SNE

```{python plot-velo-tsne, fig.height=12, fig.width=12}
ax = scv.pl.velocity_embedding(
  r.adata, basis="tsne", color="cellType", figsize=(12, 12),
  arrow_length=6, arrow_size=3, scale=2, dpi=300, show=False)
ax.set_aspect('equal', 'box')
plt.show()
```

#### UMAP

```{python plot-velo-umap, fig.height=12, fig.width=12}
ax = scv.pl.velocity_embedding(
  r.adata, basis="umap", color="cellType", figsize=(12, 12),
  arrow_length=6, arrow_size=3, scale=2, dpi=300, show=False)
ax.set_aspect('equal', 'box')
plt.show()
```

### 隐时间 {.tabset}

> scVelo’s latent time is based only on its transcriptional dynamics and represents
the cell’s internal clock. It captures aspects of the actual time better than similarity-based diffusion pseudo-time.

```{r}
scv$tl$latent_time(adata, min_likelihood=NULL)
```

#### t-SNE

```{python plot-latent-time-tsne, fig.height=7, fig.width=7}
ax = scv.pl.scatter(r.adata, basis="tsne",
  color='latent_time', color_map='gnuplot',
  size=20, figsize=(7, 7), show=False)
ax.set_aspect('equal', 'box')
plt.show()
```

#### UMAP

```{python plot-latent-time-umap, fig.height=7, fig.width=7}
ax = scv.pl.scatter(r.adata, basis="umap",
  color='latent_time', color_map='gnuplot',
  size=20, figsize=(7, 7), show=False)
ax.set_aspect('equal', 'box')
plt.show()
```

### 速度与连贯性 {.tabset}

```{r run-velocity-confidence}
scv$tl$velocity_confidence(adata)
sce_sub$velocity_length <- adata$obs$velocity_length
sce_sub$velocity_confidence <- adata$obs$velocity_confidence
```

#### UMAP

```{r plot-confidence-tsne, fig.height=7, fig.width=14}
p1 <- scater::plotReducedDim(sce_sub, dimred = "TSNE", point_size = 0.5,
                       colour_by = "velocity_length") +
  ggplot2::scale_colour_gradientn(colours = c("blue", "grey", "red")) +
  ggplot2::ggtitle("Velocity length") +
  ggplot2::coord_fixed()

p2 <- scater::plotReducedDim(sce_sub, dimred = "TSNE", point_size = 0.5,
                       colour_by = "velocity_confidence") +
  ggplot2::scale_colour_gradientn(colours = c("blue", "grey", "red")) +
  ggplot2::ggtitle("Velocity confidence") +
  ggplot2::coord_fixed()

p1 + p2
```

#### UMAP

```{r plot-confidence-umap, fig.height=7, fig.width=14}
p1 <- scater::plotReducedDim(sce_sub, dimred = "UMAP", point_size = 0.5,
                       colour_by = "velocity_length") +
  ggplot2::scale_colour_gradientn(colours = c("blue", "grey", "red")) +
  ggplot2::ggtitle("Velocity length") +
  ggplot2::coord_fixed()

p2 <- scater::plotReducedDim(sce_sub, dimred = "UMAP", point_size = 0.5,
                       colour_by = "velocity_confidence") +
  ggplot2::scale_colour_gradientn(colours = c("blue", "grey", "red")) +
  ggplot2::ggtitle("Velocity confidence") +
  ggplot2::coord_fixed()

p1 + p2
```

### 速率图 {.tabset}

#### t-SNE

```{python plot-velo-graph-tsne, fig.height=7, fig.width=7}
ax = scv.pl.velocity_graph(
  r.adata, threshold=0.8, basis="tsne",
  color="cellType", figsize=(7, 7), show=False)
ax.set_aspect('equal', 'box')
plt.show()
```

#### UMAP

```{python plot-velo-graph-umap, fig.height=7, fig.width=7}
ax = scv.pl.velocity_graph(
  r.adata, threshold=0.8, basis="umap",
  color="cellType", figsize=(7, 7), show=False)
ax.set_aspect('equal', 'box')
plt.show()
```

### PAGA 图 {.tabset}

```{python run-paga}
r.adata.uns['neighbors']['distances'] = r.adata.obsp['distances']
r.adata.uns['neighbors']['connectivities'] = r.adata.obsp['connectivities']

scv.tl.paga(r.adata, groups='cellType')
```

#### t-SNE

```{python plot-paga-tsne, fig.width=7, fig.height=7}
ax = scv.pl.paga(r.adata, basis='tsne', size=50, alpha=.1, legend_loc='on data',
            min_edge_width=2, node_size_scale=1.5, figsize=(7, 7), show=False)
ax.set_aspect('equal', 'box')
plt.show()
```

#### UMAP

```{python plot-paga-umap, fig.width=7, fig.height=7}
ax = scv.pl.paga(r.adata, basis='umap', size=50, alpha=.1, legend_loc='on data',
            min_edge_width=2, node_size_scale=1.5, figsize=(7, 7), show=False)
ax.set_aspect('equal', 'box')
plt.show()
```

## 2 PNR2 样本分析

```{r}
sce_pnr2 <- sce[, sce$treatment == "PNR2" | sce$time == "0DPI"]

sce_pnr2 <- sce_pnr2[, sce_pnr2$cluster_id %in% c(
  paste0("Me_", 1:6), paste0("Va_", 1:4),
  "MPV_1", "MPV_2", "BS")]

embed_umap <- reducedDim(sce_pnr2, "UMAP")

sce_pnr2 <- sce_pnr2[,
  embed_umap[, 1L] < 8
    & embed_umap[, 1L] > -8
    & embed_umap[, 2L] < 8
]

sce_pnr2$cluster_id <- droplevels(sce_pnr2$cluster_id)
sce_pnr2$cellType <- droplevels(sce_pnr2$cellType)

sce_sub <- sce_pnr2
```

### 降维可视化 {.tabset}

```{r}
<<run-dimred>>
```

#### t-SNE

```{r}
<<plot-dimred-tsne>>
```

#### UMAP

```{r}
<<plot-dimred-umap>>
```

#### PHATE

```{r}
<<plot-dimred-tsne>>
```

#### DiffusionMap

```{r}
<<plot-dimred-dc>>
```

### 三维可视化 {.tabset}

```{r}
<<cell-3d-colors>>
```

#### PHATE

```{r}
<<dimred-3d-phate>>
```

#### DiffusionMap

```{r}
<<dimred-3d-dc>>
```

### 运行算法

```{r}
adata <- cache_pickle(
  file = "unitvelo_adata_pnr2.pickle",
  dir = "../results/ObjectCache/TrajectoryInference/",
  expr = runUniTVelo(
    sce_pnr2,
    label_key = "cluster_id",
    use.dimred = "HARMONY",
    assay.X = "logcounts",
    GPU = 0L,
    config = list(
      R2_ADJUST = TRUE,
      IROOT = NULL,
      FIT_OPTION = "1",
      AGENES_R2 = 1L
    )
  )
)
```

```{r}
<<update-embeddings>>
```

```{python fig.height=4, fig.width=8}
<<plot-prop>>
```

### 速率流线 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-stream-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-stream-umap>>
```

### 速率网格 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-grid-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-grid-umap>>
```

### 速率细胞 {.tabset}

#### t-SNE

```{python fig.height=12, fig.width=12}
<<plot-velo-tsne>>
```

#### UMAP

```{python fig.height=12, fig.width=12}
<<plot-velo-umap>>
```

### 隐时间 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-latent-time-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-latent-time-umap>>
```

### 速度与连贯性 {.tabset}

```{r}
<<run-velocity-confidence>>
```

#### t-SNE

```{r fig.height=7, fig.width=14}
<<plot-confidence-tsne>>
```

#### UMAP

```{r fig.height=7, fig.width=14}
<<plot-confidence-umap>>
```

### 速率图 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-velo-graph-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-velo-graph-umap>>
```

### PAGA 图 {.tabset}

```{python}
<<run-paga>>
```

#### t-SNE

```{python fig.width=7, fig.height=7}
<<plot-paga-tsne>>
```

#### UMAP

```{python fig.width=7, fig.height=7}
<<plot-paga-umap>>
```


## 3 TR4 样本分析

```{r}
sce_tr4 <- sce[, sce$treatment == "TR4" | sce$time == "0DPI"]

sce_tr4 <- sce_tr4[, sce_tr4$cluster_id %in% c(
  paste0("Me_", 1:6), paste0("Va_", 1:4),
  "MPV_1", "MPV_2", "BS")]

embed_umap <- reducedDim(sce_tr4, "UMAP")

sce_tr4 <- sce_tr4[,
  embed_umap[, 1L] < 8
    & embed_umap[, 1L] > -8
    & embed_umap[, 2L] < 8
]

sce_tr4$cluster_id <- droplevels(sce_tr4$cluster_id)
sce_tr4$cellType <- droplevels(sce_tr4$cellType)

sce_sub <- sce_tr4
```

### 降维可视化 {.tabset}

```{r}
<<run-dimred>>
```

#### t-SNE

```{r}
<<plot-dimred-tsne>>
```

#### UMAP

```{r}
<<plot-dimred-umap>>
```

#### PHATE

```{r}
<<plot-dimred-tsne>>
```

#### DiffusionMap

```{r}
<<plot-dimred-dc>>
```

### 三维可视化 {.tabset}

```{r}
<<cell-3d-colors>>
```

#### PHATE

```{r}
<<dimred-3d-phate>>
```

#### DiffusionMap

```{r}
<<dimred-3d-dc>>
```

### 运行算法

```{r}
adata <- cache_pickle(
  file = "unitvelo_adata_tr4.pickle",
  dir = "../results/ObjectCache/TrajectoryInference/",
  expr = runUniTVelo(
    sce_tr4,
    label_key = "cluster_id",
    use.dimred = "HARMONY",
    assay.X = "logcounts",
    GPU = 0L,
    config = list(
      R2_ADJUST = TRUE,
      IROOT = NULL,
      FIT_OPTION = "1",
      AGENES_R2 = 1L
    )
  )
)
```

```{r}
<<update-embeddings>>
```

```{python fig.height=4, fig.width=8}
<<plot-prop>>
```

### 速率流线 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-stream-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-stream-umap>>
```

### 速率网格 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-grid-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-grid-umap>>
```

### 速率细胞 {.tabset}

#### t-SNE

```{python fig.height=12, fig.width=12}
<<plot-velo-tsne>>
```

#### UMAP

```{python fig.height=12, fig.width=12}
<<plot-velo-umap>>
```

### 隐时间 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-latent-time-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-latent-time-umap>>
```

### 速度与连贯性 {.tabset}

```{r}
<<run-velocity-confidence>>
```

#### t-SNE

```{r fig.height=7, fig.width=14}
<<plot-confidence-tsne>>
```

#### UMAP

```{r fig.height=7, fig.width=14}
<<plot-confidence-umap>>
```

### 速率图 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-velo-graph-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-velo-graph-umap>>
```

### PAGA 图 {.tabset}

```{python}
<<run-paga>>
```

#### t-SNE

```{python fig.width=7, fig.height=7}
<<plot-paga-tsne>>
```

#### UMAP

```{python fig.width=7, fig.height=7}
<<plot-paga-umap>>
```

## 4 1DPI 样本分析

```{r}
sce_1dpi <- sce[, sce$time == "1DPI"]

sce_1dpi <- sce_1dpi[, sce_1dpi$cluster_id %in% c(
  paste0("Me_", 1:6), paste0("Va_", 1:4),
  "MPV_1", "MPV_2", "BS")]

embed_umap <- reducedDim(sce_1dpi, "UMAP")

sce_1dpi <- sce_1dpi[,
  embed_umap[, 1L] < 8
    & embed_umap[, 1L] > -8
    & embed_umap[, 2L] < 8
]

sce_1dpi$cluster_id <- droplevels(sce_1dpi$cluster_id)
sce_1dpi$cellType <- droplevels(sce_1dpi$cellType)

sce_sub <- sce_1dpi
```

### 降维可视化 {.tabset}

```{r}
<<run-dimred>>
```

#### t-SNE

```{r}
<<plot-dimred-tsne>>
```

#### UMAP

```{r}
<<plot-dimred-umap>>
```

#### PHATE

```{r}
<<plot-dimred-tsne>>
```

#### DiffusionMap

```{r}
<<plot-dimred-dc>>
```

### 三维可视化 {.tabset}

```{r}
<<cell-3d-colors>>
```

#### PHATE

```{r}
<<dimred-3d-phate>>
```

#### DiffusionMap

```{r}
<<dimred-3d-dc>>
```

### 运行算法

```{r}
adata <- cache_pickle(
  file = "unitvelo_adata_1dpi.pickle",
  dir = "../results/ObjectCache/TrajectoryInference/",
  expr = runUniTVelo(
    sce_1dpi,
    label_key = "cluster_id",
    use.dimred = "HARMONY",
    assay.X = "logcounts",
    GPU = 0L,
    config = list(
      R2_ADJUST = TRUE,
      IROOT = NULL,
      FIT_OPTION = "1",
      AGENES_R2 = 1L
    )
  )
)
```

```{r}
<<update-embeddings>>
```

```{python fig.height=4, fig.width=8}
<<plot-prop>>
```

### 速率流线 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-stream-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-stream-umap>>
```

### 速率网格 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-grid-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-grid-umap>>
```

### 速率细胞 {.tabset}

#### t-SNE

```{python fig.height=12, fig.width=12}
<<plot-velo-tsne>>
```

#### UMAP

```{python fig.height=12, fig.width=12}
<<plot-velo-umap>>
```

### 隐时间 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-latent-time-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-latent-time-umap>>
```

### 速度与连贯性 {.tabset}

```{r}
<<run-velocity-confidence>>
```

#### t-SNE

```{r fig.height=7, fig.width=14}
<<plot-confidence-tsne>>
```

#### UMAP

```{r fig.height=7, fig.width=14}
<<plot-confidence-umap>>
```

### 速率图 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-velo-graph-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-velo-graph-umap>>
```

### PAGA 图 {.tabset}

```{python}
<<run-paga>>
```

#### t-SNE

```{python fig.width=7, fig.height=7}
<<plot-paga-tsne>>
```

#### UMAP

```{python fig.width=7, fig.height=7}
<<plot-paga-umap>>
```

## 5 2DPI 样本分析

```{r}
sce_2dpi <- sce[, sce$time == "2DPI"]

sce_2dpi <- sce_2dpi[, sce_2dpi$cluster_id %in% c(
  paste0("Me_", 1:6), paste0("Va_", 1:4),
  "MPV_1", "MPV_2", "BS")]

embed_umap <- reducedDim(sce_2dpi, "UMAP")

sce_2dpi <- sce_2dpi[,
  embed_umap[, 1L] < 8
    & embed_umap[, 1L] > -8
    & embed_umap[, 2L] < 8
]

sce_2dpi$cluster_id <- droplevels(sce_2dpi$cluster_id)
sce_2dpi$cellType <- droplevels(sce_2dpi$cellType)

sce_sub <- sce_2dpi
```

### 降维可视化 {.tabset}

```{r}
<<run-dimred>>
```

#### t-SNE

```{r}
<<plot-dimred-tsne>>
```

#### UMAP

```{r}
<<plot-dimred-umap>>
```

#### PHATE

```{r}
<<plot-dimred-tsne>>
```

#### DiffusionMap

```{r}
<<plot-dimred-dc>>
```

### 三维可视化 {.tabset}

```{r}
<<cell-3d-colors>>
```

#### PHATE

```{r}
<<dimred-3d-phate>>
```

#### DiffusionMap

```{r}
<<dimred-3d-dc>>
```

### 运行算法

```{r}
adata <- cache_pickle(
  file = "unitvelo_adata_2dpi.pickle",
  dir = "../results/ObjectCache/TrajectoryInference/",
  expr = runUniTVelo(
    sce_2dpi,
    label_key = "cluster_id",
    use.dimred = "HARMONY",
    assay.X = "logcounts",
    GPU = 0L,
    config = list(
      R2_ADJUST = TRUE,
      IROOT = NULL,
      FIT_OPTION = "1",
      AGENES_R2 = 1L
    )
  )
)
```

```{r}
<<update-embeddings>>
```

```{python fig.height=4, fig.width=8}
<<plot-prop>>
```

### 速率流线 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-stream-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-stream-umap>>
```

### 速率网格 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-grid-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-grid-umap>>
```

### 速率细胞 {.tabset}

#### t-SNE

```{python fig.height=12, fig.width=12}
<<plot-velo-tsne>>
```

#### UMAP

```{python fig.height=12, fig.width=12}
<<plot-velo-umap>>
```

### 隐时间 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-latent-time-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-latent-time-umap>>
```

### 速度与连贯性 {.tabset}

```{r}
<<run-velocity-confidence>>
```

#### t-SNE

```{r fig.height=7, fig.width=14}
<<plot-confidence-tsne>>
```

#### UMAP

```{r fig.height=7, fig.width=14}
<<plot-confidence-umap>>
```

### 速率图 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-velo-graph-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-velo-graph-umap>>
```

### PAGA 图 {.tabset}

```{python}
<<run-paga>>
```

#### t-SNE

```{python fig.width=7, fig.height=7}
<<plot-paga-tsne>>
```

#### UMAP

```{python fig.width=7, fig.height=7}
<<plot-paga-umap>>
```

## 6 3DPI 样本分析

```{r}
sce_3dpi <- sce[, sce$time == "3DPI"]

sce_3dpi <- sce_3dpi[, sce_3dpi$cluster_id %in% c(
  paste0("Me_", 1:6), paste0("Va_", 1:4),
  "MPV_1", "MPV_2", "BS")]

embed_umap <- reducedDim(sce_3dpi, "UMAP")

sce_3dpi <- sce_3dpi[,
  embed_umap[, 1L] < 8
    & embed_umap[, 1L] > -8
    & embed_umap[, 2L] < 8
]

sce_3dpi$cluster_id <- droplevels(sce_3dpi$cluster_id)
sce_3dpi$cellType <- droplevels(sce_3dpi$cellType)

sce_sub <- sce_3dpi
```

### 降维可视化 {.tabset}

```{r}
<<run-dimred>>
```

#### t-SNE

```{r}
<<plot-dimred-tsne>>
```

#### UMAP

```{r}
<<plot-dimred-umap>>
```

#### PHATE

```{r}
<<plot-dimred-tsne>>
```

#### DiffusionMap

```{r}
<<plot-dimred-dc>>
```

### 三维可视化 {.tabset}

```{r}
<<cell-3d-colors>>
```

#### PHATE

```{r}
<<dimred-3d-phate>>
```

#### DiffusionMap

```{r}
<<dimred-3d-dc>>
```

### 运行算法

```{r}
adata <- cache_pickle(
  file = "unitvelo_adata_3dpi.pickle",
  dir = "../results/ObjectCache/TrajectoryInference/",
  expr = runUniTVelo(
    sce_3dpi,
    label_key = "cluster_id",
    use.dimred = "HARMONY",
    assay.X = "logcounts",
    GPU = 0L,
    config = list(
      R2_ADJUST = TRUE,
      IROOT = NULL,
      FIT_OPTION = "1",
      AGENES_R2 = 1L
    )
  )
)
```

```{r}
<<update-embeddings>>
```

```{python fig.height=4, fig.width=8}
<<plot-prop>>
```

### 速率流线 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-stream-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-stream-umap>>
```

### 速率网格 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-grid-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-grid-umap>>
```

### 速率细胞 {.tabset}

#### t-SNE

```{python fig.height=12, fig.width=12}
<<plot-velo-tsne>>
```

#### UMAP

```{python fig.height=12, fig.width=12}
<<plot-velo-umap>>
```

### 隐时间 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-latent-time-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-latent-time-umap>>
```

### 速度与连贯性 {.tabset}

```{r}
<<run-velocity-confidence>>
```

#### t-SNE

```{r fig.height=7, fig.width=14}
<<plot-confidence-tsne>>
```

#### UMAP

```{r fig.height=7, fig.width=14}
<<plot-confidence-umap>>
```

### 速率图 {.tabset}

#### t-SNE

```{python fig.height=7, fig.width=7}
<<plot-velo-graph-tsne>>
```

#### UMAP

```{python fig.height=7, fig.width=7}
<<plot-velo-graph-umap>>
```

### PAGA 图 {.tabset}

```{python}
<<run-paga>>
```

#### t-SNE

```{python fig.width=7, fig.height=7}
<<plot-paga-tsne>>
```

#### UMAP

```{python fig.width=7, fig.height=7}
<<plot-paga-umap>>
```


