---
title: "Sample Correlation Analysis"
output: html_document
---

```{r, setup, include=FALSE}
library(magrittr)
library(patchwork)
knitr::opts_chunk$set(echo = TRUE)
source("../scripts/LoadUtils.R", chdir = TRUE)
```

## Define Helper Functions

```{r}
make_sample_pseudobulk <- function(seurat,
    type = c("counts", "logcounts", "cpm", "vstresiduals"),
    fun = NULL) {
  sce <- Seurat::as.SingleCellExperiment(seurat, assay = "RNA")
  sce <- muscat::prepSCE(
    sce,
    kid = "ident", # subpopulation assignments
    gid = "group", # group IDs (ctrl/stim)
    sid = "sample", # sample IDs (ctrl/stim.1234)
    drop = FALSE
  )

  sce <- sce[Matrix::rowSums(SingleCellExperiment::counts(sce) > 0) > 0, ]

  type <- match.arg(type)

  pb <- rhapsodykit::calculate_pseudo_bulk(
    sce, type, fun = fun, by = "sample_id")

  pb@assays@data[[1]]
}

sample_correlation_heatmap <- function(
  cormat,
  matrix_name = NULL,
  label = FALSE,
  ...
) {
  p <- ComplexHeatmap::Heatmap(
    cormat,
    name = matrix_name,
    row_names_side = "left",
    row_dend_side = "left",
    column_names_side = "bottom",
    column_dend_side = "top",
    column_names_rot = 45,
    show_column_dend = FALSE,
    clustering_distance_rows = \(x) as.dist(sqrt(1 - x^2)),
    clustering_distance_columns = \(x) as.dist(sqrt(1 - x^2)),
    cell_fun = if (label) function(j, i, x, y, width, height, fill) {
      grid::grid.text(
        sprintf("%.2f", cormat[i, j]),
        x, y, gp = grid::gpar(fontsize = 14)
      )
    } else NULL,

    ...
  )

  p
}

plot_correlation_heatmap <- function(cormat) {
  p <- sample_correlation_heatmap(
      cormat,
      col = viridis::viridis(256),
      row_dend_width = grid::unit(30, "mm"),
      heatmap_legend_param = list(
        title = "Pearson's R",
        title_position = "leftcenter-rot",
        legend_direction = "vertical",
        legend_height = grid::unit(4, "cm")),
      left_annotation = ComplexHeatmap::rowAnnotation(
        group = c(rep("0DPI", 2), rep("1DPI", 6),
                  rep("2DPI", 6), rep("3DPI", 6)),
        show_annotation_name = FALSE,
        annotation_legend_param = list(
          title = "Time",
          title_position = "leftcenter-rot"),
        col = list(
          group = c("0DPI" = "#ffffd4", "1DPI" = "#fed98e",
                    "2DPI" = "#fe9929", "3DPI" = "#cc4c02")))
      )

  #p <- ComplexHeatmap::draw(p, heatmap_legend_side = "right")

  p
}

#' Calculate latent space of DensityMorph
#'
#' @param Xref When calculating self-PDLR reference, if `npc` and `min_cells`
#' is same, we can just use the pre-calculated `Xref`.
calculateDensityMorph <- function(
    seurat, reduction = "pca", Xref = NULL, npc = 10) {
  # Get cell embeddings of PCA
  expr_data <- Seurat::Embeddings(seurat[[reduction]])[, 1:npc]

  # Remove duplicated cells
  keep_cells <- !duplicated(expr_data, MARGIN = 1)
  expr_data <- expr_data[keep_cells,]

  mat_split <- function(x, f) {
    stopifnot(nrow(x) == length(f))
    lapply(split(x, f), matrix, ncol = ncol(x))
  }
  expr_list <- mat_split(expr_data, seurat$sample[keep_cells])

  # Remove duplicated cells
  expr_list <- lapply(expr_list, function(x) {
    x[!duplicated(x, MARGIN = 1),]
  })

  # Calculate self-PDLR reference
  if (is.null(Xref)) {
    min_cells <- min(sapply(expr_list, nrow))
    message("Calculating Xref with npc = ", npc, ", min_cells = ", min_cells)
    Xref <- DensityMorph::reference_distribution(
      npc, min_cells*100, min_cells)
  }

  # Subsampling
  expr_list_down <- lapply(expr_list, function(x) {
    idx <- sample(1:nrow(x), min_cells, replace = FALSE)
    x[idx,]
  })

  # Calculate latent space
  LS <- DensityMorph::latent_space(expr_list_down, Xref)
  rownames(LS$latent_space) <- names(expr_list_down)

  list(LS = LS, Xref = Xref)
}

plot_density_morph <- function(morph) {
  sample_info <- stringr::str_match(rownames(morph$LS$latent_space), "^(.*)-(.*)-(.*)$")
  
  data_to_plot <- morph$LS$latent_space[, 1:2]
  data_to_plot$Sample <- sample_info[, 1]
  data_to_plot$Time <- sample_info[, 2]
  data_to_plot$Treatment <- sample_info[, 3]
  
  p <- ggplot2::ggplot(data_to_plot, ggplot2::aes(
      x = A1, y = A2,
      fill = Treatment, color = Treatment,
      shape = Time, label = Sample
    )) +
    ggplot2::geom_point(size = 3) +
    ggplot2::scale_color_brewer(type = "qual", palette = "Dark2") +
    ggplot2::scale_fill_brewer(type = "qual", palette = "Dark2") +
    ggplot2::scale_shape_manual(values = c(
      "0DPI" = 22, "1DPI" = 21, "2DPI" = 23, "3DPI" = 24)) +
    ggplot2::xlab("LSC1") +
    ggplot2::ylab("LSC2") +
    ggplot2::coord_fixed() +
    ggplot2::theme_bw() +
    NULL
  
  p
}
```

## Analysis of Raw Data

这部分数据只经历过 BD Rhapsody 官方预处理流程的二阶导数分析，移除掉了空磁珠；没有经过进一步的质量控制和批次效应移除。

### Read raw data

```{r}
samples <- c(
  "0DPI-MOCK-1",
  "0DPI-MOCK-2",

  "1DPI-MOCK-1",
  "1DPI-MOCK-2",
  "1DPI-PNR2-1",
  "1DPI-PNR2-2",
  "1DPI-TR4-1",
  "1DPI-TR4-2",

  "2DPI-MOCK-1",
  "2DPI-MOCK-2",
  "2DPI-PNR2-1",
  "2DPI-PNR2-2",
  "2DPI-TR4-1",
  "2DPI-TR4-2",

  "3DPI-MOCK-1",
  "3DPI-MOCK-2",
  "3DPI-PNR2-1",
  "3DPI-PNR2-3",
  "3DPI-TR4-1",
  "3DPI-TR4-2"
)

data_folders <- structure(
  paste0("../results/RawOutput/", samples), names = samples)

obj_list <- lapply(data_folders, function(base_dir) {
  xfun::cache_rds(
    file = paste0("obj_raw_", basename(base_dir)),
    dir = "../results/ObjectCache/QualityControl/",
    expr = local({
      expr_matrix <- rhapsodykit::read_rhapsody_wta(base_dir, TRUE)
      seurat_obj <- Seurat::CreateSeuratObject(
        counts = expr_matrix, project = basename(base_dir))
      seurat_obj$sample <- basename(base_dir)
      seurat_obj$group <- stringr::str_extract(seurat_obj$sample, "\\dDPI-(MOCK|PNR2|TR4)")
      seurat_obj$time <- stringr::str_extract(seurat_obj$sample, "\\dDPI")
      seurat_obj$treatment <- stringr::str_extract(seurat_obj$sample, "MOCK|PNR2|TR4")
      seurat_obj
    })
  )
 })

obj_merged <- merge(
  x = obj_list[[1]],
  y = obj_list[-1],
  add.cell.ids = names(obj_list)
)
```

### Sample correlation analysis

Pseudobulk 基于归一化后的表达值聚合而成，没有经历质量控制。

```{r}
pseudobulk_raw <- make_sample_pseudobulk(obj_merged, type = "logcounts", fun = "mean")
```

```{r}
cormat <- cor(pseudobulk_raw, method = "pearson", use = "complete.obs")
```

```{r}
p1 <- plot_correlation_heatmap(cormat)
p1
```

### Sample density morph analysis

```{r}
obj_merged <- obj_merged |>
  Seurat::NormalizeData(verbose = FALSE) |>
  Seurat::FindVariableFeatures(selection.method = "vst", nfeatures = 2000) |>
  Seurat::ScaleData(verbose = FALSE) |>
  Seurat::RunPCA(features = NULL, npcs = 30, verbose = FALSE)
```

```{r}
morph <- calculateDensityMorph(obj_merged, Xref = NULL)
```

```{r}
sample_info <- stringr::str_match(rownames(morph$LS$latent_space), "^(.*)-(.*)-(.*)$")

data_to_plot <- morph$LS$latent_space[, 1:2]
data_to_plot$Sample <- sample_info[, 1]
data_to_plot$Time <- sample_info[, 2]
data_to_plot$Treatment <- sample_info[, 3]

p2 <- ggplot2::ggplot(data_to_plot, ggplot2::aes(
    x = A1, y = A2,
    fill = Treatment, color = Treatment,
    shape = Time, label = Sample
  )) +
  ggplot2::geom_point(size = 3) +
  ggplot2::scale_color_brewer(type = "qual", palette = "Dark2") +
  ggplot2::scale_fill_brewer(type = "qual", palette = "Dark2") +
  ggplot2::scale_shape_manual(values = c(
    "0DPI" = 22, "1DPI" = 21, "2DPI" = 23, "3DPI" = 24)) +
  ggplot2::xlab("LSC1") +
  ggplot2::ylab("LSC2") +
  ggplot2::coord_fixed() +
  ggplot2::theme_bw() +
  NULL

p2
```

### Combine plots

```{r fig.height=6, fig.width=14}
(p <- ggplotify::as.ggplot(p1) + p2 +
   patchwork::plot_layout(widths = c(2, 1)) +
   patchwork::plot_annotation(tag_levels = "A") &
   font_plot_tag(16))
```

```{r}
rm(obj_list, obj_merged)
gc()
```

## Analysis of Quality Controlled Data

### Read quality controlled data

```{r}
obj_list <- lapply(samples, function(sample) readRDS(
  Sys.glob(paste0(
    "../results/ObjectCache/QualityControl/",
    "obj_strained_", sample, "_*.rds"))
  )
)
```

```{r}
obj_merged <- merge(
  x = obj_list[[1]],
  y = obj_list[-1],
  add.cell.ids = names(obj_list)
)
```

```{r}
head(obj_merged@assays$RNA@counts)
```

### Sample correlation analsysis

```{r}
pseudobulk_qc <- make_sample_pseudobulk(obj_merged, type = "logcounts", fun = "mean")
head(pseudobulk_qc)
```

```{r}
cormat <- cor(pseudobulk_qc, method = "pearson", use = "complete.obs")
```

```{r}
p1 <- plot_correlation_heatmap(cormat)
p1
```

### Sample density morph analysis

```{r}
obj_merged <- obj_merged |>
  Seurat::NormalizeData(verbose = FALSE) |>
  Seurat::FindVariableFeatures(selection.method = "vst", nfeatures = 2000) |>
  Seurat::ScaleData(verbose = FALSE) |>
  Seurat::RunPCA(features = NULL, npcs = 30, verbose = FALSE)
```

```{r}
morph <- calculateDensityMorph(obj_merged, Xref = NULL)
```


```{r}
p2 <- plot_density_morph(morph)
p2
```

### Combine plots

```{r fig.height=6, fig.width=14}
(p <- ggplotify::as.ggplot(p1) + p2 +
   patchwork::plot_layout(widths = c(2, 1)) +
   patchwork::plot_annotation(tag_levels = "A") &
   font_plot_tag(16))
```

```{r}
rm(obj_list, obj_merged)
gc()
```

## Analysis of Batch-effect Corrected Data

### Read integrated data

```{r}
obj_int <- readRDS(Sys.glob(
  "../results/ObjectCache/IntegrationAndAnnotation/obj_annotated_*.rds"))
```

### Sample correlation analsysis

```{r}
pseudobulk_int <- make_sample_pseudobulk(obj_int, type = "logcounts", fun = "mean")
head(pseudobulk_int)
```

```{r}
cormat <- cor(pseudobulk_int, method = "pearson", use = "complete.obs")
```

```{r}
p1 <- plot_correlation_heatmap(cormat)
p1
```

### Sample density morph analysis

```{r}
morph <- calculateDensityMorph(obj_int, reduction = "harmony")
```

```{r}
p2 <- plot_density_morph(morph)
p2
```

### Combine plots

```{r fig.height=6, fig.width=14}
(p <- ggplotify::as.ggplot(p1) + p2 +
   patchwork::plot_layout(widths = c(2, 1)) +
   patchwork::plot_annotation(tag_levels = "A") &
   font_plot_tag(16))
```

```{r}
rm(obj_list, obj_int)
gc()
```
