---
title: "Slides for 2023SSPP Presentation"
author: "Altair Wei"
date: "2023-05-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggtree, include.only = "%<+%")
source("../scripts/LoadUtils.R", chdir = TRUE)
```

## Load Data

### MOCK samples

```{r}
obj_mock <- readRDS(Sys.glob("../results/ObjectCache/IntegrationAndAnnotation/obj_mock_annotated_*.rds"))
sce_mock <- Seurat::as.SingleCellExperiment(obj_mock, assay = "RNA")
sce_mock <- muscat::prepSCE(sce_mock,
  kid = "ident", # subpopulation assignments
  gid = "group", # group IDs (ctrl/stim)
  sid = "sample", # sample IDs (ctrl/stim.1234)
  drop = FALSE
)
```

```{r}
mock_celltypes <- c(
  "Gu", "Ep_α", "Ep_β",
  "Me_α", "Me_β", "Me_γ", "Me_δ", "Me_ε",
  "BS", "CC", "Va_α", "Va_β", "Va_γ", "Va_δ", "MPV")

mock_celltype_names <- c(
  "Guard cell", "Epidermal α", "Epidermal β",
  
  "Mesophyll α", "Mesophyll β", "Mesophyll γ",
  "Mesophyll δ", "Mesophyll ε",
  
  "Vascular bundle sheath", "Phloem companion cell",
  "Vascular α", "Vascular β", "Vascular γ",
  "Vascular δ", "Midvein provascular cell"
)

mock_celltype_colors <- ggthemes::tableau_color_pal("Tableau 20")(length(mock_celltypes))
#mock_celltype_colors <- scater:::.get_palette("tableau20")[seq_along(mock_celltypes)]
names(mock_celltype_colors) <- mock_celltypes
```

```{r}
mock_body_layers <- c(
  "Gu"   = "L1",
  "Ep_α" = "L1",
  "Ep_β" = "L1",
  "Me_α" = "L2",
  "Me_β" = "L2",
  "Me_γ" = "L2",
  "Me_δ" = "L2",
  "Me_ε" = "L2",
  "BS"   = "L3",
  "CC"   = "L3",
  "Va_α" = "L3",
  "Va_β" = "L3",
  "Va_γ" = "L3",
  "Va_δ" = "L3",
  "MPV"  = "L3"
)
```

```{r}
sce_mock$cluster_id <- forcats::fct_relevel(sce_mock$cluster_id, mock_celltypes)
sce_mock$body_layer <- do.call(
  dplyr::recode, c(list(.x = sce_mock$cluster_id), as.list(mock_body_layers)))
Seurat::Idents(obj_mock) <- forcats::fct_relevel(Seurat::Idents(obj_mock), mock_celltypes)
```

### All samples

```{r}
obj <- readRDS(Sys.glob("../results/ObjectCache/IntegrationAndAnnotation/obj_annotated_*.rds"))
sce <- Seurat::as.SingleCellExperiment(obj, assay = "RNA")
sce <- muscat::prepSCE(sce,
  kid = "ident", # subpopulation assignments
  gid = "group", # group IDs (ctrl/stim)
  sid = "sample", # sample IDs (ctrl/stim.1234)
  drop = FALSE
)
```

```{r}
all_celltypes <- c(
  "Gu", "Ep_1", "Ep_2",
  "Me_1", "Me_2", "Me_3", "Me_4", "Me_5", "Me_6",
  "Va_1", "Va_2", "Va_3", "Va_4",
  "BS", "CC", "MPV_1", "MPV_2")

all_celltype_colors <- ggthemes::tableau_color_pal("Tableau 20")(length(all_celltypes))
names(all_celltype_colors) <- all_celltypes
```

```{r}
all_body_layers <- c(
  "Gu"    = "L1",
  "Ep_1"  = "L1",
  "Ep_2"  = "L1",
  "Me_1"  = "L2",
  "Me_2"  = "L2",
  "Me_3"  = "L2",
  "Me_4"  = "L2",
  "Me_5"  = "L2",
  "Me_6"  = "L2",
  "Va_1"  = "L2",
  "Va_2"  = "L2",
  "Va_3"  = "L2",
  "Va_4"  = "L2",
  "BS"    = "L3",
  "CC"    = "L3",
  "MPV_1" = "L3",
  "MPV_2" = "L3"
)

all_tissues <- c(
  Gu   = "Stomata",
  Ep_1 = "Epidermis",
  Ep_2 = "Epidermis",
  Me_1 = "Chlorenchyma",
  Me_2 = "Chlorenchyma",
  Me_3 = "Chlorenchyma",
  Me_4 = "Parenchyma",
  Me_5 = "Parenchyma",
  Me_6 = "Parenchyma",
  Va_1 = "Chlorenchyma",
  Va_2 = "Outer sheath",
  Va_3 = "Outer sheath",
  Va_4 = "Outer sheath",
  BS = "Inner sheath",
  CC = "Phloem",
  MPV_1 = "Procambium",
  MPV_2 = "Procambium"
)
```

```{r}
tissue_types <- c(
  "Stomata",
  "Epidermis",
  "Chlorenchyma",
  "Parenchyma",
  "Outer sheath",
  "Inner sheath",
  "Phloem",
  "Procambium"
)

all_tissue_colors <- ggthemes::tableau_color_pal("Tableau 10")(length(tissue_types))
names(all_tissue_colors) <- tissue_types
```

```{r}
sce$cluster_id <- forcats::fct_relevel(sce$cluster_id, all_celltypes)
Seurat::Idents(obj) <- forcats::fct_relevel(Seurat::Idents(obj), all_celltypes)
sce$body_layer <- dplyr::recode(sce$cluster_id, !!!all_body_layers)
sce$tissue <- factor(dplyr::recode(sce$cluster_id, !!!all_tissues), levels = tissue_types)
```

### Gene List

```{r}
filelist <- Sys.glob("../data/pathways/*.txt")
names(filelist) <- basename(filelist) |> stringr::str_remove(stringr::fixed(".txt"))

pathways <- lapply(filelist, function(filename) {
  genes <- stringr::str_trim(readLines(filename))
  genes[genes != ""]
})
```

### DS Results

```{r}
logfc_data <- readRDS(Sys.glob(
  "../results/ObjectCache/DifferentialState/ds_deg_logfc_data_*.rds"))
```

```{r}
ds_res <- readRDS(Sys.glob("../results/ObjectCache/DifferentialState/ds_res_*.rds"))
```

```{r}
ds_sig <- ds_res$DESeq2 |>
  rhapsodykit::diff_state_significant(fdr_limit = 0.05, logfc_limit = 0)
```

```{r}
df_degs <- rhapsodykit::diff_state_format(ds_sig)

matched <- stringr::str_match(
  df_degs$contrast, "X(\\dDPI)\\.(MOCK|PNR2|TR4)-X(\\dDPI)\\.(MOCK|PNR2|TR4)")

df_degs$time <- factor(matched[, 2], levels = c("0DPI", "1DPI", "2DPI", "3DPI"))
df_degs$contrast <- paste(matched[, 3], matched[, 5], sep = " vs. ")
df_degs$regu_type <- factor(ifelse(df_degs$logFC > 0, "UP", "DOWN"), levels = c("UP", "DOWN"))

df_degs <- df_degs |>
  dplyr::select(time, contrast, cluster_id, regu_type, gene) |>
  dplyr::filter(contrast != "PNR2 vs. TR4")
```

## Cell Type Annotation

### Cell Sample Groups

```{r}
Seurat::DimPlot(obj, reduction = "umap", group.by = "group", shuffle = TRUE) +
  ggplot2::coord_fixed() +
  theme_dimred() +
  ggplot2::theme(plot.title = ggplot2::element_blank()) +
  ggplot2::labs(color = "Sample Group")
```

### Known Marker Genes

```{r fig.height=12, fig.width=8}
p <- Seurat::DotPlot(obj_mock, features = KNOWN_MARKERS) +
  ggplot2::coord_flip() +
  ggplot2::facet_grid(
    facets = feature.groups ~ .,
    scales = "free_y",
    space = "free_y",
    switch = "y") +
  ggplot2::theme(
    axis.text.y = ggplot2::element_blank(),
    axis.text.x = ggplot2::element_text(angle = 45, hjust = 1),
    strip.text.y.left = ggplot2::element_text(angle = 0),
    strip.placement = "outside"
  )

p
```

### Cell Allocation in LCM-seq

```{r}
marker_cosg <- COSG::cosg(
 obj_mock,
 groups = "all",
 assay = "RNA",
 slot = "data",
 mu = 1,
 n_genes_user = 300
)
```

```{r}
LCM_design <- c(
  "VC_1" = "vb.",
  "VC_2" = "vb.",
  "VC_3" = "vb.",
  "CC_1" = "bp.",
  "CC_2" = "bp.",
  "CC_3" = "bp.",
  "MC3_1" = "dp.",
  "MC3_2" = "dp.",
  "MC3_3" = "dp.",
  "MC_1" = "me.",
  "MC_2" = "me.",
  "MC_3" = "me.",
  "EC_1" = "ep.",
  "EC_2" = "ep.",
  "EC_3" = "ep."
)

CIBER <- deconvLCM(
  seurat = obj_mock,
  lcm_file = "../results/LCMSeq/gene_quanti/counts/all.featureCounts",
  markers = unique(unlist(marker_cosg$names)),
  design = LCM_design,
  mc.cores = 4
)
```

```{r fig.height=7, fig.width=5}
p <- deconvScatter(CIBER, LCM_design, ncol = 3, theme_size = 14) +
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(angle = 0, hjust = 0.5, vjust = 0.5),
    axis.text.y = ggplot2::element_blank(),
    axis.ticks.y = ggplot2::element_blank())

p
```

## Fungal Invasion

### Visualize Number of DEGs

```{r}
genecount_df <- rhapsodykit::diff_state_format(ds_sig)

genecount_df$cluster_id <- factor(genecount_df$cluster_id, unique(all_tissues))
genecount_df$time <- stringr::str_extract(genecount_df$contrast, "\\dDPI")
genecount_df$contrast <- factor(genecount_df$contrast, levels = CONTRAST_NAMES)
genecount_df$regu_type <- ifelse(genecount_df$logFC > 0, "Up-regulated", "Down-regulated")

df_to_plot <- genecount_df |>
  dplyr::group_by(time, contrast, cluster_id, regu_type) |>
  dplyr::count() |>
  dplyr::mutate(n = ifelse(regu_type == "Up-regulated", n, -1 * n)) |>
  tidyr::pivot_wider(names_from = "cluster_id", values_from = "n", values_fill = NA) |>
  tidyr::pivot_longer(cols = -c(time, contrast, regu_type), names_to = "cluster_id", values_to = "n") |>
  dplyr::mutate(contrast = stringr::str_replace(contrast,
      "X(\\dDPI)\\.(.*)-X(\\dDPI)\\.(.*)", "\\2 vs. \\4"))

df_to_plot$cluster_id <- factor(df_to_plot$cluster_id, unique(all_tissues))
df_to_plot$contrast <- factor(df_to_plot$contrast, c("PNR2 vs. MOCK", "TR4 vs. MOCK", "PNR2 vs. TR4"))

p1 <- ggplot2::ggplot(
    data = df_to_plot,
    mapping = ggplot2::aes(x = cluster_id, y = n, fill = contrast)) +
  ggplot2::geom_bar(
    data = subset(df_to_plot, regu_type == "Up-regulated"),
    stat = "identity",
    position = ggplot2::position_dodge2(preserve = "single", padding = 0)) +
  ggplot2::geom_bar(
    data = subset(df_to_plot, regu_type == "Down-regulated"),
    stat = "identity", alpha = 0.5,
    position = ggplot2::position_dodge2(preserve = "single", padding = 0)) +
  ggplot2::scale_fill_brewer(palette = "Set1") +
  ggplot2::ylab("Number of DEGs") +
  ggplot2::xlab("Cell Types") +
  ggplot2::guides(fill = ggplot2::guide_legend(title = "Contrast")) +
  ggplot2::theme_bw(base_size = 16) +
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(angle = 45, hjust = 1),
    legend.position = "right",
    strip.background.x = ggplot2::element_blank(),
    strip.text.x = ggplot2::element_blank()) +
  ggplot2::facet_grid(time ~ cluster_id, scales = "free_x") +
  NULL
```

```{r}
df_degs <- rhapsodykit::diff_state_format(ds_sig)

matched <- stringr::str_match(
  df_degs$contrast, "X(\\dDPI)\\.(MOCK|PNR2|TR4)-X(\\dDPI)\\.(MOCK|PNR2|TR4)")

df_degs$time <- factor(matched[, 2], levels = c("0DPI", "1DPI", "2DPI", "3DPI"))
df_degs$contrast <- paste(matched[, 3], matched[, 5], sep = " vs. ")
df_degs$regu_type <- factor(ifelse(df_degs$logFC > 0, "UP", "DOWN"), levels = c("UP", "DOWN"))

df_degs <- df_degs |>
  dplyr::select(time, contrast, cluster_id, regu_type, gene) |>
  dplyr::filter(contrast != "PNR2 vs. TR4")

p2 <- df_degs |>
  dplyr::group_by(time, contrast, regu_type) |>
  dplyr::summarise(count = length(unique(gene)), .groups = "drop") |>
  dplyr::add_row(
    time = "0DPI",
    contrast = rep(c("PNR2 vs. MOCK", "TR4 vs. MOCK"), each = 2),
    regu_type = factor(rep(c("UP", "DOWN"), 2), levels = c("UP", "DOWN")),
    count = 0) |>
  dplyr::arrange(time) |>
  ggplot2::ggplot(ggplot2::aes(
    x = time, y = count,
    color = contrast, linetype = regu_type,
    group = paste(contrast, regu_type))) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::scale_color_brewer(palette = "Set1", name = "Contrast") +
  ggplot2::ylab("Number of DEGs") +
  ggplot2::xlab("Time") +
  ggplot2::guides(linetype = ggplot2::guide_legend(title = "Regu Type")) +
  ggplot2::theme_bw(base_size = 16)
```

```{r fig.height=6, fig.width=14}
(p <- p2 + ggplot2::theme(legend.position = c(0.22, 0.75),
                          panel.grid = ggplot2::element_blank()) +
      p1 + ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 25)))
```

### Immunity DEGs Heatmap

#### Define Pathways

```{r}
immunity_genes <- list(
  PTI = pathways[c("RLK", "MAPK", "RBOH")],
  ETI = pathways[c("NBLRR", "PR")],
  SAR = pathways[c("TaNPR", "NHP", "GOSAR")],
  Photosynthesis = pathways[c("PhotosynPSEA", "PhotosynAnt", "PhotosynCfix")],
  Metabolism = pathways[c(
    "Aromatic amino acid", "General phenylpropanoid", "Lignin biosynthesis",
    "TaGST", "Polyamine")],
  TFs = pathways[c("WRKY", "NAC", "Myb", "bZIP")]
)
```

```{r}
immunity_df <- immunity_genes |>
  tibble::enframe(name = "Category", value = "Data") |>
  dplyr::rowwise() |>
  dplyr::mutate(Data = list(tibble::enframe(
    Data, name = "Pathway", value = "Gene"))) |>
  tidyr::unnest(cols = Data) |>
  tidyr::unnest(cols = Gene) |>
  dplyr::mutate(
    Pathway = dplyr::recode(Pathway,
      RLK = "RLKs", MAPK = "MAPKs", RBOH = "RBOHs",
      NBLRR = "NBLRRs", PR = "PRs", TaNPR = "NPRs",
      NHP = "NHPs", "Aromatic amino acid" = "Aromatic",
      "General phenylpropanoid" = "Phenylpropanoid",
      TaGST = "GSTs", WRKY = "WRKYs", NAC = "NACs",
      Myb = "MYBs", bZIP = "bZIPs"))
```

#### Heatmap for Cell Types {.tabset}

```{r}
se <- SummarizedExperiment(
  assays = list(logFC = logfc_data$mtx),
  colData = logfc_data$coldata[-1])
se <- se[intersect(rownames(se), immunity_df$Gene), ]

# Duplicated genes existed
rowData(se) <- with(immunity_df, local({
  idx <- match(rownames(se), Gene)
  DataFrame(
    Pathway = factor(Pathway[idx], unique(Pathway)),
    Category = factor(Category[idx], unique(Category))
  )
}))
```

##### All cell types

```{r fig.height=10, fig.width=20}
plotImmunityHeatmap(
  se,
  row_split_by = "Category",
  row_anno_by = "Pathway",
  col_split_by = "cell_types",
  col_anno_by = c("contrast", "times"),
  clustering_method = "ward.D2",
  row_dend_width = grid::unit(40, "mm")
)
```

```{r fig.height=7, fig.width=14, eval=FALSE, include=FALSE}
p <- plotImmunityHeatmap(
  se[rowData(se)$Category != "TFs", ],
  row_split_by = "Pathway",
  show_row_dend = FALSE,
  row_title_side = "left",
  col_split_by = "cell_types",
  col_anno_by = c("contrast", "times"),
  clustering_method = "ward.D2",
  fontsize = 6,
  fontfamily = "Arial",
  simple_anno_size = grid::unit(2, "mm"),
  border = TRUE,
  border_gp = grid::gpar(lwd = 0.5),
  gap = grid::unit(0.5, "mm"),
  legend_grid_size = grid::unit(3, "mm")
)

p <- plotImmunityHeatmap(
  se[rowData(se)$Category == "TFs", ],
  row_split_by = "Pathway",
  show_row_dend = FALSE,
  row_title_side = "left",
  col_split_by = "cell_types",
  col_anno_by = c("contrast", "times"),
  clustering_method = "ward.D2",
  fontsize = 6,
  fontfamily = "Arial",
  simple_anno_size = grid::unit(2, "mm"),
  border = TRUE,
  border_gp = grid::gpar(lwd = 0.5),
  legend_grid_size = grid::unit(3, "mm")
)

plotpowerpoint(p, "../data/thesis.pptx", margins = c(top = 1.3, right = 1.3, bottom = 4, left = 1.3))
```

##### Chlorenchyma

```{r fig.height=12, fig.width=7}
plotImmunityHeatmap(
  se[, se$cell_types == "Chlorenchyma"],
  row_split_by = "Pathway",
  row_anno_by = "Pathway",
  col_split_by = "contrast",
  col_anno_by = c("contrast", "times"),
  clustering_method = "ward.D2")
```

##### Outer sheath

```{r fig.height=12, fig.width=7}
plotImmunityHeatmap(
  se[, se$cell_types == "Outer sheath"],
  row_split_by = "Pathway",
  row_anno_by = "Pathway",
  col_split_by = "contrast",
  col_anno_by = c("contrast", "times"),
  clustering_method = "ward.D2")
```

##### Chlorenchyma + Outer Sheath

```{r fig.height=10, fig.width=10, dev='svg'}
plotImmunityHeatmap(
  se[rowData(se)$Category != "TFs",
        se$cell_types %in% c("Chlorenchyma", "Outer sheath")],
  row_split_by = "Pathway",
  row_anno_by = "Pathway",
  col_split_by = "cell_types",
  col_anno_by = c("contrast", "times"),
  clustering_method = "ward.D2",
  fontsize = 16,
  border = TRUE
)
```

##### Parenchyma

```{r fig.height=12, fig.width=7}
plotImmunityHeatmap(
  se[, se$cell_types == "Parenchyma"],
  row_split_by = "Pathway",
  row_anno_by = "Pathway",
  col_split_by = "contrast",
  col_anno_by = c("contrast", "times"),
  clustering_method = "ward.D2")
```

##### Phloem

```{r fig.height=12, fig.width=7}
plotImmunityHeatmap(
  se[, se$cell_types == "Phloem"],
  row_split_by = "Pathway",
  row_anno_by = "Pathway",
  col_split_by = "contrast",
  col_anno_by = c("contrast", "times"),
  clustering_method = "ward.D2")
```

### Similarity of DEG Sets

#### Split by time {.tabset}

```{r}
df_jacmtx <- df_degs |>
  dplyr::select(-regu_type) |>
  dplyr::mutate(
    time = forcats::fct_drop(time),
    cluster_id = factor(cluster_id, levels = TISSUE_TYPES),
    contrast = factor(contrast, levels = c("PNR2 vs. MOCK", "TR4 vs. MOCK"))) |>
  dplyr::group_by(time) |>
  tidyr::nest() |>
  dplyr::mutate(
    jacmtx = lapply(data, function(x) {
      dplyr::group_by(x, contrast, cluster_id) |>
        dplyr::summarise(genes = list(gene), .groups = "drop") |>
        dplyr::rename(contrast.1 = contrast, cluster_id.1 = cluster_id) |>
        dplyr::mutate(data = lapply(
          X = seq_along(genes),
          FUN = function(idx) {
            data.frame(
              contrast.2 = contrast.1,
              cluster_id.2 = cluster_id.1,
              jaccard = vapply(
                X = genes,
                FUN = function(gs)
                  bayesbio::jaccardSets(genes[[idx]], gs),
                FUN.VALUE = numeric(1))
            )
          })) |>
        dplyr::select(contrast.1, cluster_id.1, data) |>
        tidyr::unnest(cols = data)
    }),
    anno = lapply(data, function(x) {
      dplyr::group_by(x, contrast, cluster_id) |>
        dplyr::summarise(genes = list(gene),
                         size = length(gene),
                         .groups = "drop") |>
        dplyr::mutate(
          label = paste(contrast, cluster_id, sep = " - "),
          .before = 1)
    })) |>
  dplyr::select(time, jacmtx, anno)
```

```{r results='asis'}
df_jacmtx |>
  purrr::pwalk(function(time, jacmtx, anno) {
    cat("#####", as.character(time), "\n\n")

    mtx <- jacmtx |>
      dplyr::mutate(
        set.1 = paste(contrast.1, cluster_id.1, sep = " - "),
        set.2 = paste(contrast.2, cluster_id.2, sep = " - ")) |>
      dplyr::select(set.1, set.2, jaccard) |>
      tidyr::pivot_wider(names_from = set.2, values_from = jaccard) |>
      tibble::column_to_rownames("set.1") |>
      as.matrix()

    jacdist <- as.dist(1 - mtx)

    p <- ggtree::ggtree(hclust(jacdist, method = "average")) %<+% anno +
      ggtree::geom_tippoint(ggplot2::aes(
        color = cluster_id, shape = contrast, size = size)) +
      ggtree::geom_tiplab(ggplot2::aes(
          label = cluster_id, color = cluster_id),
        offset = 0.01, show.legend = FALSE) +
      ggplot2::scale_x_continuous(expand = ggplot2::expansion(mult = c(0.05, 0.25))) +
      ggthemes::scale_color_tableau("Tableau 10")

    print(p)

    cat("\n\n")
  })
```

#### All in One Plot {.tabset}

```{r}
plist <- df_jacmtx |>
  purrr::pmap(function(time, jacmtx, anno) {
    mtx <- jacmtx |>
      dplyr::mutate(
        set.1 = paste(time, contrast.1, cluster_id.1, sep = " - "),
        set.2 = paste(time, contrast.2, cluster_id.2, sep = " - ")) |>
      dplyr::select(set.1, set.2, jaccard) |>
      tidyr::pivot_wider(names_from = set.2, values_from = jaccard) |>
      tibble::column_to_rownames("set.1") |>
      as.matrix()

    anno <- dplyr::mutate(anno, label = paste(time, label, sep = " - "))
    
    jacdist <- as.dist(1 - mtx)

    p <- ggtree::ggtree(hclust(jacdist, method = "average")) %<+% anno
    p$data
  })
```

```{r}
plist[[2]]$x <- plist[[2]]$x + max(plist[[1]]$x) + 1.2
plist[[3]]$x <- plist[[3]]$x + max(plist[[2]]$x) + 1.2
dd <- dplyr::bind_rows(plist) |>
  dplyr::filter(!is.na(label))
```

##### Raw

```{r fig.height=7, fig.width=14}
p1 <- ggtree::ggtree(tr = plist[[1]]) +
  ggtree::geom_tree(data = plist[[2]]) +
  ggtree::geom_tree(data = plist[[3]]) +
  ggtree::geom_point(
    data = dd,
    mapping = ggplot2::aes(
      x, y, color = cluster_id,
      shape = contrast, size = size)) +
  ggplot2::scale_shape_discrete(name = "Contrast") +
  ggplot2::scale_x_continuous(expand = ggplot2::expansion(mult = c(0.05, 0.05))) +
  ggplot2::scale_size_continuous(range = c(2, 8), breaks = c(10, 100, 1000)) +
  ggthemes::scale_color_tableau("Tableau 10") +
  legend_override("color", list(linetype = 0, size = 3), order = 2, title = "Cell Types") +
  legend_override("shape", list(size = 3), order = 1, title = "Contrast") +
  ggplot2::labs(size = "DEG Size") +
  ggplot2::theme(legend.title = ggplot2::element_text(size = 16),
                 legend.text = ggplot2::element_text(size = 14)) +
  NULL

p1
```

##### With lines

```{r fig.height=7, fig.width=14}
p3 <- ggtree::ggtree(tr = plist[[1]]) +
  ggtree::geom_tree(data = plist[[2]]) +
  ggtree::geom_tree(data = plist[[3]]) +
  ggplot2::geom_line(
    data = dd,
    mapping = ggplot2::aes(
        x, y, color = contrast,
        group = paste(contrast, cluster_id)),
      alpha = 0.5, linetype = "dashed") +
  ggplot2::scale_color_brewer(palette = "Set1") +
  legend_override("color", list(alpha = 1), order = 2, title = "Contrast") +
  ggnewscale::new_scale_color() +  
  ggtree::geom_point(
    data = dd,
    mapping = ggplot2::aes(
      x, y, color = cluster_id,
      shape = contrast, size = size)) +
  ggplot2::scale_shape_discrete(name = "Contrast") +
  ggplot2::scale_x_continuous(expand = ggplot2::expansion(mult = c(0.05, 0.05))) +
  ggplot2::scale_size_continuous(range = c(2, 8), breaks = c(10, 100, 1000)) +
  ggthemes::scale_color_tableau("Tableau 10") +
  legend_override("color", list(linetype = 0, size = 3), order = 2, title = "Cell Types") +
  legend_override("shape", list(size = 3), order = 1, title = "Contrast") +
  ggplot2::labs(size = "DEG Size") +
  ggplot2::theme(legend.title = ggplot2::element_text(size = 16),
                 legend.text = ggplot2::element_text(size = 14)) +
  legend_override("color", list(size = 3), order = 3, title = "Cell Types") +
  NULL

p3
```

#### Cluster all sets {.tabset}

```{r}
df_anno <- df_degs |>
  dplyr::select(-regu_type) |>
  dplyr::mutate(
    time = forcats::fct_drop(time),
    cluster_id = factor(cluster_id, levels = TISSUE_TYPES),
    contrast = factor(contrast, levels = c("PNR2 vs. MOCK", "TR4 vs. MOCK"))) |>
  dplyr::group_by(time, contrast, cluster_id) |>
  dplyr::summarise(genes = list(gene), size = length(gene), .groups = "drop")# |>
  #dplyr::filter(size > 20)

df_jaxmtx <- df_anno |>
  dplyr::rename(
    time.1 = time,
    contrast.1 = contrast,
    cluster_id.1 = cluster_id) |>
  dplyr::mutate(data = lapply(
    X = seq_along(genes),
    FUN = function(idx) {
      data.frame(
        time.2 = time.1,
        contrast.2 = contrast.1,
        cluster_id.2 = cluster_id.1,
        jaccard = vapply(
          X = genes,
          FUN = function(gs)
            bayesbio::jaccardSets(genes[[idx]], gs),
          FUN.VALUE = numeric(1))
      )
    })) |>
  dplyr::select(time.1, contrast.1, cluster_id.1, data) |>
  tidyr::unnest(cols = data) |>
  dplyr::mutate(
    set.1 = paste(time.1, contrast.1, cluster_id.1, sep = " - "),
    set.2 = paste(time.2, contrast.2, cluster_id.2, sep = " - ")) |>
  dplyr::select(set.1, set.2, jaccard) |>
  tidyr::pivot_wider(names_from = set.2, values_from = jaccard) |>
  tibble::column_to_rownames("set.1")

df_anno <- dplyr::mutate(
  df_anno,
  label = paste(time, contrast, cluster_id, sep = " - "),
  .before = 1)

jacdist <- as.dist(1 - as.matrix(df_jaxmtx))
```

##### Network

```{r fig.height=6, fig.width=9}
fit <- hclust(jacdist, method = "average")
clus <- cutree(fit, 5)
grp <- split(names(clus), clus)

p2 <- ggtree::ggtree(
    fit,
    layout = igraph::layout_with_kk,
    color = "#bed0d1") %<+% df_anno

clades <- sapply(grp, \(n) ggtree::MRCA(p2, n))
p2 <- ggtree::groupClade(p2, clades, group_name = "subtree") +
  ggplot2::aes(color = subtree)

p2 <- p2 +
  ggnewscale::new_scale_color() +
  ggtree::geom_tippoint(
    mapping = ggplot2::aes(
      x, y, color = cluster_id,
      shape = contrast, size = size)) +
  ggtree::geom_rootedge() +
  ggtree::geom_rootpoint() +
  ggtree::geom_text2(
    mapping = ggplot2::aes(subset = isTip, label = substr(time, 1, 1)),
    size = 5) +
  ggplot2::scale_size_continuous(range = c(2, 8), breaks = c(10, 100, 1000)) +
  ggthemes::scale_color_tableau("Tableau 10") +
  legend_override("color", list(size = 3), order = 2, title = "Cell Types") +
  legend_override("shape", list(size = 3), order = 1, title = "Contrast") +
  ggplot2::labs(size = "DEG Size") +
  NULL

p2
```

##### Dendrogram

```{r fig.height=10, fig.width=7}
p <- ggtree::ggtree(
    hclust(jacdist, method = "average")) %<+% df_anno +
  ggtree::geom_tippoint(
    mapping = ggplot2::aes(
      x, y, color = cluster_id,
      shape = contrast, size = size)) +
  ggtree::geom_text2(
    mapping = ggplot2::aes(subset = isTip, label = substr(time, 1, 1)),
    size = 3) +
  ggplot2::scale_size_continuous(range = c(2, 8), breaks = c(10, 100, 1000)) +
  ggthemes::scale_color_tableau("Tableau 10") +
  legend_override("color", list(size = 3), order = 2, title = "Cell Types") +
  legend_override("shape", list(size = 3), order = 1, title = "Contrast") +
  ggplot2::labs(size = "DEG Size") +
  NULL

p
```

#### Make figure

```{r fig.height=6, fig.width=14}
(p <- p2 + ggplot2::theme(legend.position = "none") +
      p3 + patchwork::plot_layout(guides = "collect"))
```


### Chlorenchyma DE 1DPI

#### Me_1 subpopulation

##### Identify DEGs

```{r}
Me_1_pnr2_degs <- Seurat::FindMarkers(
  obj, ident.1 = "1DPI-PNR2", ident.2 = "1DPI-MOCK",
  group.by = "group", subset.ident = "Me_1")

Me_1_tr4_degs <- Seurat::FindMarkers(
  obj, ident.1 = "1DPI-TR4", ident.2 = "1DPI-MOCK",
  group.by = 'group', subset.ident = "Me_1")
```

```{r}
p1 <- ggplot2::ggplot(Me_1_pnr2_degs,
    ggplot2::aes(x = avg_log2FC, y = -log(p_val_adj),
                 color = pct.1 - pct.2)) +
  ggplot2::geom_point() +
  ggplot2::ggtitle("PNR2 vs. MOCK - 1DPI in Me_1") +
  ggplot2::theme_bw()

p2 <- ggplot2::ggplot(Me_1_tr4_degs,
    ggplot2::aes(x = avg_log2FC, y = -log(p_val_adj),
                 color = pct.1 - pct.2)) +
  ggplot2::geom_point() +
  ggplot2::ggtitle("TR4 vs. MOCK - 1DPI in Me_1") +
  ggplot2::theme_bw()

p1 + p2 &
  center_plot_title() &
  ggplot2::scale_color_gradient2(
    low = "blue", mid = "grey", high = "red") &
  ggplot2::theme(legend.title.align = 0.5,
                 legend.title = ggplot2::element_text(angle = 90)) &
  ggplot2::guides(colour = ggplot2::guide_colorbar(title.position = "left"))
```

##### Enrichment

```{r}
df <- list(PNR2 = tibble::rownames_to_column(Me_1_pnr2_degs, "Gene"),
     TR4 = tibble::rownames_to_column(Me_1_tr4_degs, "Gene")) |>
  tibble::enframe("Contrast", "df") |>
  tidyr::unnest(cols = df) |>
  dplyr::filter(p_val_adj < 0.05) |>
  dplyr::mutate(Regu = ifelse(avg_log2FC > 0, "UP", "DOWN")) |>
  dplyr::select(Contrast, Regu, Gene)

enrslim <- clusterProfiler::compareCluster(
    Gene ~ Contrast + Regu,
    data = df,
    fun = "enrichGO",
    OrgDb = "org.Taestivum.iwgsc.db",
    keyType = "GID", ont = "BP",
    universe = rownames(obj)) |>
  clusterProfiler::simplify(0.9) |>
  enrichRepresent()

enrslim |>
  enrichplot::dotplot(
    showCategory = NULL,
    size = "count", label_format = 50) +
  ggplot2::facet_wrap(~ Regu, scales = "free_x") +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 15, vjust = 1, hjust = 1))
```

#### Me_2 subpopulation

##### Identify DEGs

```{r}
Me_2_pnr2_degs <- Seurat::FindMarkers(
  obj, ident.1 = "1DPI-PNR2", ident.2 = "1DPI-MOCK",
  group.by = 'group', subset.ident = "Me_2")

Me_2_tr4_degs <- Seurat::FindMarkers(
  obj, ident.1 = "1DPI-TR4", ident.2 = "1DPI-MOCK",
  group.by = 'group', subset.ident = "Me_2")
```

```{r}
p1 <- ggplot2::ggplot(Me_2_pnr2_degs,
    ggplot2::aes(x = avg_log2FC, y = -log(p_val_adj),
                 color = pct.1 - pct.2)) +
  ggplot2::geom_point() +
  ggplot2::ggtitle("PNR2 vs. MOCK - 1DPI in Me_2") +
  ggplot2::theme_bw()

p2 <- ggplot2::ggplot(Me_2_tr4_degs,
    ggplot2::aes(x = avg_log2FC, y = -log(p_val_adj),
                 color = pct.1 - pct.2)) +
  ggplot2::geom_point() +
  ggplot2::ggtitle("TR4 vs. MOCK - 1DPI in Me_2") +
  ggplot2::theme_bw()

p1 + p2 &
  center_plot_title() &
  ggplot2::scale_color_gradient2(
    low = "blue", mid = "grey", high = "red") &
  ggplot2::theme(legend.title.align = 0.5,
                 legend.title = ggplot2::element_text(angle = 90)) &
  ggplot2::guides(colour = ggplot2::guide_colorbar(title.position = "left"))
```

##### Enrichment

```{r}
df <- list(
    PNR2 = tibble::rownames_to_column(Me_2_pnr2_degs, "Gene"),
    TR4 = tibble::rownames_to_column(Me_2_tr4_degs, "Gene")) |>
  tibble::enframe("Contrast", "df") |>
  tidyr::unnest(cols = df) |>
  dplyr::filter(p_val_adj < 0.05) |>
  dplyr::mutate(Regu = ifelse(avg_log2FC > 0, "UP", "DOWN")) |>
  dplyr::select(Contrast, Regu, Gene)

enrslim <- clusterProfiler::compareCluster(
    Gene ~ Contrast + Regu,
    data = df,
    fun = "enrichGO",
    OrgDb = "org.Taestivum.iwgsc.db",
    keyType = "GID", ont = "BP",
    universe = rownames(obj)) |>
  #clusterProfiler::simplify(0.9) |>
  enrichRepresent()

enrslim |>
  enrichplot::dotplot(
    showCategory = NULL,
    size = "count", label_format = 50) +
  ggplot2::facet_wrap(~ Regu, scales = "free_x") +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 15, vjust = 1, hjust = 1))
```


### RNA Velocity

```{r}
scv <- reticulate::import("scvelo")
plt <- reticulate::import("matplotlib.pyplot")
```

```{r}
adata_all_mock <- readPickle(Sys.glob(
  "../results/ObjectCache/TrajectoryInference/unitvelo_adata_mock_*.pickle"))
adata_all_pnr2 <- readPickle(Sys.glob(
  "../results/ObjectCache/TrajectoryInference/unitvelo_adata_pnr2_*.pickle"))
adata_all_tr4 <- readPickle(Sys.glob(
  "../results/ObjectCache/TrajectoryInference/unitvelo_adata_tr4_*.pickle"))
```

```{r}
sce_sub_list <- lapply(
  X = list(
    MOCK = adata_all_mock,
    PNR2 = adata_all_pnr2,
    TR4  = adata_all_tr4),
  FUN = function(adata) {
    sce_sub <- sce[, adata$obs_names$to_list()]
    adata$obsm$update(list(X_umap = reducedDim(sce_sub, "UMAP")))
    scv$tl$velocity_embedding(adata, basis = "umap")
    reducedDim(sce_sub, "velocity_umap") <- adata$obsm["velocity_umap"]
    sce_sub
  })
```

```{r fig.height=8, fig.width=14}
plist <- purrr::imap(sce_sub_list, function(sce_sub, name) {
  colnames(reducedDim(sce_sub, "UMAP")) <- NULL
  vector_df <- velociraptor::gridVectors(
      reducedDim(sce_sub, "UMAP"),
      reducedDim(sce_sub, "velocity_umap"),
      resolution = 30)
  scater::plotReducedDim(
      sce_sub,
      colour_by = "tissue",
      dimred = "UMAP",
      point_size = 0.8,
      point_alpha = 1,
      theme_size = 14) +
    ggplot2::scale_color_manual(values = all_tissue_colors) +
    ggplot2::geom_segment(
      data = vector_df,
      mapping = ggplot2::aes_string(
        x = "start.1",
        y = "start.2", 
        xend = "end.1",
        yend = "end.2"),
      arrow = grid::arrow(
        length = grid::unit(0.06, "inches"),
        angle = 25,
        type = "open"),
      size = 0.5) +
    ggplot2::coord_fixed() +
    ggplot2::ggtitle(name) +
    center_plot_title(face = "plain") +
    theme_dimred()
})

(p <- patchwork::wrap_plots(plist) +
    patchwork::plot_layout(guides = "collect") &
    ggplot2::theme(legend.position = "bottom") &
    ggplot2::guides(color = ggplot2::guide_legend(
      override.aes = list(size = 4, alpha = 1),
      title = "Cell Types", nrow = 1, byrow = TRUE))
  )
```


